### 2016-06-29 paper.tex
feito gitignore p latex

\documentclass[review]{elsarticle}

\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother
\begin{document}
%
\begin{frontmatter}
%
\title{Open Linked Social Data for Scientific Benchmarking}
%
\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
\ead{fabbri@usp.br}
%
\author[pwr]{Osvaldo Novais de Oliveira Junior\fnref{kio-url}}
\ead{chu@ifsc.usp.br}
%
\cortext[corresponding]{Corresponding author}
\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
University, Brazil}
%
\fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
%
\begin{abstract}
The field of social network analysis and the topic of complex networks
are widely researched.
Recently, a myriad of results have been reported which are based in
diverse datasets most often not accessible to other researchers.
This work exposes an open dataset with diverse provenance and oriented
to provide the scientific community a friendly and common repertoire.
Current data was obtained from Facebook, Twitter, IRC, Email and the specific
instances of ParticipaBR, AA and Cidade Democr\'atica.
These were translated to linked data format to homonenize access,
conform to current best practices and ease analyzes which integrate third
party and provided instances.
This document presents an outline and overall statistics of given
dataset which should favor subsequent work.
\end{abstract}
%
\begin{keyword}
Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networkx
%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
\end{keyword}

\end{frontmatter}

\section{Introduction}
The enormity of the digital data makes a rapid development of various methods of data analysis,
giving an opportunity to analyze the data from a different perspectives.

\subsection{Benchmarking in the analysis of complexity}
%
Every designer of a new method wants to know how good the proposed method is in comparison with others.
Usually such comparison is made by run different methods on some commonly used benchmark datasets and use some of the proposed evaluation measures, e.g.,~\cite{Dreiseitl200128,Cooper1997107,Douglas2011544,Kampichler2010441,DBLP:journals/corr/abs-1301-7401,steinbach00comparison,Kakkar2014}.
%
% karate club, whatelse?

\section{Materials: data from diverse provenance}

\section{Methods: linked open data}
\subsection{RDF}
\subsection{Data-driven ontology synthesis}

\section{Results: data outline}
\label{outline}

\section{Conclusions}
\label{conclusions}

\section*{References}
%
\bibliography{paper}
%\bibliography{myLastBibfile.bib}
%
\end{document}

### 2016-06-30 paper.tex
first outline

@@
 %
-Every designer of a new method wants to know how good the proposed method is in comparison with others.
-Usually such comparison is made by run different methods on some commonly used benchmark datasets and use some of the proposed evaluation measures, e.g.,~\cite{Dreiseitl200128,Cooper1997107,Douglas2011544,Kampichler2010441,DBLP:journals/corr/abs-1301-7401,steinbach00comparison,Kakkar2014}.
-%
-% karate club, whatelse?
+Karate club, whatelse?
 
@@
 
-\section{Results: data outline}
+\section{Results}
 \label{outline}
+\subsection{Data outline}
+\subsection{Software tools}
+\subsection{SPARQL queries}
 

### 2016-06-30 paper.tex
minor changes

@@
 \section{Introduction}
-The enormity of the digital data makes a rapid development of various methods of data analysis,
-giving an opportunity to analyze the data from a different perspectives.
+The enormity of the digital data propels a rapid development of data analysis methods,
+implying analyzes from different perspectives.
 
@@
 
-\section{Materials: data from diverse provenance}
+\section{Materials: data from diverse provenance}\label{materials}
 
-\section{Methods: linked open data}
+\section{Methods}
+
+\subsection{Linked open data}
 \subsection{RDF}

### 2016-06-30 paper.tex
minor changes

@@
 \begin{keyword}
-Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networkx
+Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
 %Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
@@
 \section{Introduction}
-The enormity of the digital data propels a rapid development of data analysis methods,
-implying analyzes from different perspectives.
+The enormity of the digital data propels a rapid development of analysis methods
+from different perspectives.
 

### 2016-07-01 paper.tex
advances in written parts

@@
 
+\usepackage{enumitem}
 \usepackage{hyperref}
@@
 \section{Introduction}
+In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
+both research and application realms.
+However, there is is a lack of datasets for research benchmarking,
+specially in the complex networks field, yielding diverse results from 
+poorly related data.
+
 The enormity of the digital data propels a rapid development of analysis methods
 from different perspectives.
+The used datasets differ within the scope of each research with
+scarce and historical exception such as the
+karate club dataset~\cite{newmanBook}.
 
-\subsection{Benchmarking in the analysis of complexity}
-%
-Karate club, whatelse?
+This work presents a linked data dataset with data from diverse
+provenance, including Facebook, Twitter, IRC, Email and detached
+instances.
+Such data is proposed as a common repertoire for scientific
+reseach involving networks and textual content.
 
-\section{Materials: data from diverse provenance}\label{materials}
+% \subsection{Benchmarking in the analysis of complexity}
+
+\section{Materials}\label{materials}
+\subsection{Facebook data}
+\subsection{Twitter data}
+\subsection{IRC data}
+\subsection{Email data}
+\subsection{ParticipaBR data}
+\subsection{AA data}
+\subsection{Cidade Democr\'atica data}
 
 \section{Methods}
+\subsection{Linked open data}
+Linked data refers to data published in the web in such a way that it is
+machine readable and conforms to a set of best practices.
+The web of data is constructed with documents on the web 
+such as the web of hypertext.
+In practice, the idea of linked data can me summarized
+by 1) the use of RDF to publish data on the web and 2) the use of RDF
+links to interlink data from different sources.
+The web is expected to be interconnected and to grow by the systematic application of four
+steps~\cite{lee1}:
+\begin{itemize}
+    \item Use URIs to identify things~\cite{uri}.
+    \item Use HTTP URIs.
+    \item Provide useful information when an URI is accessed via HTTP.
+    \item Provide other URIs in the description of resources so human
+        and machine agents can perform discovery.
+\end{itemize}
 
-\subsection{Linked open data}
+The Linked Open Data~\cite{lod} builds an ever growing cloud of data,
+the global data space, which is usually
+conceived as centered around the DBPedia, a linked data representation
+of data from Wikipedia~\cite{dbpedia0,dbpedia}.
+
 \subsection{RDF}
+The Resource Description Framework (RDF), a W3C
+recommendation, is a model for data
+interchange.
+It is based on the idea of making statements about resources in the form
+of triples, i.e. expressions in the form ``subject - predicate -
+object''.
+RDF can be serialized in several file formats, including RDF/XML,
+Turtle and Manchester all which, in essence, represent a labeled and
+directed multi-graph.
+RDF may be stored in a type of database called a triplestore~\cite{rdf}.
+
+As an example of an RDF statement, the following triple in the Turtle
+format asserts that ``the paper has color white'':\\
+\texttt{http://example.org/paper http://example.org/hasColor\\
+http://example.org/White .}
+
 \subsection{Data-driven ontology synthesis}
+OWL Ontologies are critical tools to describe taxonomies and the
+structure of knowledge.
+Most ontologies are created by domain experts even though the data they
+arrange is often given by a software system.
+
+We developed an ontology synthesis method that probes the ontological structure in data with
+SPARQL queries and post-processing which can be divided in the following steps:
+\begin{enumerate}[leftmargin=0cm]
+    \item Obtaining all distinct classes with the query:\\
+        \texttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
+    \item Obtaining all distinct properties with the query:\\
+        \texttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
+    \item For each class, get distinct subject classes and predicates where the
+        the object is an instance of the class:\\
+        \texttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
+        a ?cs . \}}
+    \item For each class, get distinct predicates and object classes or
+        datatypes where the subject is an instance of such class:\\
+        \texttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
+                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
+    \item For each property, check if it is functional, i.e. if it
+        occurs only once with each subject:\\
+        \texttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
+            <property\_uri> ?o \} GROUP BY ?s}
+    \item For each property, find the incident range and domain with the
+        queries:\\
+        \texttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
+                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}}
+        \texttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
+    \item For each instance of each class, get all distinct predicates.
+        For each predicate, check if all instances of the class
+        hold such relationship (existential restriction):\\
+        \texttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o
+        . \}}\\
+        \texttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
+        \texttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
+                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
+    \item and if all instances that hold such relationship are instances of the class
+        (universal restriction):\\
+        \texttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
+    \item Draw each class, each property and the overall figure.
+    \item Make \texttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
+        statements to better organize knowledge and link to third party
+        ontologies and data.
+    \item
+\end{enumerate}
 

### 2016-07-01 paper.tex
finished outline without results

@@
 %
-\title{Open Linked Social Data for Scientific Benchmarking}
+\title{Linked Open Social Data for Scientific Benchmarking}
 %
@@
 to provide the scientific community a friendly and common repertoire.
-Current data was obtained from Facebook, Twitter, IRC, Email and the specific
-instances of ParticipaBR, AA and Cidade Democr\'atica.
-These were translated to linked data format to homonenize access,
+Current data was obtained from Facebook, Twitter, IRC, Email and the
+detached instances of ParticipaBR, AA and Cidade Democr\'atica.
+These were represented as linked data to homonenize access,
 conform to current best practices and ease analyzes which integrate third
 party and provided instances.
-This document presents an outline and overall statistics of given
+This document presents an outline and overall statistics of the given
 dataset which should favor subsequent work.
@@
 karate club dataset~\cite{newmanBook}.
+On the other hand, the available linked data is not 
+stable or rigorous enough to be
+a public reference on statistical physics and social networks research.
 
-This work presents a linked data dataset with data from diverse
+This work presents a linked open social data (LOSD) dataset with data from diverse
 provenance, including Facebook, Twitter, IRC, Email and detached
@@
 \section{Materials}\label{materials}
+Data was gathered either from:
+\begin{itemize}
+    \item public APIs (Twitter, Email); or
+    \item public logs (IRC and AA); or
+    \item Netvizz software~\cite{netvizz} (Facebook); or
+    \item donated data from users (Facebook); or
+    \item donated data from system administrators (AA, ParticipaBR,
+        Cidade Democr\'atica).
+\end{itemize}
+
+Integration and uniformity of access is obtained through linked data
+representation, as exposed in Section~\ref{queries}.
+
+Of central importance to presented LOSD is the concept of a snapshot.
+A snapshot is a set of data gathered together, at a contiguous time
+unit.
+Examples: the first 20 thousand email messages of an email list
+comprises a snapshot; the tweets from the MAMA event is a
+snapshot; the friendship, interaction and posts structures of a facebook
+group, prospected at the same time, is a snapshot.
+
 \subsection{Facebook data}
+Friendship ego networks (networks whose reference is an user)
+were donated from individual users in 2013 and 2014.
+Friendship and interaction networks from groups were gathered from
+groups where the first author was a participant.
+Additionally, some groups have post texts along some metadata, such as
+the number of likes.
+
 \subsection{Twitter data}
+Tweets were gathered through the streaming public API.
+Each snapshot is unified by a distinct hashtag.
+Edges are canonically yield by retweets but replies and user mentions
+are also kept in the LOSD.
+
 \subsection{IRC data}
+Public IRC logs were used to render LOSD IRC snapshots.
+LOSD has record of users to which the message is directed to or
+mentions.
+
 \subsection{Email data}
+Email snapshots refer to individual email lists.
+All messages were taken from the Gmane public email~\cite{gmane}.
+Each message has the original text and the text without some of the lines
+from previous messages or that are pasted software code.
+Most importantly, each message item holds the ID of the message it is
+a reply to, if any.
+
 \subsection{ParticipaBR data}
+The ParticipaBR is a platform for social participation once regarded as
+the Brazilian portal of social participation.
+Texts are derived from blog posts and networks are derived from
+friendship and interaction criteria.
+
 \subsection{AA data}
+The Algorithmic Autoregulation~\cite{aa} is a methodology for testifying
+and sharing ongoing work.
+The data was gathered from different versions of the system and from IRC
+logs and is presented as part of the LOSD as one of the detached
+platforms.
+
 \subsection{Cidade Democr\'atica data}
+Cidade Democr\'atica is a civil society social participation portal.
 
 \section{Methods}
+Data in the presented LOSD is represented as linked open data through
+RDF and ontologicaly described through a data-driven ontology synthesis
+method.
+These steps are described in the following sections.
+
 \subsection{Linked open data}
@@
         ontologies and data.
-    \item
 \end{enumerate}
@@
 \subsection{Data outline}
+% stats:
+%% number of chars
+%% number of users
+%% number of edges
+%% number of triples
+% diagrams
 \subsection{Software tools}
-\subsection{SPARQL queries}
+% triplification routines
+% analysis routines
+\subsection{SPARQL queries}\label{queries}
+% all text from user
+% users with the most friendships
+% users with most interactions
+% search string in messages
 
@@
 \label{conclusions}
+The Linked Open Social Data (LOSD) presented in this article
+should be available online in the \url{http://linkedopensocialdata.org}
+address in near future to fulfill the purpose of being a common
+repertoire in current research.
 

### 2016-07-13 paper.tex
first version with snapshot tables
first version with tables

@@
 \usepackage[T1]{fontenc}
+% \usepackage{changepage}
+\usepackage{longtable}
+\usepackage{tabularx}
+% \usepackage[showframe=true]{geometry}
 %
@@
 \begin{keyword}
-Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
+Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
 %Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
@@
 \label{outline}
+Knowledge Discovery in Databases finished untill data preparation phase.
+\subsection{Standardization}
+Standards: snapshot ids: <platform>-legacy-<further\_identifier>;
+ex: irc-legacy-labmacambira; email-legacy-c++;
+facebook-legacy-RenatoFabbri, etc., participants and transactions linkage to
+snapshots.
+
+Participation ontology URIs. Concept URIs are the same for different
+snapshot provenance, being differetiable by the snapshot metadata.
+
+Big data.
 \subsection{Data outline}
+Relations only (Facebook), text only (AA, Cidade Democratica).
+\input{misc/overallText}
+\input{tables/nsnapshots}
+\input{tables/basicOverall}
 % stats:
+%% number of triples
+%% number of edges
 %% number of chars
 %% number of users
-%% number of edges
-%% number of triples
 % diagrams
@@
 % analysis routines
+% envisioned multimedia routines    
 \subsection{SPARQL queries}\label{queries}

### 2016-07-13 paper.tex
added table about provenance

@@
 %% number of users
-% diagrams
+% diagrams in the supporting information file
 \subsection{Software tools}
-% triplification routines
-% analysis routines
-% envisioned multimedia routines    
+The LOSD is released with a software for rendering itself, analyses and
+multimedia artifacts.
+\subsubsection{Triplification routines}
+For each social platform there is a \emph{triplification} routine,
+i.e. a script for translating data to RDF.
+Original formats and further observations are presented in
+Table~\ref{tab:provenance}.
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{Social platforms, original formats and further observations for
+the LOSD dataset.}
+\begin{tabular}{| l || c | c |}\hline
+    \textbf{social platform} & \textbf{original format} & \textbf{further observations} \\\hline\hline
+    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users \\\hline
+    Cidade Democrática & MySQL database & donated by admins \\\hline
+    Email & mbox & obtained through Gmane public database \\\hline
+    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} \\\hline
+    IRC & plain text log & obtained through Supybot logging \\\hline
+    ParticipaBR & PostgreSQL database & donated by admins \\\hline
+    Twitter & Json & obtained through Twitter streaming API \\\hline
+\end{tabular}\end{center}
+\end{table*}                    
+\subsubsection{Topological and textual Analysis}
+\subsubsection{Multimedia rendering}
 \subsection{SPARQL queries}\label{queries}
@@
 repertoire in current research.
+One should access \url{http://wiki.nosdigitais.teia.org.br/LOSD}
+for reaching the address where LOSD is currently reachable.
 

### 2016-07-18 paper.tex
advances in results

--- 2016-07-13 paper.tex
@@
 %
+\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
 \pgfplotsset{compat=1.5}
@@
     \item Obtaining all distinct classes with the query:\\
-        \texttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
+        \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
     \item Obtaining all distinct properties with the query:\\
-        \texttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
+        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
     \item For each class, get distinct subject classes and predicates where the
         the object is an instance of the class:\\
-        \texttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
+        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
         a ?cs . \}}
@@
         datatypes where the subject is an instance of such class:\\
-        \texttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
+        \textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
                 a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
@@
         occurs only once with each subject:\\
-        \texttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
+        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
             <property\_uri> ?o \} GROUP BY ?s}
@@
         queries:\\
-        \texttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
-                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}}
-        \texttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
+        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
+                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}} \\
+        and \\
+        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
     \item For each instance of each class, get all distinct predicates.
@@
         hold such relationship (existential restriction):\\
-        \texttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o
-        . \}}\\
-        \texttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
-        \texttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
+        \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
+        \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
+        \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
                 a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
@@
         (universal restriction):\\
-        \texttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
+        \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
     \item Draw each class, each property and the overall figure.
-    \item Make \texttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
+    \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
         statements to better organize knowledge and link to third party
@@
 %% number of users
-% diagrams
+% diagrams in the supporting information file
 \subsection{Software tools}
-% triplification routines
-% analysis routines
-% envisioned multimedia routines    
+The LOSD is released with a software for rendering itself, analyses and
+multimedia artifacts.
+\subsubsection{Triplification routines}
+For each social platform there is a \emph{triplification} routine,
+i.e. a script for translating data to RDF.
+Original formats and further observations are presented in
+Table~\ref{tab:provenance}.
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{Social platforms, original formats and further observations for
+the LOSD dataset.}\label{tab:provenance}
+\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
+    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
+    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation \\\hline
+    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
+    Email & mbox & obtained through Gmane public database & Gmane \\\hline
+    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social \\\hline
+    IRC & plain text log & obtained through Supybot logging & Social \\\hline
+    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
+    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
+\end{tabular}\end{center}
+\end{table*}                    
+\subsubsection{Topological and textual Analysis}\label{ana}
+Routines are available for taking topological and textual measures from
+the dataset.
+Principal Component Analysis (PCA) is performed with available measures
+to ease pattern recognition both in a timeline and in a multiscale
+fashion.
+
+\subsubsection{Multimedia rendering}\label{media}
+It is a core purpose of LOSD framework to provide routines for rendering
+audiovisualizations of the LOSD data.
+Social structures are rendered into music, images and video animations
+through the Percolation toolbox~\cite{percolation} in association with
+the Music and Visuals toolboxes~\cite{music,visuals}.
+
+\subsubsection{Migration from deprecated toolboxes}
+Routines mentioned in Sections~\ref{ana} and~\ref{media} are being migrated from deprecated
+toolboxes~\cite{gmaneLegacy,percolationLegacy} into newly designed
+toolboxes~\cite{percolation,visuals}.
+
 \subsection{SPARQL queries}\label{queries}

### 2016-07-18 tables/nsnapshots.tex
advances in results

\begin{table*}[h!]
\begin{center}
\caption{Number of snapshots from each provenance. Every snapshot is a \texttt{po:Snapshot}; there are three types of the \texttt{po:AASnapshot} class.}
\begin{tabular}{| l | c |}\hline
\textbf{snapshot provenance} & \textbf{number of snapshots} \\\hline\hline
http://purl.org/socialparticipation/po/AAIRCSnapshot & 5 \\\hline
http://purl.org/socialparticipation/po/AAMongoSnapshot & 1 \\\hline
http://purl.org/socialparticipation/po/AAMysqlSnapshot & 1 \\\hline
http://purl.org/socialparticipation/po/AASnapshot & 7 \\\hline
http://purl.org/socialparticipation/po/CidadeDemocraticaSnapshot & 1 \\\hline
http://purl.org/socialparticipation/po/FacebookSnapshot & 88 \\\hline
http://purl.org/socialparticipation/po/IRCSnapshot & 4 \\\hline
http://purl.org/socialparticipation/po/ParticipabrSnapshot & 1 \\\hline
http://purl.org/socialparticipation/po/TwitterSnapshot & 16 \\\hline\hline
http://purl.org/socialparticipation/po/Snapshot & 116 \\\hline
\end{tabular}\end{center}
\end{table*}

### 2016-07-18 misc/overallText.tex
add of aux file

Database consists of 37467423 triples, 3088219 edges yield by interactions or relations, 382735 participants and 229008584 characters. Among all snapshots, 63 are ego snapshots, 53 are group snapshots; 45 have interaction edges, 89 have friendship edges; 42 have text content from messages.

### 2016-07-19 paper.tex
advances for first preliminary version

@@
 This work exposes an open dataset with diverse provenance and oriented
-to provide the scientific community a friendly and common repertoire.
+to furnish the scientific community with a friendly and common repertoire.
 Current data was obtained from Facebook, Twitter, IRC, Email and the
@@
 both research and application realms.
-However, there is is a lack of datasets for research benchmarking,
+However, there is a lack of datasets for research benchmarking,
 specially in the complex networks field, yielding diverse results from 
@@
 from different perspectives.
-The used datasets differ within the scope of each research with
-scarce and historical exception such as the
+However, the used datasets in the literature differ within the scope of each research with
+scarce and historical exceptions such as the
 karate club dataset~\cite{newmanBook}.
@@
 
-This work presents a linked open social data (LOSD) dataset with data from diverse
+This work presents a linked open social data (LOSD) dataset with diverse
 provenance, including Facebook, Twitter, IRC, Email and detached
@@
     \item Netvizz software~\cite{netvizz} (Facebook); or
-    \item donated data from users (Facebook); or
-    \item donated data from system administrators (AA, ParticipaBR,
+    \item donated data by users (Facebook); or
+    \item donated data by system administrators (AA, ParticipaBR,
         Cidade Democr\'atica).
@@
 \label{outline}
-Knowledge Discovery in Databases finished untill data preparation phase.
+Current overall results concern data selection and preparation for knowledge discovery.
+The main result is the data made available, which enables benchmarking of scientific results
+and easy experimentations.
+Secondary results include software support and SparQL queries made beforehand.
+
 \subsection{Standardization}
-Standards: snapshot ids: <platform>-legacy-<further\_identifier>;
-ex: irc-legacy-labmacambira; email-legacy-c++;
-facebook-legacy-RenatoFabbri, etc., participants and transactions linkage to
-snapshots.
+The LOSD is a collection of data translated into RDF and embedded into standard URIs and triples.
+Such standards constitute the LOSD itself, but some of the conventions are outlined in this section.
 
-Participation ontology URIs. Concept URIs are the same for different
-snapshot provenance, being differetiable by the snapshot metadata.
+URIs have the root \url{http://purl.org/socialparticipation/participationontology/} which are identified with
+the radical \textttt{po:}.
+Classes and properties are build by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
+Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
+All participants and transactions (messages, friendships, interactions) are linked to
+snapshots through the triple \textttt{<resource> po:snapshot <snapshot\_uri>}.
+Message texts, including comments, are objectives in the triple: \textttt{<message\_id> po:text <message\_text>}.
+Individuals (also called instances) are built on top of the class they derive from plus a hashtag character,
+the snapshot id of the snapshot they refer to, and an identifier;
+i.e. \textttt{po:Participant\#<snapshot\_id>-<local\_id>}.
+All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
+All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
+\textttt{irc-legacy-labmacambira} or \textttt{email-legacy-c++...}.
 
-Big data.
 \subsection{Data outline}
-Relations only (Facebook), text only (AA, Cidade Democratica).
 \input{misc/overallText}
+
+\input{tables/basicOverall}
+
 \input{tables/nsnapshots}
-\input{tables/basicOverall}
 % stats:
@@
 \subsection{SPARQL queries}\label{queries}
-% all text from user
-% users with the most friendships
-% users with most interactions
-% search string in messages
+There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
+Here we write some of the most basic of such queries selected by their potential to be varied.
+All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
+\begin{enumerate}[leftmargin=0cm]
+	\item Retrieve the number of participants
+	\item Retrieve the number of relations, be them interactions of friendships
+	\item Retrieve the number of relations, be them interactions of friendships
+	\item Retrieve all text produced by an specific user
+	\item List 100 users with the most friendships
+	\item List 100 users with most interactions
+	\item Search string in messages
+\end{enumerate}
 

### 2016-07-19 paper.tex
finished example sparql queries

@@
 \begin{enumerate}[leftmargin=0cm]
-	\item Retrieve the number of participants
-	\item Retrieve the number of relations, be them interactions of friendships
-	\item Retrieve the number of relations, be them interactions of friendships
-	\item Retrieve all text produced by an specific user
-	\item List 100 users with the most friendships
-	\item List 100 users with most interactions
-	\item Search string in messages
+	\item Retrieve the number of participants:\\
+            \textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
+            ?author a po:Participant . \} }
+	\item Retrieve the number of relations, be them interactions or
+            friendships:\\
+            \textttt{ SELECT (COUNT(?interaction) as ?c) WHERE \{ \{
+                    ?interaction a po:Friendship \} UNION \{ ?interaction
+                    a po:Interaction \} UNION \{ ?interaction po:retweetOf
+                    ?message \} UNION \{ ?interaction po:replyTo ?message
+                    \} UNION \{ ?interaction po:directedTo ?participant
+                    \} \} }
+	\item Retrieve all text produced by an specific user:\\
+            \textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{ ?activity
+            po:author <user\_uri> . ?activity po:text ?text . \}}
+	\item List users with the most friendships and the number of
+            friendships in descending order by the number of friendships:\\
+            \textttt{SELECT DISTINCT ?participant (COUNT(?interaction)
+            as ?c) WHERE \{ ?interaction a po:Friendship . ?interaction
+        po:member ?participant . \} ORDER BY DESC(?c)}
+	\item Search string ``pineapple'' in LOSD messages:\\
+            \textttt{SELECT ?text WHERE \{ ?activity po:text ?text .
+            FILTER regex(?text, 'pineapple', 'i') \}}
 \end{enumerate}

### 2016-07-21 paper.tex
advances in query formating

--- 2016-07-19 paper.tex
@@
 \newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
+\newcommand{\h} {\hphantom ~ }
+% \newcommand{\textttt}[1] {\mbox{\texttt{\footnotesize#1}}}
+% \newcommand{\textttt}[1] {
+% \begin{verbatim} #1 \end{verbatim}
+% }
 \pgfplotsset{compat=1.5}
@@
 We developed an ontology synthesis method that probes the ontological structure in data with
-SPARQL queries and post-processing which can be divided in the following steps:
+SPARQL queries and post-processing.
+The results are OWL code and figures which are available in the
+Supporting Information file.
+The method can be improved to comprise further OWL axioms,
+but is currently performed to fit present needs as follows:
 \begin{enumerate}[leftmargin=0cm]
     \item Obtaining all distinct classes with the query:\\
+        % \textttt{SELECT DISTINCT ?class WHERE { ?s a ?class }}
         \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
@@
 \begin{enumerate}[leftmargin=0cm]
-	\item Retrieve the number of participants
-	\item Retrieve the number of relations, be them interactions of friendships
-	\item Retrieve the number of relations, be them interactions of friendships
-	\item Retrieve all text produced by an specific user
-	\item List 100 users with the most friendships
-	\item List 100 users with most interactions
-	\item Search string in messages
+	\item Retrieve the number of participants:\\
+            \textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
+            ?author a po:Participant . \} }
+	\item Retrieve the number of relations, be them interactions or
+            friendships:\\
+            \textttt{SELECT (COUNT(?interaction) as ?c) WHERE \{\\
+                    \h \{ ?interaction a po:Friendship \} UNION \{ ?interaction
+                    a po:Interaction \} UNION\\
+                    \h \{ ?interaction po:retweetOf
+                    ?message \} UNION \{ ?interaction po:replyTo ?message
+                    \}\\
+                    \h UNION \{ ?interaction po:directedTo ?participant
+                    \}\\ \} }
+	\item Retrieve all text produced by an specific user:\\
+            \textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{\\
+                    \h ?activity po:author <user\_uri> . ?activity po:text ?text .\\
+    \}}
+        \item List 1000 users (URIs and names) with the most friendships and the number of
+            friendships in descending order by the number of friendships:\\
+            \textttt{SELECT DISTINCT ?participant (COUNT(?interaction)
+            as ?c) WHERE \{\\
+                \h ?interaction a po:Friendship . ?interaction po:member ?participant . \\
+            \} ORDER BY DESC(?c) LIMIT 1000}
+	\item Search string ``pineapple'' in LOSD messages:\\
+            \textttt{SELECT ?text WHERE \{ \\
+                    \h ?activity po:text ?text . FILTER regex(?text, 'pineapple', 'i')\\
+    \}}
 \end{enumerate}

### 2016-07-25 paper.tex
updating to change workstation

@@
 Recently, a myriad of results have been reported which are based in
-diverse datasets most often not accessible to other researchers.
+diverse datasets most often not accessible to researchers other than the publishing authors.
 This work exposes an open dataset with diverse provenance and oriented
@@
 both research and application realms.
-However, there is a lack of datasets for research benchmarking,
+However, there is a lack of datasets for benchmarking results,
 specially in the complex networks field, yielding diverse results from 
@@
 Such data is proposed as a common repertoire for scientific
-reseach involving networks and textual content.
+research involving networks and textual content.
 
@@
 \section{Materials}\label{materials}
-Data was gathered either from:
+Data was gathered from:
 \begin{itemize}
-    \item public APIs (Twitter, Email); or
-    \item public logs (IRC and AA); or
-    \item Netvizz software~\cite{netvizz} (Facebook); or
-    \item donated data by users (Facebook); or
-    \item donated data by system administrators (AA, ParticipaBR,
-        Cidade Democr\'atica).
+    \item public APIs (Twitter, Email);
+    \item public logs (IRC and AA);
+    \item Netvizz software~\cite{netvizz} (Facebook);
+    \item donated data by users (Facebook);
+    \item donated data by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
 \end{itemize}
@@
 
-Of central importance to presented LOSD is the concept of a snapshot.
+\subsection{Snapshots}
+Of central importance to presented dataset is the concept of a snapshot.
 A snapshot is a set of data gathered together, at a contiguous time
@@
 Examples: the first 20 thousand email messages of an email list
-comprises a snapshot; the tweets from the MAMA event is a
+comprises a snapshot; the tweets from the MAMA music event is a
 snapshot; the friendship, interaction and posts structures of a facebook
@@
 \subsection{Twitter data}
-Tweets were gathered through the streaming public API.
+Tweets were gathered through the Twitter streaming public API.
 Each snapshot is unified by a distinct hashtag.
 Edges are canonically yield by retweets but replies and user mentions
-are also kept in the LOSD.
+are also kept in the dataset.
 
 \subsection{IRC data}
-Public IRC logs were used to render LOSD IRC snapshots.
-LOSD has record of users to which the message is directed to or
+Public IRC logs were used to render IRC snapshots.
+The dataset has records of users to which the message is directed to or
 mentions.
@@
 Email snapshots refer to individual email lists.
-All messages were taken from the Gmane public email~\cite{gmane}.
+All messages were taken from the Gmane public email database~\cite{gmane}.
 Each message has the original text and the text without some of the lines
@@
 
-We developed an ontology synthesis method that probes the ontological structure in data with
+We developed a simple ontology synthesis method that probes
+the ontological structure in data with
 SPARQL queries and post-processing.
@@
 Supporting Information file.
-The method can be improved to comprise further OWL axioms,
-but is currently performed to fit present needs as follows:
+The method can be extended to comprise further OWL axioms and restrictions,
+but is currently performed to fit present needs with maximum simplicity.
+Present needs are limited to informative figures and
+the steps implemented are as follows:
 \begin{enumerate}[leftmargin=0cm]
-    \item Obtaining all distinct classes with the query:\\
+    \item Obtain all distinct classes with the query:\\
         % \textttt{SELECT DISTINCT ?class WHERE { ?s a ?class }}
-        \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
-    \item Obtaining all distinct properties with the query:\\
+        \textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri \}}
+%    \item For each class, get distinct subject classes and predicates where the
+%        the object is an instance of the class:\\
+%        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
+%        a ?cs . \}}
+% \item For each class, obtain distinct predicates (\textttt{?p}) and object classes or
+% 	datatypes (\textttt{?co}, \textttt{?do}) where the subject is an instance (\textttt{?i}) of such class:\\
+% 	\textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ \\
+% 		\h ?i a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \\
+% \}}
+\item For each class, obtain the properties that occur as predicates in triples where the subject is an instance of the class:\\
+	\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o . \}}\\
+Such properties are used to assert existential and universal restrictions for the class.
+\item Compare the total number of individuals (\textttt{?cs1}) of the class (\textttt{class\_uri}) with
+	the number of such individuals (\textttt{?cs2}) that are subjects of at least one triple where 
+	the predicate is the property (\textttt{property\_uri}).
+	If the numbers match, there is an existential restriction of the class. The queries are:\\
+	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
+	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
+	\h ?s a <class\_uri>. ?s <property\_uri> ?o .\\ \}}
+\item Find the number of instances which are subjects of triples where the predicate is the property but are not instances of the class.
+	If there is zero of such instances, there is an universal restriction:\\
+	\textttt{SELECT (COUNT(DISTINCT ?s)=0 as ?cs) WHERE \{\\
+	\h ?s <property\_uri> ?o . ?s a ?ca . FILTER(str(?ca) != 'class\_uri')\\ \}}
+\item To keep a record of the restrictions (and occurring triples), get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
+	\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
+		\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
+\}}
+    \item Obtain all distinct properties:\\
         \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
-    \item For each class, get distinct subject classes and predicates where the
-        the object is an instance of the class:\\
-        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
-        a ?cs . \}}
-    \item For each class, get distinct predicates and object classes or
-        datatypes where the subject is an instance of such class:\\
-        \textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
-                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
-    \item For each property, check if it is functional, i.e. if it
+    \item Check if each property is functional, i.e. if it
         occurs only once with each subject:\\
@@
         queries:\\
-        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
-                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}} \\
+        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
+			\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\\}} \\
         and \\
         \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
-    \item For each instance of each class, get all distinct predicates.
-        For each predicate, check if all instances of the class
-        hold such relationship (existential restriction):\\
-        \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
-        \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
-        \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
-                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
-    \item and if all instances that hold such relationship are instances of the class
-        (universal restriction):\\
-        \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
-    \item Draw each class, each property and the overall figure.
-    \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
-        statements to better organize knowledge and link to third party
-        ontologies and data.
+%     \item For each instance of each class, get all distinct predicates.
+%         For each predicate, check if all instances of the class
+%         hold such relationship (existential restriction):\\
+%         \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
+%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
+%         \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
+%                 a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
+%     \item and if all instances that hold such relationship are instances of the class
+%         (universal restriction):\\
+%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
+%     \item Draw each class, each property and the overall figure.
+\item Draw the overall figure. This results in diagrams such as given in Figure~\ref{fig:facebook}.
+	Other diagrams are given in the Supporting Information file.
+%      \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
+%          statements to better organize knowledge and link to third party
+%          ontologies and data. This step is left aside in current LOSD as 
 \end{enumerate}

### 2016-07-25 supportingInformation.tex
finished gathering methods for rendering correct supporting information document

--- 2016-07-19 paper.tex
@@
 \documentclass[review]{elsarticle}
+% \usepackage[active, tightpage]{preview}
 
@@
 \usepackage{tabularx}
+\usepackage{pdfpages}
+\usepackage{incgraph,tikz}
 % \usepackage[showframe=true]{geometry}
@@
 \begin{abstract}
-The field of social network analysis and the topic of complex networks
-are widely researched.
-Recently, a myriad of results have been reported which are based in
-diverse datasets most often not accessible to other researchers.
-This work exposes an open dataset with diverse provenance and oriented
-to furnish the scientific community with a friendly and common repertoire.
-Current data was obtained from Facebook, Twitter, IRC, Email and the
-detached instances of ParticipaBR, AA and Cidade Democr\'atica.
-These were represented as linked data to homonenize access,
-conform to current best practices and ease analyzes which integrate third
-party and provided instances.
-This document presents an outline and overall statistics of the given
-dataset which should favor subsequent work.
 \end{abstract}
@@
 \end{frontmatter}
-
+\pdfpageheight 6in
 \section{Introduction}
-In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
-both research and application realms.
-However, there is a lack of datasets for research benchmarking,
-specially in the complex networks field, yielding diverse results from 
-poorly related data.
-
-The enormity of the digital data propels a rapid development of analysis methods
-from different perspectives.
-However, the used datasets in the literature differ within the scope of each research with
-scarce and historical exceptions such as the
-karate club dataset~\cite{newmanBook}.
-On the other hand, the available linked data is not 
-stable or rigorous enough to be
-a public reference on statistical physics and social networks research.
-
-This work presents a linked open social data (LOSD) dataset with diverse
-provenance, including Facebook, Twitter, IRC, Email and detached
-instances.
-Such data is proposed as a common repertoire for scientific
-reseach involving networks and textual content.
-
-% \subsection{Benchmarking in the analysis of complexity}
-
-\section{Materials}\label{materials}
-Data was gathered either from:
-\begin{itemize}
-    \item public APIs (Twitter, Email); or
-    \item public logs (IRC and AA); or
-    \item Netvizz software~\cite{netvizz} (Facebook); or
-    \item donated data by users (Facebook); or
-    \item donated data by system administrators (AA, ParticipaBR,
-        Cidade Democr\'atica).
-\end{itemize}
-
-Integration and uniformity of access is obtained through linked data
-representation, as exposed in Section~\ref{queries}.
-
-Of central importance to presented LOSD is the concept of a snapshot.
-A snapshot is a set of data gathered together, at a contiguous time
-unit.
-Examples: the first 20 thousand email messages of an email list
-comprises a snapshot; the tweets from the MAMA event is a
-snapshot; the friendship, interaction and posts structures of a facebook
-group, prospected at the same time, is a snapshot.
-
-\subsection{Facebook data}
-Friendship ego networks (networks whose reference is an user)
-were donated from individual users in 2013 and 2014.
-Friendship and interaction networks from groups were gathered from
-groups where the first author was a participant.
-Additionally, some groups have post texts along some metadata, such as
-the number of likes.
-
-\subsection{Twitter data}
-Tweets were gathered through the streaming public API.
-Each snapshot is unified by a distinct hashtag.
-Edges are canonically yield by retweets but replies and user mentions
-are also kept in the LOSD.
-
-\subsection{IRC data}
-Public IRC logs were used to render LOSD IRC snapshots.
-LOSD has record of users to which the message is directed to or
-mentions.
-
-\subsection{Email data}
-Email snapshots refer to individual email lists.
-All messages were taken from the Gmane public email~\cite{gmane}.
-Each message has the original text and the text without some of the lines
-from previous messages or that are pasted software code.
-Most importantly, each message item holds the ID of the message it is
-a reply to, if any.
-
-\subsection{ParticipaBR data}
-The ParticipaBR is a platform for social participation once regarded as
-the Brazilian portal of social participation.
-Texts are derived from blog posts and networks are derived from
-friendship and interaction criteria.
-
-\subsection{AA data}
-The Algorithmic Autoregulation~\cite{aa} is a methodology for testifying
-and sharing ongoing work.
-The data was gathered from different versions of the system and from IRC
-logs and is presented as part of the LOSD as one of the detached
-platforms.
-
-\subsection{Cidade Democr\'atica data}
-Cidade Democr\'atica is a civil society social participation portal.
-
-\section{Methods}
-Data in the presented LOSD is represented as linked open data through
-RDF and ontologicaly described through a data-driven ontology synthesis
-method.
-These steps are described in the following sections.
-
-\subsection{Linked open data}
-Linked data refers to data published in the web in such a way that it is
-machine readable and conforms to a set of best practices.
-The web of data is constructed with documents on the web 
-such as the web of hypertext.
-In practice, the idea of linked data can me summarized
-by 1) the use of RDF to publish data on the web and 2) the use of RDF
-links to interlink data from different sources.
-The web is expected to be interconnected and to grow by the systematic application of four
-steps~\cite{lee1}:
-\begin{itemize}
-    \item Use URIs to identify things~\cite{uri}.
-    \item Use HTTP URIs.
-    \item Provide useful information when an URI is accessed via HTTP.
-    \item Provide other URIs in the description of resources so human
-        and machine agents can perform discovery.
-\end{itemize}
-
-The Linked Open Data~\cite{lod} builds an ever growing cloud of data,
-the global data space, which is usually
-conceived as centered around the DBPedia, a linked data representation
-of data from Wikipedia~\cite{dbpedia0,dbpedia}.
-
-\subsection{RDF}
-The Resource Description Framework (RDF), a W3C
-recommendation, is a model for data
-interchange.
-It is based on the idea of making statements about resources in the form
-of triples, i.e. expressions in the form ``subject - predicate -
-object''.
-RDF can be serialized in several file formats, including RDF/XML,
-Turtle and Manchester all which, in essence, represent a labeled and
-directed multi-graph.
-RDF may be stored in a type of database called a triplestore~\cite{rdf}.
-
-As an example of an RDF statement, the following triple in the Turtle
-format asserts that ``the paper has color white'':\\
-\texttt{http://example.org/paper http://example.org/hasColor\\
-http://example.org/White .}
-
-\subsection{Data-driven ontology synthesis}
-OWL Ontologies are critical tools to describe taxonomies and the
-structure of knowledge.
-Most ontologies are created by domain experts even though the data they
-arrange is often given by a software system.
-
-We developed an ontology synthesis method that probes the ontological structure in data with
-SPARQL queries and post-processing which can be divided in the following steps:
-\begin{enumerate}[leftmargin=0cm]
-    \item Obtaining all distinct classes with the query:\\
-        \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
-    \item Obtaining all distinct properties with the query:\\
-        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
-    \item For each class, get distinct subject classes and predicates where the
-        the object is an instance of the class:\\
-        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
-        a ?cs . \}}
-    \item For each class, get distinct predicates and object classes or
-        datatypes where the subject is an instance of such class:\\
-        \textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
-                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
-    \item For each property, check if it is functional, i.e. if it
-        occurs only once with each subject:\\
-        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
-            <property\_uri> ?o \} GROUP BY ?s}
-    \item For each property, find the incident range and domain with the
-        queries:\\
-        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
-                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}} \\
-        and \\
-        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
-    \item For each instance of each class, get all distinct predicates.
-        For each predicate, check if all instances of the class
-        hold such relationship (existential restriction):\\
-        \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
-        \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
-        \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
-                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
-    \item and if all instances that hold such relationship are instances of the class
-        (universal restriction):\\
-        \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
-    \item Draw each class, each property and the overall figure.
-    \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
-        statements to better organize knowledge and link to third party
-        ontologies and data.
-\end{enumerate}
-
-\section{Results}
-\label{outline}
-Current overall results concern data selection and preparation for knowledge discovery.
-The main result is the data made available, which enables benchmarking of scientific results
-and easy experimentations.
-Secondary results include software support and SparQL queries made beforehand.
-
-\subsection{Standardization}
-The LOSD is a collection of data translated into RDF and embedded into standard URIs and triples.
-Such standards constitute the LOSD itself, but some of the conventions are outlined in this section.
-
-URIs have the root \url{http://purl.org/socialparticipation/participationontology/} which are identified with
-the radical \textttt{po:}.
-Classes and properties are build by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
-Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
-All participants and transactions (messages, friendships, interactions) are linked to
-snapshots through the triple \textttt{<resource> po:snapshot <snapshot\_uri>}.
-Message texts, including comments, are objectives in the triple: \textttt{<message\_id> po:text <message\_text>}.
-Individuals (also called instances) are built on top of the class they derive from plus a hashtag character,
-the snapshot id of the snapshot they refer to, and an identifier;
-i.e. \textttt{po:Participant\#<snapshot\_id>-<local\_id>}.
-All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
-All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
-\textttt{irc-legacy-labmacambira} or \textttt{email-legacy-c++...}.
-
-\subsection{Data outline}
-\input{misc/overallText}
-
-\input{tables/basicOverall}
-
-\input{tables/nsnapshots}
-% stats:
-%% number of triples
-%% number of edges
-%% number of chars
-%% number of users
-% diagrams in the supporting information file
-\subsection{Software tools}
-The LOSD is released with a software for rendering itself, analyses and
-multimedia artifacts.
-\subsubsection{Triplification routines}
-For each social platform there is a \emph{triplification} routine,
-i.e. a script for translating data to RDF.
-Original formats and further observations are presented in
-Table~\ref{tab:provenance}.
-\begin{table*}[h!]\scriptsize
-\begin{center}
-\caption{Social platforms, original formats and further observations for
-the LOSD dataset.}\label{tab:provenance}
-\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
-    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
-    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation \\\hline
-    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
-    Email & mbox & obtained through Gmane public database & Gmane \\\hline
-    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social \\\hline
-    IRC & plain text log & obtained through Supybot logging & Social \\\hline
-    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
-    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
-\end{tabular}\end{center}
-\end{table*}                    
-\subsubsection{Topological and textual Analysis}\label{ana}
-Routines are available for taking topological and textual measures from
-the dataset.
-Principal Component Analysis (PCA) is performed with available measures
-to ease pattern recognition both in a timeline and in a multiscale
-fashion.
-
-\subsubsection{Multimedia rendering}\label{media}
-It is a core purpose of LOSD framework to provide routines for rendering
-audiovisualizations of the LOSD data.
-Social structures are rendered into music, images and video animations
-through the Percolation toolbox~\cite{percolation} in association with
-the Music and Visuals toolboxes~\cite{music,visuals}.
-
-\subsubsection{Migration from deprecated toolboxes}
-Routines mentioned in Sections~\ref{ana} and~\ref{media} are being migrated from deprecated
-toolboxes~\cite{gmaneLegacy,percolationLegacy} into newly designed
-toolboxes~\cite{percolation,visuals}.
-
-\subsection{SPARQL queries}\label{queries}
-There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
-Here we write some of the most basic of such queries selected by their potential to be varied.
-All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
-\begin{enumerate}[leftmargin=0cm]
-	\item Retrieve the number of participants:\\
-            \textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
-            ?author a po:Participant . \} }
-	\item Retrieve the number of relations, be them interactions or
-            friendships:\\
-            \textttt{ SELECT (COUNT(?interaction) as ?c) WHERE \{ \{
-                    ?interaction a po:Friendship \} UNION \{ ?interaction
-                    a po:Interaction \} UNION \{ ?interaction po:retweetOf
-                    ?message \} UNION \{ ?interaction po:replyTo ?message
-                    \} UNION \{ ?interaction po:directedTo ?participant
-                    \} \} }
-	\item Retrieve all text produced by an specific user:\\
-            \textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{ ?activity
-            po:author <user\_uri> . ?activity po:text ?text . \}}
-	\item List users with the most friendships and the number of
-            friendships in descending order by the number of friendships:\\
-            \textttt{SELECT DISTINCT ?participant (COUNT(?interaction)
-            as ?c) WHERE \{ ?interaction a po:Friendship . ?interaction
-        po:member ?participant . \} ORDER BY DESC(?c)}
-	\item Search string ``pineapple'' in LOSD messages:\\
-            \textttt{SELECT ?text WHERE \{ ?activity po:text ?text .
-            FILTER regex(?text, 'pineapple', 'i') \}}
-\end{enumerate}
+% \includepdf{ontologies/aairc.ttl/draw.pdf}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Paper sized to picture};}
+    paper=graphics
+    ]{ontologies/aairc.ttl/draw.png}
 
@@
 \label{conclusions}
+\pdfpageheight 2in
 The Linked Open Social Data (LOSD) presented in this article

### 2016-07-26 supportingInformation.tex
added starting version of supporting information

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black},
+    pdfborder={0 0 0}
+}
 \usepackage{multirow}
@@
 \usepackage{incgraph,tikz}
+\usepackage[titles]{tocloft}
+\setlength{\cftbeforesecskip}{-.5ex}
+\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
+
 % \usepackage[showframe=true]{geometry}
@@
 %
-\title{Linked Open Social Data for Scientific Benchmarking}
+\title{Linked Open Social Data for Scientific Benchmarking (Supporting
+Information document)}
 %
@@
 \begin{abstract}
+This is a Supporting Information document which exposes ontological
+diagrams and auxiliary tables for the Linked Open Social Data (LOSD)
+database. The main article is in~\cite{losd}.
 \end{abstract}
@@
 \end{frontmatter}
-\pdfpageheight 6in
-\section{Introduction}
+\pdfpageheight 8in
+\tableofcontents
+\section{General guidance}
+In this document we provide diagrams
+the provenances in the LOSD:
+Facebook, Twitter, IRC, Email, ParticipaBR, Cidade Democrática and AA.
+Each provenance diagram was broken in two, one present the relations
+among main classes and datatypes, the other present metadata on the
+snapshot.
+
+Furthermore, this document ends with two sets of tables, one with countings of
+triples, participants, edges/interactions/relations and characters,
+the other with references for snapshot groups, such as wikipedia or
+contact links.
+
+\section{Facebook data}
+
 % \includepdf{ontologies/aairc.ttl/draw.pdf}
 \incgraph[
-  overlay={\node[red,below right] at (page.north west) {\Huge Paper sized to picture};}
+  overlay={\node[red,below right] at (page.north west) {\Huge Facebook};}
     paper=graphics
-    ]{ontologies/aairc.ttl/draw.png}
+][scale=.6]{ontologies/facebook-legacy-Auricultura10042013Friendship.ttl/draw.png}
 
-\section{Conclusions}
-\label{conclusions}
-\pdfpageheight 2in
-The Linked Open Social Data (LOSD) presented in this article
-should be available online in the \url{http://linkedopensocialdata.org}
-address in near future to fulfill the purpose of being a common
-repertoire in current research.
-One should access \url{http://wiki.nosdigitais.teia.org.br/LOSD}
-for reaching the address where LOSD is currently reachable.
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Facebook};}
+    paper=graphics
+][scale=.5]{ontologies/facebook-legacy-Auricultura10042013Meta.ttl/draw.png}
+
+\section{Twitter data}
+\pdfpageheight 4in
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge
+  Twitter };}
+    paper=graphics
+][scale=.5]{ontologies/twitter-legacy-arenaNETmundialTweet00000.ttl/draw.png}
+
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Twitter};}
+    paper=graphics
+][scale=.5]{ontologies/twitter-legacy-arenaNETmundialMeta.ttl/draw.png}
+
+\section{IRC data}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge IRC};}
+    paper=graphics
+][scale=.7]{ontologies/irc-legacy-hackerspace-cpsLog00000.ttl/draw.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge IRC};}
+    paper=graphics
+][scale=.6]{ontologies/irc-legacy-hackerspace-cpsMeta.ttl/draw.png}
+
+\section{Email data}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Email};}
+    paper=graphics
+][scale=.4]{ontologies/gmane-legacy-.linux.audio.devel1-20000Email00000.ttl/draw.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Email};}
+    paper=graphics
+][scale=.5]{ontologies/gmane-legacy-.linux.audio.devel1-20000Meta.ttl/draw.png}
+
+\section{ParticipaBR data}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge
+  ParticipaBR};}
+    paper=graphics
+][scale=.5]{ontologies/participabr.ttl/draw.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge
+  ParticipaBR};}
+    paper=graphics
+][scale=.7]{ontologies/participabr.ttl/draw_circo.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge
+  ParticipaBR};}
+    paper=graphics
+][scale=.7]{ontologies/participabrMeta.ttl/draw.png}
+
+\section{Cidade Democrática data}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Cidade
+  Democrática};}
+    paper=graphics
+][scale=.4]{ontologies/cidadedemocratica00000.ttl/draw.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Cidade
+  Democrática};}
+    paper=graphics
+][scale=.4]{ontologies/cidadedemocratica00000.ttl/draw_circo.png}
+
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge Cidade
+  Democrática};}
+    paper=graphics
+][scale=.7]{ontologies/cidadedemocraticaMeta.ttl/draw.png}
+\section{AA data}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge AA};}
+    paper=graphics
+][scale=.6]{ontologies/aairc.ttl/draw.png}
+\incgraph[
+  overlay={\node[red,below right] at (page.north west) {\Huge AA};}
+    paper=graphics
+][scale=.7]{ontologies/aaircMeta.ttl/draw.png}
+
+\section{Snapshot references}
+\label{sreferences}
+\pdfpageheight 10in
+\section{Elementary counting in snapshots}
 

### 2016-07-26 supportingInformation.tex
mainly advances in description of supporting information diagrams

@@
 diagrams and auxiliary tables for the Linked Open Social Data (LOSD)
-database. The main article is in~\cite{losd}.
+database. The main document of the article is in~\cite{losd}.
 \end{abstract}
@@
 \begin{keyword}
-Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
+Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email,
+Complex Networks, Text Mining
 %Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
@@
 \end{frontmatter}
-\pdfpageheight 8in
+% \pdfpageheight 8in
 \tableofcontents
 \section{General guidance}
+% \paperheight = 200pt
 In this document we provide diagrams
-the provenances in the LOSD:
+for the provenances in the LOSD:
 Facebook, Twitter, IRC, Email, ParticipaBR, Cidade Democrática and AA.
-Each provenance diagram was broken in two, one present the relations
-among main classes and datatypes, the other present metadata on the
+Each provenance diagram was broken in two, one presents the relations
+among main classes (blue nodes) and data types (orange nodes), the other presents metadata on the
 snapshot.
+Every class instance is related to the snapshot instance
+by the triple \textttt{class\_uri po:snapshot snapshot\_uri}.
+Such triples are omitted for simplicity.
+Due to the large number of relations, the rendering of diagrams are
+automatized and displays imperfections.
+Even so, the images are useful for grasping what is in current LOSD
+and for conducting explorations.
+Edges in the diagrams have:
+\begin{itemize}
+    \item green color if representing an OWL existential
+class restriction (all individuals from the class present at least one triple with
+the property as predicate);
+    \item inverted nip if representing an OWL universal class
+        restriction (all individuals presenting triples with the
+        property as predicate are from the class);
+    \item full edges (non-dashed) if representing a functional property
+        axiom (there is at most one triple with the property as the
+        predicate for each individual).
+\end{itemize}
 
-Furthermore, this document ends with two sets of tables, one with countings of
+% \textheight = 100pt
+% \pdfpageheight 300pt
+Furthermore, this document ends with two sets of tables, one with counts of
 triples, participants, edges/interactions/relations and characters,
@@
 \section{Facebook data}
+Each Facebook snapshot is yield through either an user, from which the
+friends constitute a friendship network, or a group, which participants
+can yield friendship and interaction networks and posts information with
+text and some metadata.
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 
@@
     paper=graphics
-][scale=.5]{ontologies/twitter-legacy-arenaNETmundialMeta.ttl/draw.png}
+][scale=.7]{ontologies/twitter-legacy-arenaNETmundialMeta.ttl/draw.png}
 
@@
     paper=graphics
-][scale=.5]{ontologies/gmane-legacy-.linux.audio.devel1-20000Meta.ttl/draw.png}
+][scale=.5]{ontologies/gmane-legacy-linux.audio.devel1-20000Meta.ttl/draw.png}
 

### 2016-07-26 supportingInformation.tex
minor adds in SI

@@
 Due to the large number of relations, the rendering of diagrams are
-automatized and displays imperfections.
+automatized and displays some overlaps.
 Even so, the images are useful for grasping what is in current LOSD
@@
 \section{Facebook data}
-Each Facebook snapshot is yield through either an user, from which the
+Each Facebook snapshot is yield by either an user, from which the
 friends constitute a friendship network, or a group, which participants
@@
 \section{Twitter data}
+Each Twitter snapshot is yield by a hashtag.
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
+
 \pdfpageheight 4in

### 2016-07-27 supportingInformation.tex
advances in SI

@@
 Each Twitter snapshot is yield by a hashtag.
+Retweets (\textttt{po:retweetOf} are usually considered to yield the interactions between users.
+Users are identified through authors \textttt{po:numericID} (global as given by Twitter API)
+or \textttt{}.
+The database present also \textttt{po:replyTo} and \textttt{po:userMention}
+which might also be useful in understanding the networking.
 Further information is found on the following diagrams, the tables on
@@
 \section{IRC data}
+Each IRC snapshot is yield by an IRC channel.
+IRC messages are either server messages (e.g. join and exit channel)
+marked with \textttt{po:systemMessage true} and having an \textttt{po:impliedUser user\_uri},
+or user messages, which yield interactions through \textttt{po:directedTo} and \textttt{po:mentions} properties.
+Text messages without the user names are delivered through the \textttt{po:cleanText} property.
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 \incgraph[
@@
 \section{Email data}
+Each IRC snapshot is yield by an Email list.
+Interactions might be considered through \textttt{po:replyTo} relations
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 \incgraph[

### 2016-07-27 paper.tex
added starting version of supporting information
finished most of the supporting information document

@@
 %
-\title{Linked Open Social Data for Scientific Benchmarking}
+\title{A Linked Open Social Database for Scientific Benchmarking}
 %
@@
 
-URIs have the root \url{http://purl.org/socialparticipation/participationontology/} which are identified with
+URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/} which are identified with
 the radical \textttt{po:}.

### 2016-07-27 supportingInformation.tex
finished most of the supporting information document

@@
 \tableofcontents
+\clearpage
 \section{General guidance}
@@
 Each IRC snapshot is yield by an Email list.
-Interactions might be considered through \textttt{po:replyTo} relations
+Interactions is yield through \textttt{po:replyTo} relations
+although \textttt{po:to} and \textttt{po:cc} can also be considered.
+the email body is given by \textttt{po:text} relations while
+\textttt{po:cleanText} has the text with lines removed where they are
+trivially from previous messages or computer code.
 Further information is found on the following diagrams, the tables on
@@
 \section{ParticipaBR data}
+The ParticipaBR snapshot is yield by a data dump donated by the system
+administrators of the federal portal of social participation ParticipaBR.
+Articles can have parent articles (\textttt{po:parent}), be step of a
+collection of articles (\textttt{po:stepOf}) and be a mediation of other
+articles (\textttt{po:mediationOf}).
+Interactions are yield by comments which are \textttt{po:replyTo} other
+comments or which are made directly to an article.
+This snapshot holds also friendship structures.
+The language used is mainly Brazilian Portuguese, but English and
+Spanish are also incident.
+Due to the higher complexity of the diagram, an additional figure is
+given rendered with another layout algorithm
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 \incgraph[
@@
 \section{Cidade Democrática data}
+The Cidade Democrática snapshot is yield by a data dump donated by the system
+administrators of the civil society social participation portal Cidade
+Democrática.
+This snapshot holds a complex structure of both
+Topics/Inspirations/Observatories/Supports/Competitions/Prizes
+and of State/City/Neighborhood/Place.
+The language used is mainly Brazilian Portuguese.
+Due to the higher complexity of the diagram, an additional figure is
+given rendered with another layout algorithm
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 \incgraph[
@@
 ][scale=.7]{ontologies/cidadedemocraticaMeta.ttl/draw.png}
+
 \section{AA data}
+The AA (Algorithmic Autoregulation) snapshots are yield by a data dump donated by the system
+administrators and by a mined IRC log.
+The system pursue simplicity and most of data consists of detached
+shouts with \textttt{po:text} and \textttt{po:author}.
+Further information is found on the following diagrams, the tables on
+the end of this document or in the main document of this
+article~\cite{losd}.
 \incgraph[
@@
 \pdfpageheight 10in
-\section{Elementary counting in snapshots}
+All the Facebook snapshots are either the result of individuals who downloaded
+their data (and donated to the first author) or data downloaded from groups.
+In the first case, it is senseless to present references and the. In the second
+case, we present the group name and a link to a post in the group where
+data and figures were delivered to the group.
+
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{Different Twitter snapshots are yield by different hashtags. In this
+table we present each snapshot with the respective hashtag and a
+reference to the subject.}\label{tab:provenance}
+\begin{tabular}{| l || p{4cm} | p{4cm} | }\hline
+    \textbf{snapshot hashtag} & \textbf{observation} & \textbf{reference} \\\hline\hline
+    \#arenaNETmundial & a Brazilian discussion hub about free culture, democracy and the internet & \url{http://www.participa.br/netmundial} \\\hline
+    \#art & tweets with the generic hashtag \#art & \url{https://en.wikipedia.org/wiki/Art} \\\hline
+    \#ChennaiFloods & heavy rainfall generated by the annual northeast monsoon in November–December 2015 & \url{https://en.wikipedia.org/wiki/2015_South_Indian_floods} \\\hline
+    \#dilma & the 36th President of Brazil & \url{https://en.wikipedia.org/wiki/Dilma_Rousseff} \\\hline
+    \#ForaDilma & 2015-16 anti-government protests in Brazil & \url{https://en.wikipedia.org/wiki/2015-16_protests_in_Brazil} \\\hline
+    \#ForaCunha & 2015-16 anti-corruption protests in Brazil & \url{https://en.wikipedia.org/wiki/2015-16_protests_in_Brazil} \\\hline
+    \#fuck & tweets with the generic hashtag \#fuck & \url{https://en.wikipedia.org/wiki/Fuck} \\\hline
+    \#game & tweets with the generic hashtag \#game & \url{https://en.wikipedia.org/wiki/Game} \\\hline
+    \#god & tweets with the generic hashtag \#god & \url{https://en.wikipedia.org/wiki/God} \\\hline
+    \#MAMA2015 & the grand 2015 Mnet Asian Music Awards & \url{https://en.wikipedia.org/wiki/2015_Mnet_Asian_Music_Awards} \\\hline
+    \#music & tweets with the generic hashtag \#music & \url{https://en.wikipedia.org/wiki/Music} \\\hline
+    \#obama & the 44th President of the United States & \url{https://en.wikipedia.org/wiki/Barack_Obama} \\\hline
+    \#python & the Python programming language & \url{https://en.wikipedia.org/wiki/Python_(programming_language)} \\\hline
+    \#QuartaSemRacismoClubeSDV & an anti-racism netweaving & \url{https://twitter.com/hashtag/quartasemracismoclubesdv} \\\hline
+    \#science & tweets with the generic hashtag \#science & \url{https://en.wikipedia.org/wiki/Science} \\\hline
+    \#SnapDetremura & reference for Snapchat about a celebrated person & \url{https://twitter.com/detremura} \\\hline
+\end{tabular}\end{center}
+\end{table*}
+
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{Different IRC snapshots are yield by different channels. In this
+table we present each snapshot with the respective channel and a
+reference to the subject.}\label{tab:provenance}
+\begin{tabular}{| l || p{4cm} | p{4cm} | }\hline
+    \textbf{snapshot channel} & \textbf{observation} & \textbf{reference} \\\hline\hline
+    \#foradoeixo & a Brazilian network of culture related collectives & \url{https://pt.wikipedia.org/wiki/Fora_do_Eixo} \\\hline
+    \#hackerspace-cps & a hackerspace in Campinas, Brazil & \url{https://lhc.net.br/wiki/P%C3%A1gina_principal} \\\hline
+    \#hackerspaces-br & Brazilian hackerspaces channel & \url{https://garoa.net.br/wiki/Hackerspaces_Brasileiros} \\\hline
+    \#labmacambira & Brazilian channel for the labMacambira collective & \url{http://labmacambira.sourceforge.net/} \\\hline
+\end{tabular}\end{center}
+\end{table*}
+
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{Different Email snapshots are yield by different email lists. In this
+table we present each snapshot with the respective list and a
+reference to the subject.}\label{tab:provenance}
+\begin{tabular}{| l || p{4cm} | p{4cm} | }\hline
+    \textbf{Gmane ID} & \textbf{observation} & \textbf{reference} \\\hline\hline
+    gmane.linux.audio.users & the Linux Audio Users & \url{http://linuxaudio.org} \\\hline
+    gmane.politics.organizations.metareciclagem & a network about technology and social transformation  & \url{https://metareciclagem.github.io} \\\hline
+    gmane.linux.audio.devel & the Linux Audio Developers & \url{http://lists.linuxaudio.org/listinfo/linux-audio-dev} \\\hline
+    gmane.comp.gcc.libstdc++.devel & the C++ standard library & \url{https://gcc.gnu.org/libstdc++/} \\\hline
+\end{tabular}\end{center}
+\end{table*}
+
+\begin{table*}[h!]\scriptsize
+\begin{center}
+\caption{References for the snapshots of the detached instances
+ParticipaBR, Cidade Democrática and AA.}\label{tab:provenance}
+\begin{tabular}{| l || p{4cm} | p{3cm} | }\hline
+    \textbf{social protocol} & \textbf{observations} & \textbf{reference} \\\hline\hline
+    ParticipaBR & a Brazilian federal portal of social participation & \url{http://www.participa.br/} \\\hline
+    Cidade Demorática & a Brazilian civil society portal of social participation & \url{http://www.cidadedemocratica.org.br/} \\\hline
+    AA & the Algorithmic Autoregulation software development methodology & \cite{aarticle} \\\hline
+\end{tabular}\end{center}
+\end{table*}
+
+\clearpage
 
 \section*{References}
+
 %

### 2016-07-28 misc/overallText.tex
enhanced page sizes

@@
-Database consists of 37467423 triples, 3088219 edges yield by interactions or relations, 382735 participants and 229008584 characters.
-Among all snapshots, 63 are ego snapshots, 53 are group snapshots; 45 have interaction edges, 89 have friendship edges; 42 have text content from messages.
+The LOSD database consists of 34,120,026 triples, 3,172,927 edges yield by interactions or relations, 382,568 participants and 253,155,020 characters.
+Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 49 have interaction edges, 89 have friendship edges; 43 have text content from messages.

### 2016-07-28 supportingInformation.tex
enhanced page sizes

@@
 \end{frontmatter}
+\newcommand{\foo}{\textheight}
+\newcommand{\foobar}{\pdfpageheight}
 % \pdfpageheight 8in
@@
 
+\textheight = 2in
+\pdfpageheight 5in
 \incgraph[
@@
 
-\pdfpageheight 4in
 \incgraph[
@@
     paper=graphics
-][scale=.4]{ontologies/gmane-legacy-.linux.audio.devel1-20000Email00000.ttl/draw.png}
+][scale=.4]{ontologies/gmane-legacy-linux.audio.devel1-20000Email00000.ttl/draw.png}
+\textheight = 3in
+\pdfpageheight 6in
 \incgraph[
@@
 This snapshot holds a complex structure of both
-Topics/Inspirations/Observatories/Supports/Competitions/Prizes
+Topics/Inspirations/
+Observatories/Supports/Competitions/Prizes
 and of State/City/Neighborhood/Place.
@@
 ][scale=.4]{ontologies/cidadedemocratica00000.ttl/draw.png}
+\textheight = 2in
+\pdfpageheight 5in
 \incgraph[
@@
 ][scale=.6]{ontologies/aairc.ttl/draw.png}
+\textheight = 7in
+\pdfpageheight \foobar
 \incgraph[

### 2016-07-28 tables/nsnapshots.tex
enhanced page sizes

@@
 \begin{tabular}{| l | c |}\hline
-\textbf{snapshot provenance} & \textbf{number of snapshots} \\\hline\hline
-http://purl.org/socialparticipation/po/AAIRCSnapshot & 5 \\\hline
-http://purl.org/socialparticipation/po/AAMongoSnapshot & 1 \\\hline
-http://purl.org/socialparticipation/po/AAMysqlSnapshot & 1 \\\hline
-http://purl.org/socialparticipation/po/AASnapshot & 7 \\\hline
-http://purl.org/socialparticipation/po/CidadeDemocraticaSnapshot & 1 \\\hline
-http://purl.org/socialparticipation/po/FacebookSnapshot & 88 \\\hline
-http://purl.org/socialparticipation/po/IRCSnapshot & 4 \\\hline
-http://purl.org/socialparticipation/po/ParticipabrSnapshot & 1 \\\hline
-http://purl.org/socialparticipation/po/TwitterSnapshot & 16 \\\hline\hline
-http://purl.org/socialparticipation/po/Snapshot & 116 \\\hline
+\textbf{snapshot provenance} & \textbf{number of snapshots} \\\hline\hline\hline
+http://purl.org/socialparticipation/po/Snapshot & 117 \\\hline
 \end{tabular}\end{center}

### 2016-07-28 supportingInformation.tex
finished rendering tables

@@
 \pdfpageheight 10in
-All the Facebook snapshots are either the result of individuals who downloaded
-their data (and donated to the first author) or data downloaded from groups.
-In the first case, it is senseless to present references and the. In the second
-case, we present the group name and a link to a post in the group where
-data and figures were delivered to the group.
+\input{tables/facebookReferences}
 
@@
 \clearpage
+\section{Trivial counts in each snapshot}
+\input{tables/basicOverall}
+
+\clearpage
 

### 2016-07-28 tables/nsnapshots.tex
finished rendering tables

@@
 \begin{center}
-\caption{Number of snapshots from each provenance. Every snapshot is a \texttt{po:Snapshot}; there are three types of the \texttt{po:AASnapshot} class.}
+\caption{Number of snapshots from each provenance.}
 \begin{tabular}{| l | c |}\hline
-\textbf{snapshot provenance} & \textbf{number of snapshots} \\\hline\hline\hline
-http://purl.org/socialparticipation/po/Snapshot & 117 \\\hline
+\textbf{social protocol} & \textbf{number of snapshots} \\\hline\hline
+Algorithmic Autoregulation & 3 \\\hline
+Cidade Democrática & 1 \\\hline
+Email & 4 \\\hline
+Facebook & 88 \\\hline
+IRC & 4 \\\hline
+ParticipaBR & 1 \\\hline
+Twitter & 16 \\\hline\hline
+all & 117 \\\hline
 \end{tabular}\end{center}

### 2016-07-28 paper.tex
finished rendering tables
finished license issues

@@
 
-\input{tables/basicOverall}
-
 \input{tables/nsnapshots}
@@
 
+\section{License issues}
+% Facebook, Twitter, IRC, Gmane, Participa, CD, AA.
+The database presented in this article is released under public domain.
+Computer scripts are in public git repositories and PyPI Python packages.
+Although most data is already in open licenses (Twitter, Email, Participabr,
+Cidade Democrática, and AA data), IRC and Facebook data was collected
+and donated by the individuals which yield the data.
+This rises the the understanding of the right to study such data as the right to access the self,
+in parity with anthropological endaviors~\cite{antphy,antphy2}.
+% Of central importance to the obtainment of the data is the concept of
+% anthropological physics~\cite{antphy}
 \section{Conclusions}
@@
 %
+\bibliographystyle{plainurl}
 \bibliography{paper}

### 2016-07-29 paper.tex
finished a first version of the main document

@@
 \begin{abstract}
-The field of social network analysis and the topic of complex networks
+The fields of social network analysis and complex networks
 are widely researched.
@@
 diverse datasets most often not accessible to researchers other than the publishing authors.
-This work exposes an open dataset with diverse provenance and oriented
+This work exposes an open database with diverse provenance and oriented
 to furnish the scientific community with a friendly and common repertoire.
@@
 This document presents an outline and overall statistics of the given
-dataset which should favor subsequent work.
+database which should favor subsequent work.
 \end{abstract}
@@
 
-This work presents a linked open social data (LOSD) dataset with diverse
+This work presents a linked open social database (LOSD) with diverse
 provenance, including Facebook, Twitter, IRC, Email and detached
@@
     \item public logs (IRC and AA);
-    \item Netvizz software~\cite{netvizz} (Facebook);
-    \item donated data by users (Facebook);
-    \item donated data by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
+    \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
+    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
 \end{itemize}
@@
 representation, as exposed in Section~\ref{queries}.
+This section introduce the underlying data in very concise terms.
+One should access the Supporting Information document for
+a thorough presentation of the database.
 
 \subsection{Snapshots}
-Of central importance to presented dataset is the concept of a snapshot.
-A snapshot is a set of data gathered together, at a contiguous time
+Of central importance to presented database is the concept of a snapshot.
+A snapshot is herein a set of data gathered together, at a contiguous time
 unit.
@@
 \subsection{Facebook data}
-Friendship ego networks (networks whose reference is an user)
+Friendship ego networks (networks whose constituents are friends of an user)
 were donated from individual users in 2013 and 2014.
@@
 Edges are canonically yield by retweets but replies and user mentions
-are also kept in the dataset.
+are also kept in the database.
 
@@
 Public IRC logs were used to render IRC snapshots.
-The dataset has records of users to which the message is directed to or
+The database has records of users to which the message is directed to or
 mentions.
@@
 Email snapshots refer to individual email lists.
-All messages were taken from the Gmane public email database~\cite{gmane}.
+All messages were obtained from the Gmane public email database~\cite{gmane}.
 Each message has the original text and the text without some of the lines
-from previous messages or that are pasted software code.
-Most importantly, each message item holds the ID of the message it is
+from previous messages or that are software code.
+Most importantly, each message instance holds the ID of the message it is
 a reply to, if any.
@@
 \subsection{ParticipaBR data}
-The ParticipaBR is a platform for social participation once regarded as
-the Brazilian portal of social participation.
+The ParticipaBR is a Brazilian federal platform for social participation.
 Texts are derived from blog posts and networks are derived from
@@
 
+\subsection{Cidade Democr\'atica data}
+Cidade Democr\'atica is a Brazilian civil society social participation portal.
+Data gathered is complex in the number of types of instances.
+
 \subsection{AA data}
-The Algorithmic Autoregulation~\cite{aa} is a methodology for testifying
-and sharing ongoing work.
-The data was gathered from different versions of the system and from IRC
-logs and is presented as part of the LOSD as one of the detached
-platforms.
-
-\subsection{Cidade Democr\'atica data}
-Cidade Democr\'atica is a civil society social participation portal.
+The Algorithmic Autoregulation~\cite{aa} is a software development
+methodology based on testifying and sharing ongoing work.
+The data was gathered from different versions of the system and from an IRC
+log.
 
 \section{Methods}
-Data in the presented LOSD is represented as linked open data through
+Data is represented as linked open data through
 RDF and ontologicaly described through a data-driven ontology synthesis
 method.
-These steps are described in the following sections.
 
@@
 The web of data is constructed with documents on the web 
-such as the web of hypertext.
+such as the web of HTML documents.
 In practice, the idea of linked data can me summarized
@@
 RDF can be serialized in several file formats, including RDF/XML,
-Turtle and Manchester all which, in essence, represent a labeled and
+Turtle and Manchester, all which, in essence, represent a labeled and
 directed multi-graph.
@@
 format asserts that ``the paper has color white'':\\
-\texttt{http://example.org/paper http://example.org/hasColor\\
-http://example.org/White .}
+\texttt{http://example.org/Things\#Paper http://example.org/hasColor\\
+http://example.org/Colors\#White .}
 
@@
 Most ontologies are created by domain experts even though the data they
-arrange is often given by a software system.
+arrange is often given by a software system and has a predefined
+structure.
 
@@
 SPARQL queries and post-processing.
-The results are OWL code and figures which are available in the
-Supporting Information file.
+The results are OWL code and diagrams which are available in the
+Supporting Information document.
 The method can be extended to comprise further OWL axioms and restrictions,
@@
 	the predicate is the property (\textttt{property\_uri}).
-	If the numbers match, there is an existential restriction of the class. The queries are:\\
+	If the numbers match, there is an existential restriction for the class. The queries are:\\
 	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
@@
     \item Check if each property is functional, i.e. if it
-        occurs only once with each subject:\\
+        occurs at most once with each subject.
+        This is performed by counting the objects and further verifying
+        that they are at most one. The query is:\\
         \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
@@
 %     \item Draw each class, each property and the overall figure.
-\item Draw the overall figure. This results in diagrams such as given in Figure~\ref{fig:facebook}.
-	Other diagrams are given in the Supporting Information file.
+\item Render diagrams as exposed in the next section and in the Supporting Information file.
 %      \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
@@
 and easy experimentations.
-Secondary results include software support and SparQL queries made beforehand.
+Secondary results include data outline through figures and tables,
+software support and example SparQL queries.
 
 \subsection{Standardization}
-The LOSD is a collection of data translated into RDF and embedded into standard URIs and triples.
-Such standards constitute the LOSD itself, but some of the conventions are outlined in this section.
+The data is embedded into standard URIs and triples, i.e. translated to RDF.
+URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
+which are identified herein with the prefix \textttt{po:}.
+Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
+Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
+All class instances, such as participants, messages, friendships and
+interactions, are linked to
+snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
+Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
+Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
+More specialized predicates are used for delivering text when necessary,
+such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
+for ParticipaBR articles.
+A participant URI is unique throughout the provenance (e.g. the same for
+the same participant in all Twitter snapshots).
+To enable annotations which differ when the snapshot changes,
+\texttt{po:Observation} class instances are used in the triple
+\textttt{<participant\_uri> po:observation <observation\_uri>}.
+The observation instances are then linked to the snapshot and the
+data.
 
-URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/} which are identified with
-the radical \textttt{po:}.
-Classes and properties are build by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
-Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
-All participants and transactions (messages, friendships, interactions) are linked to
-snapshots through the triple \textttt{<resource> po:snapshot <snapshot\_uri>}.
-Message texts, including comments, are objectives in the triple: \textttt{<message\_id> po:text <message\_text>}.
-Individuals (also called instances) are built on top of the class they derive from plus a hashtag character,
-the snapshot id of the snapshot they refer to, and an identifier;
-i.e. \textttt{po:Participant\#<snapshot\_id>-<local\_id>}.
+Instances are built on top of the class they derive from plus a hashtag character,
+a provenance string (e.g. \textttt{facebook-legacy} or
+\textttt{participabr-legacy}) of the snapshot they refer to, and an identifier;
+i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
 All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
 All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
-\textttt{irc-legacy-labmacambira} or \textttt{email-legacy-c++...}.
+\textttt{irc-legacy-labmacambira} or
+\textttt{email-legacy-linux.audio.devel1-20000}.
 
@@
 \subsection{Software tools}
-The LOSD is released with a software for rendering itself, analyses and
+The database is released with software for rendering itself, analyses and
 multimedia artifacts.
@@
 \caption{Social platforms, original formats and further observations for
-the LOSD dataset.}\label{tab:provenance}
+the database.}\label{tab:provenance}
 \begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
     \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
-    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation \\\hline
+    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
     Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
-    Email & mbox & obtained through Gmane public database & Gmane \\\hline
-    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social \\\hline
+    Email & mbox & obtained through Gmane public database & Gmane~\cite{gmane} \\\hline
+    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
     IRC & plain text log & obtained through Supybot logging & Social \\\hline
@@
 \end{table*}                    
-\subsubsection{Topological and textual Analysis}\label{ana}
+\subsubsection{Topological and textual analysis}\label{ana}
 Routines are available for taking topological and textual measures from
-the dataset.
-Principal Component Analysis (PCA) is performed with available measures
-to ease pattern recognition both in a timeline and in a multiscale
-fashion.
+the database.
+Auxiliary routines, such as performing principal component analysis
+and taking Kolmogorov-Smirnov measures, are available
+to ease pattern recognition.
+Single, timeline and multi-scale analyzes are automated.
 
 \subsubsection{Multimedia rendering}\label{media}
-It is a core purpose of LOSD framework to provide routines for rendering
-audiovisualizations of the LOSD data.
+It is a core purpose of framework to provide routines for rendering
+audiovisualizations of the data.
 Social structures are rendered into music, images and video animations
@@
 
+\subsection{Diagrams of the data and auxiliary tables}
+The database exploration can be assisted through diagrams which expose
+the structure from each provenance.
+Such diagrams are in the Supporting Information document
+with some tables to ease understanding of the provided data.
+A simplified example is given in Figure~\ref{dia} where the friendship
+structure of the Facebook snapshots are exposed.
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=0.5\textwidth]{ontologies/facebook-legacy-AntonioAnzoategui18022013Friendship.ttl/draw}
+    \caption{A diagram of the structure involved in the friendship networks
+    of the Facebook snapshot.
+A green edge denotes an OWL existential class restriction; an inverted nip denotes an OWL universal class restriction; a full (non-dashed) edge denotes an OWL functional property axiom.
+Further information and complete diagrams for each provenance are in the Supporting Information document.}\label{dia}
+\end{figure}
+
 \subsection{SPARQL queries}\label{queries}
-There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
+There are numerous useful and general purpose SPARQL queries to be performed against the database.
 Here we write some of the most basic of such queries selected by their potential to be varied.
-All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
+All queries assume the use the preamble \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
 \begin{enumerate}[leftmargin=0cm]
@@
 The database presented in this article is released under public domain.
-Computer scripts are in public git repositories and PyPI Python packages.
-Although most data is already in open licenses (Twitter, Email, Participabr,
-Cidade Democrática, and AA data), IRC and Facebook data was collected
+Computer scripts are in git repositories and PyPI Python packages, also under public domain.
+Although most data is already in open licenses (Twitter, Email, Participabr, Cidade Democrática, and AA data), IRC and Facebook data was collected
 and donated by the individuals which yield the data.
@@
 \label{conclusions}
-The Linked Open Social Data (LOSD) presented in this article
-should be available online in the \url{http://linkedopensocialdata.org}
+The database presented in this article
+constitutes a large database with diverse provenance.
+Even so, database should be expanded in upon need or requests from feedback.
+All data should be available online in the \url{http://linkedopensocialdata.org}
 address in near future to fulfill the purpose of being a common
 repertoire in current research.
-One should access \url{http://wiki.nosdigitais.teia.org.br/LOSD}
-for reaching the address where LOSD is currently reachable.
+One should reach the diagrams and tables of the 
+Supporting Information document of this article
+for further directions
+on the available structures and for an overview complement.
 
+% Supporting Information
+% diverse provenance, large database, should be expanded, waits feedback
 \section*{References}
 %
-\bibliographystyle{plainurl}
+% \bibliographystyle{plainurl}
 \bibliography{paper}

### 2016-07-29 supportingInformation.tex
finished first version of SI

@@
 This is a Supporting Information document which exposes ontological
-diagrams and auxiliary tables for the Linked Open Social Data (LOSD)
-database. The main document of the article is in~\cite{losd}.
+diagrams and auxiliary tables for the Linked Open Social Database.
+The main document of the article is in~\cite{losd}.
 \end{abstract}
@@
 In this document we provide diagrams
-for the provenances in the LOSD:
+for the social protocols in the Linked Open Social Database:
 Facebook, Twitter, IRC, Email, ParticipaBR, Cidade Democrática and AA.
-Each provenance diagram was broken in two, one presents the relations
-among main classes (blue nodes) and data types (orange nodes), the other presents metadata on the
-snapshot.
+Each social protocol diagram was broken in two, one presents the relations
+among main classes (blue nodes) and data types (orange nodes),
+the other presents metadata for the
+snapshots.
 Every class instance is related to the snapshot instance
@@
 automatized and displays some overlaps.
-Even so, the images are useful for grasping what is in current LOSD
-and for conducting explorations.
+Even so, the images are useful for grasping what is in the current
+database and for conducting explorations.
 Edges in the diagrams have:
@@
 % \pdfpageheight 300pt
-Furthermore, this document ends with two sets of tables, one with counts of
-triples, participants, edges/interactions/relations and characters,
-the other with references for snapshot groups, such as wikipedia or
-contact links.
+Furthermore, this document ends with two sets of tables, one with
+references for snapshot groups, such as wikipedia or
+contact links, 
+the other with 
+counts of
+triples, participants, edges/interactions/relations and characters.
 
@@
 Each Twitter snapshot is yield by a hashtag.
-Retweets (\textttt{po:retweetOf} are usually considered to yield the interactions between users.
-Users are identified through authors \textttt{po:numericID} (global as given by Twitter API)
-or \textttt{}.
+Retweets (\textttt{po:retweetOf} are usually considered the interactions between users.
 The database present also \textttt{po:replyTo} and \textttt{po:userMention}
-which might also be useful in understanding the networking.
+which might be useful in understanding the networking.
 Further information is found on the following diagrams, the tables on
@@
 Each IRC snapshot is yield by an IRC channel.
-IRC messages are either server messages (e.g. join and exit channel)
+IRC messages are either server messages (e.g. join and exit)
 marked with \textttt{po:systemMessage true} and having an \textttt{po:impliedUser user\_uri},
@@
 \section{Email data}
-Each IRC snapshot is yield by an Email list.
-Interactions is yield through \textttt{po:replyTo} relations
-although \textttt{po:to} and \textttt{po:cc} can also be considered.
-the email body is given by \textttt{po:text} relations while
-\textttt{po:cleanText} has the text with lines removed where they are
+Each email snapshot is yield by an Email list.
+Interactions are yield through \textttt{po:replyTo} relations
+although \textttt{po:to} and \textttt{po:cc} might also be considered.
+The email body is given by \textttt{po:text} relations while
+\textttt{po:cleanText} links to text with lines removed where they are
 trivially from previous messages or computer code.
@@
 Due to the higher complexity of the diagram, an additional figure is
-given rendered with another layout algorithm
+given rendered with another layout algorithm.
 Further information is found on the following diagrams, the tables on
@@
 This snapshot holds a complex structure of both
-Topics/Inspirations/
-Observatories/Supports/Competitions/Prizes
-and of State/City/Neighborhood/Place.
+Topics/Inspirati-ons/Observatories/Supports/Competitions/Prizes
+and of State/City/Neighbor-hood/Place.
 The language used is mainly Brazilian Portuguese.
 Due to the higher complexity of the diagram, an additional figure is
-given rendered with another layout algorithm
+given rendered with another layout algorithm.
 Further information is found on the following diagrams, the tables on
@@
 reference to the subject.}\label{tab:provenance}
-\begin{tabular}{| l || p{4cm} | p{4cm} | }\hline
+\begin{tabular}{| p{4cm} || p{4cm} | p{4cm} | }\hline
     \textbf{Gmane ID} & \textbf{observation} & \textbf{reference} \\\hline\hline
     gmane.linux.audio.users & the Linux Audio Users & \url{http://linuxaudio.org} \\\hline
-    gmane.politics.organizations.metareciclagem & a network about technology and social transformation  & \url{https://metareciclagem.github.io} \\\hline
+    gmane.politics.organizations.me-tareciclagem & a network about technology and social transformation  & \url{https://metareciclagem.github.io} \\\hline
     gmane.linux.audio.devel & the Linux Audio Developers & \url{http://lists.linuxaudio.org/listinfo/linux-audio-dev} \\\hline

### 2016-08-01 paper.tex
some queries added and diagrams updated

@@
             friendships in descending order by the number of friendships:\\
-            \textttt{SELECT DISTINCT ?participant (COUNT(?interaction)
+            \textttt{SELECT DISTINCT ?participant (COUNT(?friendship)
             as ?c) WHERE \{\\
-                \h ?interaction a po:Friendship . ?interaction po:member ?participant . \\
+                \h ?friendship a po:Friendship . ?friendship po:member ?participant . \\
             \} ORDER BY DESC(?c) LIMIT 1000}
-	\item Search string ``pineapple'' in LOSD messages:\\
+        \item Retrieve text messages with the word ``pineapple'' (case insensitive):\\
             \textttt{SELECT ?text WHERE \{ \\
                     \h ?activity po:text ?text . FILTER regex(?text, 'pineapple', 'i')\\
+    \}}
+\item List participants and respective full names whose name has the substring ``Amanda'':\\
+    \textttt{SELECT DISTINCT ?participant ?name WHERE \{\\
+    \h ?participant po:observation ?obs . ?obs po:name ?name .\\
+    \h FILTER regex(?name, 'Amanda', 'i') \\
+    \}}
+\item Return all pairs of friends of a participant which are friends themselves:\\
+    \textttt{SELECT DISTINCT ?friend1 ?friend2 WHERE \{\\
+     \h       ?friendship1 po:member <participant\_uri> .  ?friendship1 po:member ?friend1 .\\
+     \h       ?friendship2 po:member <participant\_uri> .  ?friendship2 po:member ?friend2 .\\
+     \h       ?friendship3 po:member ?friend1 .  ?friendship3 po:member ?friend2 .\\
+    \}}
+\item Return all interactions from replies in a snapshot:\\
+    \textttt{SELECT ?from ?to WHERE \{\\
+          \h  ?message1 po:snapshot <snapshot\_uri> .  ?message2 po:replyTo ?message1 .\\
+          \h  ?message1 po:author ?from .  ?message2 po:author ?to .\\
     \}}

### 2016-08-01 supportingInformation.tex
some queries added and diagrams updated

@@
     paper=graphics
-][scale=.4]{ontologies/gmane-legacy-linux.audio.devel1-20000Email00000.ttl/draw.png}
+][scale=.4]{ontologies/email-legacy-linux.audio.devel1-20000Email00000.ttl/draw.png}
 \textheight = 3in
@@
     paper=graphics
-][scale=.5]{ontologies/gmane-legacy-linux.audio.devel1-20000Meta.ttl/draw.png}
+][scale=.5]{ontologies/email-legacy-linux.audio.devel1-20000Meta.ttl/draw.png}
 

### 2016-08-01 paper.tex
add of directions from chu

@@
 \section{Introduction}
-In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
-both research and application realms.
+
+% In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
+% both research and application realms.
+Open machine-readable contents and tools are desired
+in performing knowledge discovery in social data,
+% there is a lack of datasets for benchmarking results,
 However, there is a lack of datasets for benchmarking results,
-specially in the complex networks field, yielding diverse results from 
-poorly related data.
+specially associated the complex networks field,
+yielding diverse results from poorly related sources.
 
@@
 
-This work presents a linked open social database (LOSD) with diverse
+This work presents a framework that provides machine-readable contents with a
+database on social networks.
+This has been made by linkng social data into an open social database with diverse
 provenance, including Facebook, Twitter, IRC, Email and detached
@@
 Such data is proposed as a common repertoire for scientific
-research involving networks and textual content.
+research involving networks and textual content and may be extended.
 

### 2016-08-03 paper.tex
pequena melhora na escrita da introp

@@
 Open machine-readable contents and tools are desired
-in performing knowledge discovery in social data,
+in performing knowledge discovery in social data.
 % there is a lack of datasets for benchmarking results,
-However, there is a lack of datasets for benchmarking results,
-specially associated the complex networks field,
+However, there is a lack of open datasets for benchmarking results,
+specially associated to the complex networks field,
 yielding diverse results from poorly related sources.
@@
 from different perspectives.
-However, the used datasets in the literature differ within the scope of each research with
+Even so, the used datasets in the literature differ within the scope of each research with
 scarce and historical exceptions such as the
 karate club dataset~\cite{newmanBook}.
-On the other hand, the available linked data is not 
+Beyond that, the available linked data is not 
 stable or rigorous enough to be
@@
 database on social networks.
-This has been made by linkng social data into an open social database with diverse
+This has been made by linkng social data into an open database with diverse
 provenance, including Facebook, Twitter, IRC, Email and detached

### 2017-01-10 tables/facebookReferences.tex
adicionada tabela de referencias do fb

\begin{table*}[h!]
\begin{center}\scriptsize
\caption{All the Facebook snapshots are either the result of individuals who downloaded
their data (and donated to the first author) or data downloaded from groups.
In the first case, it is senseless to present references. In the second
case, we present the group name and a link to a post in the group where
data and figures were delivered back to the group.}
\begin{tabular}{| l | p{9.9cm} |}\hline
\textbf{group name} & \textbf{url(s)} \\\hline
Adorno Nao Eh Enfeite & \url{https://www.facebook.com/groups/265217103529531/permalink/525654127485826/} \\\hline
Ativistas Da Inclusao Digital & \url{https://www.facebook.com/groups/423602557691243/permalink/525201037531394/} \\\hline
Ciencias Com Fronteiras & \url{https://www.facebook.com/groups/contraaexclusao/permalink/269103356558439/} \\\hline
Computer Art & \url{https://www.facebook.com/groups/computerart/permalink/259389137529870/} \\\hline
Coolmeia &  \url{https://www.facebook.com/groups/coolmeia/permalink/380091142098291/} , \url{https://www.facebook.com/groups/coolmeia/permalink/489757754464962/} \\\hline
Democracia Direta Ja & \url{https://www.facebook.com/groups/ddjbrasil/permalink/347023325397298/} \\\hline
Democracia Pura & \url{https://www.facebook.com/groups/democraciapura/permalink/310907215704321/} \\\hline
Economia & \url{https://www.facebook.com/groups/economa1/permalink/586007714743535/} \\\hline
Economia Criativa Digital & \url{https://www.facebook.com/groups/economiacriativadigital/permalink/438313682916103/} \\\hline
Educacoes E Aprendizagens XXI & \url{https://www.facebook.com/groups/geaxxi/permalink/433229973421182/} \\\hline
Latesfip & \url{https://www.facebook.com/groups/183557128478424/permalink/266610616839741/} \\\hline
Living Bridges Planet & \url{https://www.facebook.com/groups/livingbridgesplanet/permalink/352950408144951/} \\\hline
Mobilizacoes Culturais Interior SP & \url{https://www.facebook.com/groups/131639147005593/permalink/144204529082388/} \\\hline
Partido Pirata & \url{https://www.facebook.com/groups/partidopiratabrasil/permalink/10151409024509317/} \\\hline
Politicas Culturas Brasileiras & \url{https://www.facebook.com/groups/pcult/permalink/519626544747423/} \\\hline
Praca Popular & \url{https://www.facebook.com/groups/215924991863921/permalink/319279541528465/} \\\hline
Rede Tranzmidias & \url{https://www.facebook.com/groups/318333384951196/permalink/346658712118663/} \\\hline
Silicon Valley Global Network & \url{https://www.facebook.com/groups/109971182359978/permalink/589326757757749/} \\\hline
Solidarity Economy & \url{https://www.facebook.com/groups/9149038282/permalink/10151461945623283/} \\\hline
Study Group SNA & \url{https://www.facebook.com/groups/140630009439814/permalink/151470598355755/} \\\hline
Tecnoxamanismo &  \url{https://www.facebook.com/groups/505090906188661/permalink/733144993383250/} , \url{https://www.facebook.com/groups/505090906188661/permalink/733157380048678/} \\\hline
\end{tabular}\end{center}
\end{table*}

### 2019-06-20 misc/overallText.tex
finished a first version of the main document
adiantado um pouco do artigo

--- 2016-07-18 misc/overallText.tex
@@
-Database consists of 37467423 triples, 3088219 edges yield by interactions or relations, 382735 participants and 229008584 characters.
-Among all snapshots, 63 are ego snapshots, 53 are group snapshots; 45 have interaction edges, 89 have friendship edges; 42 have text content from messages.
+The database consists of 34,120,026 triples, 3,172,927 network links yield by selected interactions or relations, 382,568 participants and 253,155,020 characters in text data.
+Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 49 have network interaction links, 89 have network friendship links; 43 have text content from messages.

### 2019-06-20 paper.tex
adicionada tabela de referencias do fb
adiantado um pouco do artigo

--- 2016-08-01 paper.tex
@@
 \usepackage{hyperref}
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
 \hypersetup{
@@
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+% \usepackage[T1]{fontenc}
 % \usepackage{changepage}
@@
 \usepackage{tabularx}
+\usepackage{listings}
+\lstdefinelanguage[]{spq}
+{
+    morecomment=[s][\color{ForestGreen}\bf]{<}{>},
+    basicstyle = \ttfamily\footnotesize,
+    commentstyle=\color{green},
+    keywordstyle=\color{Plum},
+    morecomment=[s][\color{Blue}\bf]{?}{\ },
+    morekeywords={BASE,PREFIX,SELECT,DISTINCT,CONSTRUCT,DESCRIBE,ASK,%
+    FROM,NAMED,WHERE,ORDER,BY,ASC,DESC,LIMIT,OFFSET,OPTIONAL,%
+    GRAPH,UNION,FILTER,a,STR,LANG,LANGMATCHES,DATATYPE,BOUND,%
+    isIRI,isURI,isBLANK,isLITERAL,REGEX,true,false},%
+    sensitive=false,%
+    % morecomment=[l]\#,%
+    morestring=[d][\color{red}]',%
+    morestring=[d]"%
+}
 % \usepackage[showframe=true]{geometry}
 %
+
 \newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
+\newcommand{\te}[1] {\texttt{\footnotesize#1}}
 \newcommand{\h} {\hphantom ~ }
@@
 % }
+\linespread{1}
 \pgfplotsset{compat=1.5}
@@
 }
-
+\usepackage{booktabs}
 %% `Elsevier LaTeX' style
@@
 %
-\title{A Linked Open Social Database for Scientific Benchmarking}
+\title{A Linked Open Social Dataset for Scientific Benchmarking}
 %
-\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
-\ead{fabbri@usp.br}
+\author[pws]{Renato Fabbri\corref{corresponding}}
+\ead{renato.fabbri@gmail.com}
 %
-\author[pwr]{Osvaldo Novais de Oliveira Junior\fnref{kio-url}}
+\author[pwr]{Osvaldo Novais de Oliveira Junior}
 \ead{chu@ifsc.usp.br}
@@
 \cortext[corresponding]{Corresponding author}
-\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
-University, Brazil}
-%
-\fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
+  \address[pws]{Institute of Mathematics and Computer Sciences, University of São Paulo (ICMC/USP), Brazil}
+\address[pwr]{S\~ao Carlos Institute of Physics, University of  São Paulo
+(IFSC-USP), Brazil}
 %
@@
 are widely researched.
-Recently, a myriad of results have been reported which are based in
-diverse datasets most often not accessible to researchers other than the publishing authors.
-This work exposes an open database with diverse provenance and oriented
+In fact, a myriad of results have been reported which are based in
+diverse data, most often not accessible to researchers other than the publishing authors.
+This work presents an open dataset with diverse provenance of 
+human social online networking 
+  and oriented
 to furnish the scientific community with a friendly and common repertoire.
-Current data was obtained from Facebook, Twitter, IRC, Email and the
-detached instances of ParticipaBR, AA and Cidade Democr\'atica.
-These were represented as linked data to homonenize access,
-conform to current best practices and ease analyzes which integrate third
+Current data was obtained in 2013-16 from well-known online social networking platforms:
+Facebook, Twitter, IRC, Email.
+Data from ParticipaBR, AA, and Cidade Democrática were also added as samples of
+less known instances.
+Authors recently recognized that no similar dataset was made available,
+and the potential benefit of the content for the research community.
+The linked data representation was adopted in order to
+comply with current best practices, homogenize access, enable discovery,
+and facilitate analyzes which integrate third
 party and provided instances.
+Futhermore, the method for obtaing and making the data available is potentially novel, and may of use for other research groups.
 This document presents an outline and overall statistics of the given
-database which should favor subsequent work.
+dataset, which should favor subsequent work in complex and social networks,
+  and in enhancing the available linked and social data.
 \end{abstract}
@@
 \begin{keyword}
-Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email,
-Complex Networks, Text Mining
+Big Data, Dataset, Benchmark Data, Facebook, Twitter, IRC, Email,
+Complex Networks, Social Networks, Text Mining
 %Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
@@
 \section{Introduction}
-In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
-both research and application realms.
-However, there is a lack of datasets for benchmarking results,
-specially in the complex networks field, yielding diverse results from 
-poorly related data.
-
-The enormity of the digital data propels a rapid development of analysis methods
-from different perspectives.
-However, the used datasets in the literature differ within the scope of each research with
-scarce and historical exceptions such as the
-karate club dataset~\cite{newmanBook}.
-On the other hand, the available linked data is not 
-stable or rigorous enough to be
-a public reference on statistical physics and social networks research.
-
-This work presents a linked open social database (LOSD) with diverse
-provenance, including Facebook, Twitter, IRC, Email and detached
-instances.
-Such data is proposed as a common repertoire for scientific
+% In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
+% both research and application realms.
+The research on human social and complex networks often realy on data that express
+real phenomena~\cite{c1,c2,c3}.
+Such data is often not made publicly available, which hinders validation of the results and does not favor makes subsequent work.
+Furthermore, there is a lack of open datasets for benchmarking results,
+specially associated to the complex and social networks field,
+yielding diverse results from poorly related sources.
+Scarce and historical exceptions are the karate club or the dolphin datasets~\cite{newmanBook,dolphin}.
+For expressing the information,
+open machine-readable content and tools are desired for enabling
+data linkage and discovery~\cite{o1,o2},
+and the favoured protocol is thus RDF ``linked data''.
+Social networks linked data is currently not stable or rigorous enough to be
+a public reference on statistical physics and social networks research,
+with scarce exceptions~\cite{l1,l2}
+This work presents a collection of linked and open data
+about online human social networks,
+named LOSD (Linked Open Social Data).
+This has been achieved by linking social data into a dataset using diverse
+provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
+Such data is proposed as a common and public repertoire for scientific
 research involving networks and textual content.
 
+\subsection{Related work}
+Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
+on specialized communities such as related to research and movies~\cite{s1,s2,s3,s4,s5},
+or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
+Contributions which are most similar to LOSD are mentioned in this section.
+In~\cite{fb1}, data obtained only from Facebook is reported.
+With such a focus, the authors make available a number of statistics,
+but the data is offline at least since 2010.
+Linked human social network data is scarcely found in the scientific community.
+In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud, and made available a dataset from the Advogato platform,
+dedicated to free software developers linked by trust relatioships.
+There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
+most often with a specific focus (users, tweets, hashtags, etc)
+and not presented to the scientific community by means of a research article.
+In summary, no human social network dataset presented to the
+scientific community containts the diverse provenance used in LOSD,
+potentially because researchers did not have the means to acquire and
+share the data described in Section~\ref{acq}.
+
 % \subsection{Benchmarking in the analysis of complexity}
+
+\section{Methods}
+LOSD is expressed as linked data through RDF
+and is ontologicaly described through the
+data-driven ontology synthesis method presented in Section~\ref{ont}.
+
+\subsection{Data acquisition}\label{acq}
+Research involving human subjects raise well-known ethic issues~\cite{}.
+Furthermore, the online social networking platforms often restrict access
+to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
+In this context, research using data, e.g. from Facebook, is considered
+problematic if not prohibited.
+We reached a potential solution to this issue through continuated research with
+collaborators in the fields of phylosophy, social sciences, and psychology.
+The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
+consists in researching first the social networks related to the researchers,
+and keeping the procedures, data, algorithms and results publicly available.
+This practices enabled collecting and sharing the data in LOSD.
+Notice that, although Twitter data is available through public APIs, the other sources are more problematic.
+
+% download via netviz when fb allowed it, donations
+% twitter, bots, gmane
+% contact with developers of AA, PBR, CD
+% Anthropological physics
+
+\subsection{Linked open data}
+Linked data refers to data published in the web in such a way that it is
+machine readable and conforms to a set of best practices.
+The yielded web of data is composed of documents on the web
+such as the web of HTML documents.
+In practice, the publication of linked data can me summarized
+by 1) the use of RDF to publish data on the web and 2) the use of RDF
+links to interlink data from different sources~\cite{ld1,ld2}.
+The web is expected to be interconnected and to grow by the systematic application of four
+steps~\cite{lee1}:
+\begin{itemize}
+    \item Use URIs to identify things~\cite{uri}.
+    \item Use HTTP URIs.
+    \item Provide useful information when a URI is accessed via HTTP.
+    \item Provide other URIs in the description of resources so human
+        and machine agents can perform discovery.
+\end{itemize}
+
+The Linked Open Data~\cite{lod} is a constantly growing cloud of data,
+the global data space, which is usually
+conceived as centered around the DBPedia, a linked data representation
+of data from Wikipedia~\cite{dbpedia0,dbpedia},
+and sometimes called the Giant Global Graph (GGG, akin to the WWW).
+
+\subsection{RDF}
+The Resource Description Framework (RDF), a W3C
+recommendation, is a model for data
+interchange~\cite{w3cr}.
+It is based on the idea of making statements about resources in the form
+of triples, i.e. expressions in the form ``subject - predicate -
+object''.
+RDF can be serialized in several file formats, including RDF/XML,
+Turtle and Manchester, which all, in essence, represent a labeled and
+directed multi-graph.
+RDF may be stored in a type of database called a triplestore~\cite{rdf}.
+
+As an example of an RDF statement, the following triple in the Turtle
+format asserts that ``the paper has color white'':\\
+\texttt{http://example.org/Thing\#Paper http://example.org/hasColor\\
+http://example.org/Color\#White .}
+
+\subsection{RDFS and OWL ontologies}\label{sont}
+One of the most important notions of the semantic web is that of an ontology~\cite{ont}.
+An ontology, in this context, is a formalized conceptualization, comprised by
+concepts and relations between the concepts and between the relations themselves.
+In current semantic web, most simple ontologies are written using the RDFS protocol,
+by which one can specify, among other things~\cite{rdfs}:
+\begin{itemize}
+  \item concepts;
+  \item properties, which are concepts that are used as predicates and thus relate concepts;
+  \item special relations between concepts that state that one concept is more general than the other\footnote{This
+          kind of relation is called hypernymy. Examples: mammal is a hypernym of monkey, drink is a hypernym of beer.};
+  \item the subjects and objects that can occurs in a triple where a specific property is a predicate.
+\end{itemize}
+
+One can also write an ontology using the OWL protocol,
+with which all the expressive capabilities of RDFS are
+available, but one can also, among other things~\cite{owl}:
+\begin{itemize}
+  \item state ``property axioms'', i.e. specify if a property is e.g. reflexive or transitive;
+  \item state ``class restrictions'', i.e. specify if a class instance e.g. necessarily holds a relation to another class instance or data.
+\end{itemize}
+OWL has a richer vocabulary than RDFS and is (way more) complex.
+This complexity is a drawback together with the greater computational cost
+for performing inference.
+The advantage is the greater power to represent conceptualizations.
+Using ontologies enables automated reasoning.
+For example, if a property \textttt{:r} is known to relate only mokeys to trees,
+then, if there is a triple \textttt{:a :r :b}, the instance \textttt{:a} is considered a monkey,
+and instance \textttt{:b} is considered a tree,
+even if there is no explicit declaration similar
+to \textttt{:a rdf:type :Mokey}.
+
+For the data described in this article, the following features are most relevant~\cite{rdfs,owl}:
+\begin{itemize}
+  \item Using RDFS, one may define a property domain, i.e. the class instances that are allowed as the object in a triple with the property.
+  \item Using RDFS, one may define a property range, i.e. the class instances that are allowed as the predicate in a triple with the property.
+  \item Using OWL, one may define an existential restriction for a class \textttt{:C},
+    i.e. declare that an instance \textttt{:C\#I} of such class has at least one relation with a property \textttt{:p} and a class \textttt{:C2} in the form
+    \textttt{:C\#I :p :C2\#I}, where \textttt{:C2\#I} is an instance of the class \textttt{:C2}.
+  \item Using OWL, one may define an universal restriction for a class \textttt{:C},
+    i.e. declare that any relation of an instance \textttt{:C\#I} of such class, with a property \textttt{:p} of the form \textttt{:C\#I :p :CX\#I},
+    implies that \textttt{:CX\#I} is an instance of a specific class, e.g. \textttt{:C2}.
+\end{itemize}
+
+\subsection{Data-driven ontology synthesis}\label{ont}
+OWL Ontologies are tools to describe taxonomies and the
+structure of knowledge~\cite{owl1}.
+Most ontologies are created by domain experts even though the data they
+arrange is often given by a software system and has a predefined
+structure~\cite{owl2}.
+
+We developed a simple ontology synthesis method that probes
+the ontological structure in the data with
+SPARQL queries and post-processing.
+The results are OWL code and diagrams which are available in the
+Supporting Information document of this article.
+The method can be extended to comprise further OWL axioms and restrictions,
+but is currently performed to fit present needs with maximum simplicity.
+Present needs are limited to informative figures and
+the steps implemented are as follows:
+\begin{enumerate}[leftmargin=0cm]
+    \item Obtain all distinct classes with the query:\\
+        \h\textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri .\}}
+\item For each class, obtain the properties that occur as predicates in triples where the subject is an instance of the class:\\
+	\h\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o .\}}\\
+Such properties are used to assert existential and universal restrictions for the class.
+\item Compare the total number of individuals (\textttt{?cs1}) of the class (\textttt{class\_uri}) with
+	the number of such individuals (\textttt{?cs2}) that are subjects of at least one triple where 
+        the predicate is the property (\textttt{property\_uri}) and the object is
+        an instance of the same class (\textttt{class2\_uri}).
+	The queries are:\\
+	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
+        \h\textttt{SELECT DISTINCT ?oc WHERE \{\\
+	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a ?oc . \\ 
+        \h\}}\\
+        Then, for each object class (\textttt{object\_curi}) in \textttt{?oc}:\\
+	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs2) WHERE \{\\
+	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a <object\_curi> .\\         \h\}}\\
+        When \textttt{?cs1} matches \textttt{?cs2}, an existential restriction is found. And with the query:
+	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
+	\h\h ?s a <class\_uri> . ?s <property\_uri> ?o . ?o a ?oc .\\
+        \h\h FILTER(str(?oc) != 'object\_curi')\\
+        \h\}}\\
+       When \textttt{?cs} is 0, a universal restriction is found. 
+\item To keep a record of further restrictions, get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
+	\h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
+        \h\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
+        \h\}}
+    \item Obtain all distinct properties:\\
+        \h\textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
+    \item Check if each property is functional, i.e. if it
+        occurs at most once with each subject.
+        This is performed by counting the objects and further verifying
+        that they are at most one. The query is:\\
+        \h\textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
+            <property\_uri> ?o \} GROUP BY ?s}
+          \item For each property, find the incident range (possible subjects) and domain (possible objects) with the
+        queries:\\
+        \h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
+        \h\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
+        \h\}} \\
+        and \\
+        \h\textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
+%     \item For each instance of each class, get all distinct predicates.
+%         For each predicate, check if all instances of the class
+%         hold such relationship (existential restriction):\\
+%         \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
+%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
+%         \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
+%                 a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
+%     \item and if all instances that hold such relationship are instances of the class
+%         (universal restriction):\\
+%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
+%     \item Draw each class, each property and the overall figure.
+      \item Render diagrams to achieve figures as exemplified in Section~\ref{sdia} and in the Supporting Information document.
+%      \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
+%          statements to better organize knowledge and link to third party
+%          ontologies and data. This step is left aside in current LOSD as 
+\end{enumerate}
 
@@
 Integration and uniformity of access is obtained through linked data
-representation, as exposed in Section~\ref{queries}.
-This section introduce the underlying data in very concise terms.
+representation, as described in Section~\ref{queries}.
+This section introduces the underlying data in very concise terms.
 One should access the Supporting Information document for
-a thorough presentation of the database.
+a more information about the dataset.
 
@@
 \subsection{Facebook data}
-Friendship ego networks (networks whose constituents are friends of an user)
+Friendship ego networks (networks whose nodes represent friends of a user)
 were donated from individual users in 2013 and 2014.
@@
 Each snapshot is unified by a distinct hashtag.
-Edges are canonically yield by retweets but replies and user mentions
-are also kept in the database.
+Network links are canonically yield by retweets,
+but replies and user mentions are also kept in the database.
 
 \subsection{IRC data}
-Public IRC logs were used to render IRC snapshots.
-The database has records of users to which the message is directed to or
+Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
+The database has records of relations yield by directed messages and
 mentions.
@@
 Email snapshots refer to individual email lists.
-All messages were obtained from the Gmane public email database~\cite{gmane}.
+All messages were obtained from the Gmane public email archive~\cite{gmane}.
 Each message has the original text and the text without some of the lines
-from previous messages or that are software code.
+from previous messages or that are software code or author signature.
 Most importantly, each message instance holds the ID of the message it is
-a reply to, if any.
+a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}
 
 \subsection{ParticipaBR data}
-The ParticipaBR is a Brazilian federal platform for social participation.
-Texts are derived from blog posts and networks are derived from
+ParticipaBR was a Brazilian federal platform for social participation,
+the most important platform from the myriad of social participation platforms
+that were available in 2011-2018~\cite{spbr}.
+Texts are derived from blog posts and networks may be obtained from
 friendship and interaction criteria.
@@
 \subsection{Cidade Democr\'atica data}
-Cidade Democr\'atica is a Brazilian civil society social participation portal.
+Cidade Democrática is a Brazilian civil society social participation portal.
 Data gathered is complex in the number of types of instances.
@@
 \subsection{AA data}
-The Algorithmic Autoregulation~\cite{aa} is a software development
-methodology based on testifying and sharing ongoing work.
+AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
+it is based on testifying and sharing ongoing work.
 The data was gathered from different versions of the system and from an IRC
@@
 
-\section{Methods}
-Data is represented as linked open data through
-RDF and ontologicaly described through a data-driven ontology synthesis
-method.
-
-\subsection{Linked open data}
-Linked data refers to data published in the web in such a way that it is
-machine readable and conforms to a set of best practices.
-The web of data is constructed with documents on the web 
-such as the web of HTML documents.
-In practice, the idea of linked data can me summarized
-by 1) the use of RDF to publish data on the web and 2) the use of RDF
-links to interlink data from different sources.
-The web is expected to be interconnected and to grow by the systematic application of four
-steps~\cite{lee1}:
-\begin{itemize}
-    \item Use URIs to identify things~\cite{uri}.
-    \item Use HTTP URIs.
-    \item Provide useful information when an URI is accessed via HTTP.
-    \item Provide other URIs in the description of resources so human
-        and machine agents can perform discovery.
-\end{itemize}
-
-The Linked Open Data~\cite{lod} builds an ever growing cloud of data,
-the global data space, which is usually
-conceived as centered around the DBPedia, a linked data representation
-of data from Wikipedia~\cite{dbpedia0,dbpedia}.
-
-\subsection{RDF}
-The Resource Description Framework (RDF), a W3C
-recommendation, is a model for data
-interchange.
-It is based on the idea of making statements about resources in the form
-of triples, i.e. expressions in the form ``subject - predicate -
-object''.
-RDF can be serialized in several file formats, including RDF/XML,
-Turtle and Manchester, all which, in essence, represent a labeled and
-directed multi-graph.
-RDF may be stored in a type of database called a triplestore~\cite{rdf}.
-
-As an example of an RDF statement, the following triple in the Turtle
-format asserts that ``the paper has color white'':\\
-\texttt{http://example.org/Things\#Paper http://example.org/hasColor\\
-http://example.org/Colors\#White .}
-
-\subsection{Data-driven ontology synthesis}
-OWL Ontologies are critical tools to describe taxonomies and the
-structure of knowledge.
-Most ontologies are created by domain experts even though the data they
-arrange is often given by a software system and has a predefined
-structure.
-
-We developed a simple ontology synthesis method that probes
-the ontological structure in data with
-SPARQL queries and post-processing.
-The results are OWL code and diagrams which are available in the
-Supporting Information document.
-The method can be extended to comprise further OWL axioms and restrictions,
-but is currently performed to fit present needs with maximum simplicity.
-Present needs are limited to informative figures and
-the steps implemented are as follows:
-\begin{enumerate}[leftmargin=0cm]
-    \item Obtain all distinct classes with the query:\\
-        % \textttt{SELECT DISTINCT ?class WHERE { ?s a ?class }}
-        \textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri \}}
-%    \item For each class, get distinct subject classes and predicates where the
-%        the object is an instance of the class:\\
-%        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
-%        a ?cs . \}}
-% \item For each class, obtain distinct predicates (\textttt{?p}) and object classes or
-% 	datatypes (\textttt{?co}, \textttt{?do}) where the subject is an instance (\textttt{?i}) of such class:\\
-% 	\textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ \\
-% 		\h ?i a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \\
-% \}}
-\item For each class, obtain the properties that occur as predicates in triples where the subject is an instance of the class:\\
-	\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o . \}}\\
-Such properties are used to assert existential and universal restrictions for the class.
-\item Compare the total number of individuals (\textttt{?cs1}) of the class (\textttt{class\_uri}) with
-	the number of such individuals (\textttt{?cs2}) that are subjects of at least one triple where 
-	the predicate is the property (\textttt{property\_uri}).
-	If the numbers match, there is an existential restriction for the class. The queries are:\\
-	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
-	\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
-	\h ?s a <class\_uri>. ?s <property\_uri> ?o .\\ \}}
-\item Find the number of instances which are subjects of triples where the predicate is the property but are not instances of the class.
-	If there is zero of such instances, there is an universal restriction:\\
-	\textttt{SELECT (COUNT(DISTINCT ?s)=0 as ?cs) WHERE \{\\
-	\h ?s <property\_uri> ?o . ?s a ?ca . FILTER(str(?ca) != 'class\_uri')\\ \}}
-\item To keep a record of the restrictions (and occurring triples), get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
-	\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
-		\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
-\}}
-    \item Obtain all distinct properties:\\
-        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
-    \item Check if each property is functional, i.e. if it
-        occurs at most once with each subject.
-        This is performed by counting the objects and further verifying
-        that they are at most one. The query is:\\
-        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
-            <property\_uri> ?o \} GROUP BY ?s}
-    \item For each property, find the incident range and domain with the
-        queries:\\
-        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
-			\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\\}} \\
-        and \\
-        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
-%     \item For each instance of each class, get all distinct predicates.
-%         For each predicate, check if all instances of the class
-%         hold such relationship (existential restriction):\\
-%         \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
-%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
-%         \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
-%                 a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
-%     \item and if all instances that hold such relationship are instances of the class
-%         (universal restriction):\\
-%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
-%     \item Draw each class, each property and the overall figure.
-\item Render diagrams as exposed in the next section and in the Supporting Information file.
-%      \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
-%          statements to better organize knowledge and link to third party
-%          ontologies and data. This step is left aside in current LOSD as 
-\end{enumerate}
-
 \section{Results}
 \label{outline}
-Current overall results concern data selection and preparation for knowledge discovery.
+Current overall results concern data selection and preparation for data discovery.
 The main result is the data made available, which enables benchmarking of scientific results
-and easy experimentations.
+and easy experimentations, and no compliant dataset was found by the authors.
 Secondary results include data outline through figures and tables,
@@
 \subsection{Standardization}
-The data is embedded into standard URIs and triples, i.e. translated to RDF.
+The data is embedded into URIs and triples, i.e. translated to RDF.
 URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
-which are identified herein with the prefix \textttt{po:}.
+which are identified herein with the prefix ``\textttt{po:}''.
 Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
@@
 
-Instances are built on top of the class they derive from plus a hashtag character,
+Instance URIs are built using the URI from the class they derive from plus a hashtag character,
 a provenance string (e.g. \textttt{facebook-legacy} or
-\textttt{participabr-legacy}) of the snapshot they refer to, and an identifier;
+\textttt{participabr-legacy}), and an identifier;
 i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
@@
 \subsection{Software tools}
-The database is released with software for rendering itself, analyses and
-multimedia artifacts.
+The authors made available software for rendering analyses and
+the data set itself.
 \subsubsection{Triplification routines}
 For each social platform there is a \emph{triplification} routine,
-i.e. a script for translating data to RDF.
+i.e. a script for expressinf the data as RDF.
 Original formats and further observations are presented in
 Table~\ref{tab:provenance}.
+  {\renewcommand{\arraystretch}{1.2}
 \begin{table*}[h!]\scriptsize
@@
 the database.}\label{tab:provenance}
-\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
-    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
+\begin{tabular}{ l | p{3cm} p{3cm} c }
+    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
     AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
     Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
-    Email & mbox & obtained through Gmane public database & Gmane~\cite{gmane} \\\hline
+    Email & mbox & obtained through the Gmane public archive & Gmane~\cite{gmane} \\\hline
     Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social~\cite{social} \\\hline
@@
     ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
-    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
+    Twitter & JSON & obtained through Twitter streaming API & Social \\
 \end{tabular}\end{center}
 \end{table*}                    
+  }
 \subsubsection{Topological and textual analysis}\label{ana}
 Routines are available for taking topological and textual measures from
-the database.
-Auxiliary routines, such as performing principal component analysis
-and taking Kolmogorov-Smirnov measures, are available
-to ease pattern recognition.
-Single, timeline and multi-scale analyzes are automated.
+the LOSD data.
+  Auxiliary routines, such as for performing principal component analysis (PCA)
+  and obtaining statistics derived from the Kolmogorov-Smirnov test~\cite{kol}, are available to ease pattern recognition.
 
-\subsubsection{Multimedia rendering}\label{media}
-It is a core purpose of framework to provide routines for rendering
-audiovisualizations of the data.
-Social structures are rendered into music, images and video animations
-through the Percolation toolbox~\cite{percolation} in association with
-the Music and Visuals toolboxes~\cite{music,visuals}.
+% \subsubsection{Multimedia rendering}\label{media}
+% It is a core purpose of framework to provide routines for rendering
+% audiovisualizations of the data.
+% Social structures are rendered into music, images and video animations
+% through the Percolation toolbox~\cite{percolation} in association with
+% the Music and Visuals toolboxes~\cite{music,visuals}.
 
-\subsubsection{Migration from deprecated toolboxes}
-Routines mentioned in Sections~\ref{ana} and~\ref{media} are being migrated from deprecated
-toolboxes~\cite{gmaneLegacy,percolationLegacy} into newly designed
-toolboxes~\cite{percolation,visuals}.
+% \subsubsection{Migration from deprecated toolboxes}
+% Routines mentioned in Sections~\ref{ana} and~\ref{media} are being migrated from deprecated
+% toolboxes~\cite{gmaneLegacy,percolationLegacy} into newly designed
+% toolboxes~\cite{percolation,visuals}. TTM, recover?
 
-\subsection{Diagrams of the data and auxiliary tables}
-The database exploration can be assisted through diagrams which expose
+\subsection{Diagrams of the data and auxiliary tables}\label{sdia}
+LOSD navigation may be assisted through diagrams which describes
 the structure from each provenance.
 Such diagrams are in the Supporting Information document
-with some tables to ease understanding of the provided data.
-A simplified example is given in Figure~\ref{dia} where the friendship
-structure of the Facebook snapshots are exposed.
+together with tables to facilitate the appreciation of the provided data.
+A very simple example is given in Figure~\ref{dia} where the friendship
+structure of the Facebook snapshots is revealed.
 
@@
     \includegraphics[width=0.5\textwidth]{ontologies/facebook-legacy-AntonioAnzoategui18022013Friendship.ttl/draw}
-    \caption{A diagram of the structure involved in the friendship networks
-    of the Facebook snapshot.
-A green edge denotes an OWL existential class restriction; an inverted nip denotes an OWL universal class restriction; a full (non-dashed) edge denotes an OWL functional property axiom.
-Further information and complete diagrams for each provenance are in the Supporting Information document.}\label{dia}
+    \caption{A diagram of the core structure that entails the friendship networks
+    of the Facebook snapshots.
+A green edge denotes an OWL existential class restriction; an inverted nip denotes an OWL universal class restriction; a full (non-dashed) edge denotes an OWL functional property axiom (see Section~\ref{sont}).
+Further information and complete diagrams for each provenance are available in the Supporting Information document.}\label{dia}
 \end{figure}
@@
 \subsection{SPARQL queries}\label{queries}
-There are numerous useful and general purpose SPARQL queries to be performed against the database.
-Here we write some of the most basic of such queries selected by their potential to be varied.
-All queries assume the use the preamble \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
+There are numerous useful and general purpose SPARQL queries to be performed against LOSD.
+In this section, a selection of some of the most basic queries are made available due to their potential to be varied.
+All queries assume the preamble \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
 \begin{enumerate}[leftmargin=0cm]
 	\item Retrieve the number of participants:\\
-            \textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
+            \h\textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
             ?author a po:Participant . \} }
@@
             friendships:\\
-            \textttt{SELECT (COUNT(?interaction) as ?c) WHERE \{\\
-                    \h \{ ?interaction a po:Friendship \} UNION \{ ?interaction
-                    a po:Interaction \} UNION\\
-                    \h \{ ?interaction po:retweetOf
-                    ?message \} UNION \{ ?interaction po:replyTo ?message
+            \h\textttt{SELECT (COUNT(?interaction) as ?c) WHERE \{\\
+            \h        \h \{ ?interaction a po:Friendship \} UNION \{ ?interaction
+                    a po:Interaction \}\\
+            \h        \h  UNION \{ ?interaction po:retweetOf
+                    ?message \} \\
+            \h\h  UNION \{ ?interaction po:replyTo ?message \}\\
+            \h        \h UNION \{ ?interaction po:directedTo ?participant
                     \}\\
-                    \h UNION \{ ?interaction po:directedTo ?participant
-                    \}\\ \} }
-	\item Retrieve all text produced by an specific user:\\
-            \textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{\\
-                    \h ?activity po:author <user\_uri> . ?activity po:text ?text .\\
-    \}}
+            \h\} }
+                  \item Retrieve all text produced by a specific user with URI \textttt{user\_uri}:\\
+            \h\textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{\\
+            \h        \h ?activity po:author <user\_uri> . ?activity po:text ?text .\\
+            \h\}}
         \item List 1000 users (URIs and names) with the most friendships and the number of
             friendships in descending order by the number of friendships:\\
-            \textttt{SELECT DISTINCT ?participant (COUNT(?friendship)
-            as ?c) WHERE \{\\
-                \h ?friendship a po:Friendship . ?friendship po:member ?participant . \\
-            \} ORDER BY DESC(?c) LIMIT 1000}
-        \item Retrieve text messages with the word ``pineapple'' (case insensitive):\\
-            \textttt{SELECT ?text WHERE \{ \\
-                    \h ?activity po:text ?text . FILTER regex(?text, 'pineapple', 'i')\\
-    \}}
+            \h\textttt{SELECT DISTINCT ?participant (COUNT(?friendship) as ?c) WHERE \{\\
+            \h    \h ?friendship a po:Friendship . ?friendship po:member ?participant . \\
+            \h\} ORDER BY DESC(?c) LIMIT 1000}
+        \item Retrieve text messages with the word ``crucial'' (case insensitive):\\
+            \h\textttt{SELECT ?text WHERE \{ \\
+            \h        \h ?activity po:text ?text . FILTER regex(?text, 'crucial', 'i')\\
+            \h\}}
 \item List participants and respective full names whose name has the substring ``Amanda'':\\
-    \textttt{SELECT DISTINCT ?participant ?name WHERE \{\\
-    \h ?participant po:observation ?obs . ?obs po:name ?name .\\
-    \h FILTER regex(?name, 'Amanda', 'i') \\
-    \}}
-\item Return all pairs of friends of a participant which are friends themselves:\\
-    \textttt{SELECT DISTINCT ?friend1 ?friend2 WHERE \{\\
-     \h       ?friendship1 po:member <participant\_uri> .  ?friendship1 po:member ?friend1 .\\
-     \h       ?friendship2 po:member <participant\_uri> .  ?friendship2 po:member ?friend2 .\\
-     \h       ?friendship3 po:member ?friend1 .  ?friendship3 po:member ?friend2 .\\
-    \}}
+    \h\textttt{SELECT DISTINCT ?participant ?name WHERE \{\\
+    \h\h ?participant po:observation ?obs . ?obs po:name ?name .\\
+    \h\h FILTER regex(?name, 'Amanda', 'i') \\
+    \h\}}
+  \item Return all pairs of friends of a participant (with URI \textttt{participant\_uri}) which are friends themselves:\\
+     \h\textttt{SELECT DISTINCT ?friend1 ?friend2 WHERE \{\\
+     \h\h       ?friendship1 po:member <participant\_uri> . ?friendship1 po:member ?friend1 .\\
+     \h\h       ?friendship2 po:member <participant\_uri> . ?friendship2 po:member ?friend2 .\\
+     \h\h       ?friendship3 po:member ?friend1 . ?friendship3 po:member ?friend2 .\\
+     \h\}}
 \item Return all interactions from replies in a snapshot:\\
-    \textttt{SELECT ?from ?to WHERE \{\\
-          \h  ?message1 po:snapshot <snapshot\_uri> .  ?message2 po:replyTo ?message1 .\\
-          \h  ?message1 po:author ?from .  ?message2 po:author ?to .\\
-    \}}
+    \h\textttt{SELECT ?from ?to WHERE \{\\
+    \h      \h  ?message1 po:snapshot <snapshot\_uri> .  ?message2 po:replyTo ?message1 .\\
+    \h      \h  ?message1 po:author ?from .  ?message2 po:author ?to .\\
+    \h\}}
 \end{enumerate}
+
+\subsubsection{Obtaining the networks}\label{snet}
+A complex network $N=(n,l,m)$ may be regarded essentially as
+a set of nodes $n$, a set of links $l$ that related the nodes,
+and metadata $m$ such as node and link weights, text related to the nodes, etc.
+This section presents the queries for obtaining networks using LOSD
+which are promptly envisioned by the authors.
+Many other networks may be obtained because the criteria that entail
+links, and restrains in the set of nodes and of links,
+are determined by the research being developed.
+Furthermore, we do not focus here on multilayer networks (e.g. bipartite networks), or on multigraphs (or muldimensional networks), i.e. on networks with more than one type of node or of link.
+Here is a list with short descriptions of the networks and the corresponding SPARQL queries:
+\begin{itemize}
+  \item Overall facebook friendship network available in LOSD:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?f a po:Friendship .  ?f po:member ?a1 . ?f po:member ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Overall facebook friendship network derived only from ego networks:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?f a po:Friendship . ?f po:snapshot ?s . ?s po:isEgo True .
+  ?f po:member ?a1 . ?f po:member ?a2 .
+
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Overall facebook interaction network available in LOSD:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?f a po:Interaction .
+  ?f po:interactionFrom ?a1 . ?f po:interactionTo ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Email interaction network for email list shapshot with URI \verb|http://xxxx|, with the all (preliminarily cleaned) text related to each message (node atribute):
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 ?t1 ?t2 WHERE { 
+  ?m po:snapshot <http://xxxx> . ?m a po:EmailMessage .
+  ?m po:author ?a1 . ?m po:replyTo ?m2 . ?m2 po:author ?a2 .
+  ?m po:cleanText ?t1 . ?m2 po:cleanText ?t2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item IRC users related by messages directed to each other, and the text of each message (link attribute):
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 ?t WHERE { 
+  ?m a po:IRCMessage .  ?m po:author ?a1 . ?m po:directedTo ?a2 .
+  ?m po:cleanText ?t .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Twitter users related by retweets:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?m1 po:retweetOf ?m2 . ?m1 po:author ?a1 . ?m2 po:author ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Twitter users related by user mentions and the text of the tweets:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 ?t WHERE { 
+  ?m a po:Tweet . ?m po:author ?a1 . ?m po:userMention ?a2 .
+  ?m po:text ?t .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Friendship network from ParticipaBR, with the date when each frienship was established:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 ?d WHERE { 
+  ?s po:shapshotID 'participabr-legacy' . ?f po:snapshot ?s .
+  ?f a po:Friendship . 
+  ?f po:member ?a1 . ?f po:member ?a2 . ?f po:createdAt ?d .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Interaction network from ParticipaBR obtained by comments on articles:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?s po:shapshotID 'participabr-legacy' . ?a po:snapshot ?s .
+  ?a a po:Article .
+  ?a po:author ?a1 . ?c po:article ?a . ?c po:author ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Interaction network from ParticipaBR obtained by comments and votes on comments:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?s po:shapshotID 'participabr-legacy' . ?c po:snapshot ?s .
+  ?c a po:Comment .  ?c po:author ?a1 .
+  { ?v a po:Vote . ?v po:reference ?c . ?v po:author ?a2 .
+  } UNION {
+    ?c2 a po:Comment . ?c2 po:replyTo ?c . ?c2 po:author ?a2 .
+  }
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item Interaction network from Cidade Democrática obtained by comments on Topics:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?s po:shapshotID 'cidadedemocratica-legacy' . ?t po:snapshot ?s .
+  ?t a po:Topic .  ?t po:author ?a1 .
+  ?c a po:Comment . ?c po:topic ?t . ?c po:author ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+  \item AA users, related by AA session validations:
+\begin{lstlisting}[language=spq]
+SELECT ?a1 ?a2 WHERE { 
+  ?s po:author ?a1 . ?s po:checkParticipant ?a2 .
+}
+\end{lstlisting}\vspace{-0.2cm}
+\end{itemize}
+
+Again, this is not an exaustive list.
+For example, other instances which are related in LOSD, and have authors,
+also yield valid criteria for interaciton networks.
+For a less canonical example for obtaining networks, links may be derived from sufficiently similar vocabulary used by participants, or from activity is sufficiently similar dates, which yield networks which are not interaction or friendship networks.
+% mk figure with two participants and their relation though response, retweet, or friendship
 
@@
 % Facebook, Twitter, IRC, Gmane, Participa, CD, AA.
-The database presented in this article is released under public domain.
+The data presented in this article is released under public domain.
 Computer scripts are in git repositories and PyPI Python packages, also under public domain.
 Although most data is already in open licenses (Twitter, Email, Participabr, Cidade Democrática, and AA data), IRC and Facebook data was collected
-and donated by the individuals which yield the data.
-This rises the the understanding of the right to study such data as the right to access the self,
-in parity with anthropological endaviors~\cite{antphy,antphy2}.
+and donated by the individuals which yield the data, in a practice compliant with the framework described in Section~\ref{acq}.
 % Of central importance to the obtainment of the data is the concept of
@@
 \label{conclusions}
-The database presented in this article
-constitutes a large database with diverse provenance.
-Even so, database should be expanded in upon need or requests from feedback.
+This article outlines LOSD, a large human online social networking dataset (with almost 35 million triples) with diverse provenance.
+Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, are also described to the convenience of the newcommer
+and for a consistent exposition.
+The resulting dataset should be expanded upon need or request from interested parties.
 All data should be available online in the \url{http://linkedopensocialdata.org}
@@
 repertoire in current research.
-One should reach the diagrams and tables of the 
-Supporting Information document of this article
+All data is publicly available though the endpoint \url{xxxyyyzzz},
+through which the scientific community now has consistent and curated data
+that enables research validation, continuity, and benchmarking.
+The diagrams and tables of the Supporting Information document of this article
 for further directions
 on the available structures and for an overview complement.
+
+% further work: link participants in Facebook with those in ParticipaBR, for example
+% crowsorce more data
+% assure data is 5 starts TTM
 

### 2019-06-20 tables/nsnapshots.tex
adiantado um pouco do artigo

@@
 \caption{Number of snapshots from each provenance.}
-\begin{tabular}{| l | c |}\hline
-\textbf{social protocol} & \textbf{number of snapshots} \\\hline\hline
-Algorithmic Autoregulation & 3 \\\hline
-Cidade Democrática & 1 \\\hline
-Email & 4 \\\hline
-Facebook & 88 \\\hline
-IRC & 4 \\\hline
-ParticipaBR & 1 \\\hline
-Twitter & 16 \\\hline\hline
-all & 117 \\\hline
+\begin{tabular}{ l | c }
+\textbf{social protocol} & \textbf{number of snapshots} \\\hline
+Algorithmic Autoregulation & 3 \\
+Cidade Democrática & 1 \\
+Email & 4 \\
+Facebook & 88 \\
+IRC & 4 \\
+ParticipaBR & 1 \\
+Twitter & 16 \\\hline
+all & 117 \\
 \end{tabular}\end{center}

### 2019-06-20 scripts/dworld.py

@@ -1,0 +1,12 @@
+import datadotworld as d
+
+td = intro_dataset = d.load_dataset('rfabbri/test1')
+
+q = """SELECT * WHERE {?s ?p ?o}"""
+
+q2 = """
+PREFIX po: <http://purl.org/socialparticipation/po/>
+SELECT ?s WHERE {?s a po:Participant}
+"""
+
+r = d.query('rfabbri/test1', q2, query_type='sparql')

### 2019-06-29 mdpi/template.tex
iniciada passagem do artigo para o template do Data

--- 2019-06-20 paper.tex
@@
-\documentclass[review]{elsarticle}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+%=================================================================
+\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
 \usepackage{enumitem}
-\usepackage{hyperref}
-\usepackage[dvipsnames]{xcolor}
 \hypersetup{
@@
 }
+\usepackage{hyperref}
 \usepackage{multirow}
@@
 {
-    morecomment=[s][\color{ForestGreen}\bf]{<}{>},
+    morecomment=[s][\color{olive}\bf]{<}{>},
     basicstyle = \ttfamily\footnotesize,
     commentstyle=\color{green},
-    keywordstyle=\color{Plum},
-    morecomment=[s][\color{Blue}\bf]{?}{\ },
+    keywordstyle=\color{brown},
+    morecomment=[s][\color{blue}\bf]{?}{\ },
     morekeywords={BASE,PREFIX,SELECT,DISTINCT,CONSTRUCT,DESCRIBE,ASK,%
@@
 \usepackage{booktabs}
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%
+% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+\firstpage{1} 
 \makeatletter 
-\def\ps@pprintTitle{%
-    \let\@oddhead\@empty
-    \let\@evenhead\@empty
-    \def\@oddfoot{}%
-\let\@evenfoot\@oddfoot}
+\setcounter{page}{\@firstpage} 
 \makeatother
-\begin{document}
-%
-\begin{frontmatter}
-%
-\title{A Linked Open Social Dataset for Scientific Benchmarking}
-%
-\author[pws]{Renato Fabbri\corref{corresponding}}
-\ead{renato.fabbri@gmail.com}
-%
-\author[pwr]{Osvaldo Novais de Oliveira Junior}
-\ead{chu@ifsc.usp.br}
-%
-\cortext[corresponding]{Corresponding author}
-  \address[pws]{Institute of Mathematics and Computer Sciences, University of São Paulo (ICMC/USP), Brazil}
-\address[pwr]{S\~ao Carlos Institute of Physics, University of  São Paulo
-(IFSC-USP), Brazil}
-%
-\begin{abstract}
+\pubvolume{xx}
+\issuenum{1}
+\articlenumber{5}
+\pubyear{2019}
+\copyrightyear{2019}
+%\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+%\updates{yes} % If there is an update available, un-comment this line
+
+%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
+%\continuouspages{yes}
+
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
+
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{A Linked Open Social Dataset for Scientific Benchmarking}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{https://orcid.org/0000-0002-9699-629X} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{https://orcid.org/0000-0002-5399-5860} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Renato Fabbri $^{1}$*\orcidA{} and Osvaldo N. Oliveira Junior $^{2}$\orcidB{}}
+
+% Authors, for metadata in PDF
+\AuthorNames{Renato Fabbri, Osvaldo Novais de Oliveira Junior}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Institute of Mathematical and Computer Sciences, University of São Paulo (ICMC/USP), São Carlos, Brazil; renato.fabbri@gmail.com\\
+$^{2}$ \quad São Carlos Institute of Physics, University of São Paulo (IFSC/USP), São Carlos, Brazil; chu@ifsc.usp.br}
+
+% Contact information of the corresponding author
+\corres{Correspondence: renato.fabbri@gmail.com}
+
+% Current address and/or shared authorship
+% \firstnote{Current address: Affiliation 3} 
+% \secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{
 The fields of social network analysis and complex networks
@@
 diverse data, most often not accessible to researchers other than the publishing authors.
-This work presents an open dataset with diverse provenance of 
-human social online networking 
-  and oriented
-to furnish the scientific community with a friendly and common repertoire.
+In order to furnish the scientific community with a common repertoire,
+this work presents an open dataset with diverse provenance of 
+human social online networking.
 Current data was obtained in 2013-16 from well-known online social networking platforms:
 Facebook, Twitter, IRC, Email.
-Data from ParticipaBR, AA, and Cidade Democrática were also added as samples of
-less known instances.
-Authors recently recognized that no similar dataset was made available,
-and the potential benefit of the content for the research community.
+Data from ParticipaBR, AA, and Cidade Democrática are also supplied as samples of less known instances.
 The linked data representation was adopted in order to
-comply with current best practices, homogenize access, enable discovery,
-and facilitate analyzes which integrate third
-party and provided instances.
+comply with current best practices, homogenize access, facilitate discovery
+and analyzes which integrate third party and provided instances.
 Futhermore, the method for obtaing and making the data available is potentially novel, and may of use for other research groups.
-This document presents an outline and overall statistics of the given
-dataset, which should favor subsequent work in complex and social networks,
-  and in enhancing the available linked and social data.
-\end{abstract}
-%
-\begin{keyword}
-Big Data, Dataset, Benchmark Data, Facebook, Twitter, IRC, Email,
-Complex Networks, Social Networks, Text Mining
-%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
-\end{keyword}
+This document presents a description, usage remarks and overall statistics of the dataset, which should favor subsequent work in complex and social networks
+  and enhancements to the dataset.
+}
 
-\end{frontmatter}
+% Keywords
+\keyword{social networks; complex networks; linked data; semantic web; big data; benchmark data; Facebook; Twitter; irc; email}
 
-\section{Introduction}
-% In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
-% both research and application realms.
-The research on human social and complex networks often realy on data that express
-real phenomena~\cite{c1,c2,c3}.
-Such data is often not made publicly available, which hinders validation of the results and does not favor makes subsequent work.
-Furthermore, there is a lack of open datasets for benchmarking results,
-specially associated to the complex and social networks field,
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+\dataset{https://rfabbri.linked.data.world/d/linked-open-social-data/}
+
+\datasetlicense{CC0}
+
+\begin{document}
+\section{Summary}
+The research on human social and complex networks often rely on data that express real phenomena~\cite{c1,c2,c3}.
+Such data is regularly not made publicly available,
+which hinders validation of the results
+and does not favor subsequent work.
+Furthermore, there is a lack of open datasets for benchmarking results
+related to complex and social networks,
 yielding diverse results from poorly related sources.
-Scarce and historical exceptions are the karate club or the dolphin datasets~\cite{newmanBook,dolphin}.
-For expressing the information,
-open machine-readable content and tools are desired for enabling
-data linkage and discovery~\cite{o1,o2},
-and the favoured protocol is thus RDF ``linked data''.
-Social networks linked data is currently not stable or rigorous enough to be
-a public reference on statistical physics and social networks research,
-with scarce exceptions~\cite{l1,l2}
+Social networks linked data is currently scarce and not stable
+or rigorous enough to be a public reference on statistical physics
+and social networks research~\cite{l1,l2}.
+
 This work presents a collection of linked and open data
-about online human social networks,
-named LOSD (Linked Open Social Data).
-This has been achieved by linking social data into a dataset using diverse
+derived from online human social networks.
+The dataset is named LOSD (Linked Open Social Data),
+and has been achieved by linking social data into a dataset derived from diverse
 provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
@@
 research involving networks and textual content.
+In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ), and should be used in further research in complex networks and data visualization (FAPESP grant ) which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
 
@@
 In~\cite{fb1}, data obtained only from Facebook is reported.
-With such a focus, the authors make available a number of statistics,
+With such a focus, the authors made available a number of statistics,
 but the data is offline at least since 2010.
@@
 potentially because researchers did not have the means to acquire and
-share the data described in Section~\ref{acq}.
+share the data as described in Section~\ref{acq}.
 
-% \subsection{Benchmarking in the analysis of complexity}
+\section{Data Description}\label{materials}
+\subsection{Data outline}
+\input{../misc/overallText}
 
+\input{../tables/nsnapshots}
+\label{outline}
+Current overall results concern data selection and preparation for data discovery.
+The main result is the data made available, which enables benchmarking of scientific results
+and easy experimentations, and no compliant dataset was found by the authors.
+Secondary results include data outline through figures and tables,
+software support and example SparQL queries.
+
+\subsection{Standardization}
+The data is embedded into URIs and triples, i.e. translated to RDF.
+URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
+which are identified herein with the prefix ``\textttt{po:}''.
+Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
+Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
+All class instances, such as participants, messages, friendships and
+interactions, are linked to
+snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
+Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
+Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
+More specialized predicates are used for delivering text when necessary,
+such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
+for ParticipaBR articles.
+A participant URI is unique throughout the provenance (e.g. the same for
+the same participant in all Twitter snapshots).
+To enable annotations which differ when the snapshot changes,
+\texttt{po:Observation} class instances are used in the triple
+\textttt{<participant\_uri> po:observation <observation\_uri>}.
+The observation instances are then linked to the snapshot and the
+data.
+
+Instance URIs are built using the URI from the class they derive from plus a hashtag character,
+a provenance string (e.g. \textttt{facebook-legacy} or
+\textttt{participabr-legacy}), and an identifier;
+i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
+All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
+All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
+\textttt{irc-legacy-labmacambira} or
+\textttt{email-legacy-linux.audio.devel1-20000}.
+
+Data was gathered from:
+\begin{itemize}
+    \item public APIs (Twitter, Email);
+    \item public logs (IRC and AA);
+    \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
+    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
+\end{itemize}
+
+Integration and uniformity of access is obtained through linked data
+representation, as described in Section~\ref{queries}.
+This section introduces the underlying data in very concise terms.
+One should access the Supporting Information document for
+a more information about the dataset.
+
+\subsection{Snapshots}
+Of central importance to presented database is the concept of a snapshot.
+A snapshot is herein a set of data gathered together, at a contiguous time
+unit.
+Examples: the first 20 thousand email messages of an email list
+comprises a snapshot; the tweets from the MAMA music event is a
+snapshot; the friendship, interaction and posts structures of a facebook
+group, prospected at the same time, is a snapshot.
+
+\subsection{Facebook data}
+Friendship ego networks (networks whose nodes represent friends of a user)
+were donated from individual users in 2013 and 2014.
+Friendship and interaction networks from groups were gathered from
+groups where the first author was a participant.
+Additionally, some groups have post texts along some metadata, such as
+the number of likes.
+
+\subsection{Twitter data}
+Tweets were gathered through the Twitter streaming public API.
+Each snapshot is unified by a distinct hashtag.
+Network links are canonically yield by retweets,
+but replies and user mentions are also kept in the database.
+
+\subsection{IRC data}
+Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
+The database has records of relations yield by directed messages and
+mentions.
+
+\subsection{Email data}
+Email snapshots refer to individual email lists.
+All messages were obtained from the Gmane public email archive~\cite{gmane}.
+Each message has the original text and the text without some of the lines
+from previous messages or that are software code or author signature.
+Most importantly, each message instance holds the ID of the message it is
+a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}
+
+\subsection{ParticipaBR data}
+ParticipaBR was a Brazilian federal platform for social participation,
+the most important platform from the myriad of social participation platforms
+that were available in 2011-2018~\cite{spbr}.
+Texts are derived from blog posts and networks may be obtained from
+friendship and interaction criteria.
+
+\subsection{Cidade Democr\'atica data}
+Cidade Democrática is a Brazilian civil society social participation portal.
+Data gathered is complex in the number of types of instances.
+
+\subsection{AA data}
+AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
+it is based on testifying and sharing ongoing work.
+The data was gathered from different versions of the system and from an IRC
+log.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
@@
 
-\section{Materials}\label{materials}
-Data was gathered from:
-\begin{itemize}
-    \item public APIs (Twitter, Email);
-    \item public logs (IRC and AA);
-    \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
-    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
-\end{itemize}
-
-Integration and uniformity of access is obtained through linked data
-representation, as described in Section~\ref{queries}.
-This section introduces the underlying data in very concise terms.
-One should access the Supporting Information document for
-a more information about the dataset.
-
-\subsection{Snapshots}
-Of central importance to presented database is the concept of a snapshot.
-A snapshot is herein a set of data gathered together, at a contiguous time
-unit.
-Examples: the first 20 thousand email messages of an email list
-comprises a snapshot; the tweets from the MAMA music event is a
-snapshot; the friendship, interaction and posts structures of a facebook
-group, prospected at the same time, is a snapshot.
-
-\subsection{Facebook data}
-Friendship ego networks (networks whose nodes represent friends of a user)
-were donated from individual users in 2013 and 2014.
-Friendship and interaction networks from groups were gathered from
-groups where the first author was a participant.
-Additionally, some groups have post texts along some metadata, such as
-the number of likes.
-
-\subsection{Twitter data}
-Tweets were gathered through the Twitter streaming public API.
-Each snapshot is unified by a distinct hashtag.
-Network links are canonically yield by retweets,
-but replies and user mentions are also kept in the database.
-
-\subsection{IRC data}
-Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
-The database has records of relations yield by directed messages and
-mentions.
-
-\subsection{Email data}
-Email snapshots refer to individual email lists.
-All messages were obtained from the Gmane public email archive~\cite{gmane}.
-Each message has the original text and the text without some of the lines
-from previous messages or that are software code or author signature.
-Most importantly, each message instance holds the ID of the message it is
-a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}
-
-\subsection{ParticipaBR data}
-ParticipaBR was a Brazilian federal platform for social participation,
-the most important platform from the myriad of social participation platforms
-that were available in 2011-2018~\cite{spbr}.
-Texts are derived from blog posts and networks may be obtained from
-friendship and interaction criteria.
-
-\subsection{Cidade Democr\'atica data}
-Cidade Democrática is a Brazilian civil society social participation portal.
-Data gathered is complex in the number of types of instances.
-
-\subsection{AA data}
-AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
-it is based on testifying and sharing ongoing work.
-The data was gathered from different versions of the system and from an IRC
-log.
-
-\section{Results}
-\label{outline}
-Current overall results concern data selection and preparation for data discovery.
-The main result is the data made available, which enables benchmarking of scientific results
-and easy experimentations, and no compliant dataset was found by the authors.
-Secondary results include data outline through figures and tables,
-software support and example SparQL queries.
-
-\subsection{Standardization}
-The data is embedded into URIs and triples, i.e. translated to RDF.
-URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
-which are identified herein with the prefix ``\textttt{po:}''.
-Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
-Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
-All class instances, such as participants, messages, friendships and
-interactions, are linked to
-snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
-Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
-Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
-More specialized predicates are used for delivering text when necessary,
-such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
-for ParticipaBR articles.
-A participant URI is unique throughout the provenance (e.g. the same for
-the same participant in all Twitter snapshots).
-To enable annotations which differ when the snapshot changes,
-\texttt{po:Observation} class instances are used in the triple
-\textttt{<participant\_uri> po:observation <observation\_uri>}.
-The observation instances are then linked to the snapshot and the
-data.
-
-Instance URIs are built using the URI from the class they derive from plus a hashtag character,
-a provenance string (e.g. \textttt{facebook-legacy} or
-\textttt{participabr-legacy}), and an identifier;
-i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
-All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
-All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
-\textttt{irc-legacy-labmacambira} or
-\textttt{email-legacy-linux.audio.devel1-20000}.
-
-\subsection{Data outline}
-\input{misc/overallText}
-
-\input{tables/nsnapshots}
-% stats:
-%% number of triples
-%% number of edges
-%% number of chars
-%% number of users
-% diagrams in the supporting information file
+\section{Usage notes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt} 
 \subsection{Software tools}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{ontologies/facebook-legacy-AntonioAnzoategui18022013Friendship.ttl/draw}
+    \includegraphics[width=0.5\textwidth]{../ontologies/facebook-legacy-AntonioAnzoategui18022013Friendship.ttl/draw}
     \caption{A diagram of the core structure that entails the friendship networks
@@
 For a less canonical example for obtaining networks, links may be derived from sufficiently similar vocabulary used by participants, or from activity is sufficiently similar dates, which yield networks which are not interaction or friendship networks.
-% mk figure with two participants and their relation though response, retweet, or friendship
+% mk figure with two participants and their relation though response, retweet, or friendship TTM
 
-\section{License issues}
-% Facebook, Twitter, IRC, Gmane, Participa, CD, AA.
-The data presented in this article is released under public domain.
-Computer scripts are in git repositories and PyPI Python packages, also under public domain.
-Although most data is already in open licenses (Twitter, Email, Participabr, Cidade Democrática, and AA data), IRC and Facebook data was collected
-and donated by the individuals which yield the data, in a practice compliant with the framework described in Section~\ref{acq}.
-% Of central importance to the obtainment of the data is the concept of
-% anthropological physics~\cite{antphy}
-\section{Conclusions}
+\section{Conclusions and further work}
 \label{conclusions}
@@
 
-% further work: link participants in Facebook with those in ParticipaBR, for example
-% crowsorce more data
-% assure data is 5 starts TTM
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
 
-% Supporting Information
-% diverse provenance, large database, should be expanded, waits feedback
-\section*{References}
-%
-% \bibliographystyle{plainurl}
-\bibliography{paper}
-%\bibliography{myLastBibfile.bib}
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Conceptualization, writing, formal analysis, R.F; senior supervision, final writing, O.N.O.Jr.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\funding{This research was funded by CNPq (grant number XXX) and FAPESP (grant number ).}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acknowledgments{The authors thank all participants in the experiments which  donated the data they gathered as described in Section~\ref{}.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conflictsofinterest{The authors declare no conflict of interest.
+The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendix
+\section{}
+\unskip
+\subsection{}
+The appendix is an optional section that can contain details and data supplemental to the main text.
+For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+\section{}
+All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\externalbibliography{yes}
+\bibliography{../paper}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+% \begin{thebibliography}{999}
+% % Reference 1
+% \bibitem[Author1(year)]{ref-journal}
+% Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% % Reference 2
+% \bibitem[Author2(year)]{ref-book}
+% Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% \end{thebibliography}
+
+% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2019-06-30 mdpi/template.tex
adiantada adaptacao p revista data

@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[data,datadescriptor,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 % Full title of the paper (Capitalized)
-\Title{A Linked Open Social Dataset for Scientific Benchmarking}
+\Title{LOSD: A Linked Open Social Dataset for Scientific Benchmarking}
 
@@
 diverse data, most often not accessible to researchers other than the publishing authors.
-In order to furnish the scientific community with a common repertoire,
+In order to provide the scientific community a common repertoire,
 this work presents an open dataset with diverse provenance of 
@@
 comply with current best practices, homogenize access, facilitate discovery
-and analyzes which integrate third party and provided instances.
-Futhermore, the method for obtaing and making the data available is potentially novel, and may of use for other research groups.
-This document presents a description, usage remarks and overall statistics of the dataset, which should favor subsequent work in complex and social networks
-  and enhancements to the dataset.
+and analyzes which integrate third party and provided data.
+Futhermore, the method for obtaing and making the data available is potentially novel,
+and may be of use for other research groups.
+This document presents a description, usage remarks and overall statistics of the dataset,
+which should favor subsequent work in complex and social networks and enhancements to the dataset.
 }
@@
 \section{Summary}
-The research on human social and complex networks often rely on data that express real phenomena~\cite{c1,c2,c3}.
+The research on human social and complex networks often
+rely on data that express real phenomena~\cite{c1,c2,c3}.
 Such data is regularly not made publicly available,
@@
 
-This work presents a collection of linked and open data
+This work presents a linked and open dataset
 derived from online human social networks.
@@
 provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
-Such data is proposed as a common and public repertoire for scientific
-research involving networks and textual content.
-In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ), and should be used in further research in complex networks and data visualization (FAPESP grant ) which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
+Such data may be of use as a common and public repertoire for scientific
+research involving complex networks, specially if social networks and textual content are pinpointed.
+In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ),
+and should be further used e.g. in researching complex networks and data visualization (FAPESP grant )
+which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
+
+Data collection was performed by a potentially innovative method which places the researcher
+at the core of data collection, and keeps the resources (e.g. code, data, documents) open,
+in order to ammeliorate ethical issues involved in studying human social systems.
+
+This paper is organized as follows. Next subsection briefly discusses related work.
+Section~\ref{materials} describes the data currently available.
+Section~\ref{methods} describes the methods used for data gathering, translation into RDF linked data, and the synthesis of auxiliary structures.
+Section~\ref{usage} contains direction for using the data.
 
@@
 Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
-on specialized communities such as related to research and movies~\cite{s1,s2,s3,s4,s5},
+on specialized communities such as related to research or movies~\cite{s1,s2,s3,s4,s5},
 or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
@@
 There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
-most often with a specific focus (users, tweets, hashtags, etc)
+most often with a specific focus (users, retweets, hashtags, etc)
 and not presented to the scientific community by means of a research article.
@@
 \section{Data Description}\label{materials}
-\subsection{Data outline}
-\input{../misc/overallText}
+\subsection{Data outline}\label{outline}
+% scripts/percolationUse2.py
+The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters.
+Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.
+Appendix~\ref{aoverall} holds the queries used to obtain such data.
+This section describes de data in LOSD in very brief terms.
+The Supplementary document is provided for further information about the dataset.
 
-\input{../tables/nsnapshots}
-\label{outline}
-Current overall results concern data selection and preparation for data discovery.
-The main result is the data made available, which enables benchmarking of scientific results
-and easy experimentations, and no compliant dataset was found by the authors.
-Secondary results include data outline through figures and tables,
-software support and example SparQL queries.
-
-\subsection{Standardization}
-The data is embedded into URIs and triples, i.e. translated to RDF.
-URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
-which are identified herein with the prefix ``\textttt{po:}''.
-Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
-Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
-All class instances, such as participants, messages, friendships and
-interactions, are linked to
-snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
-Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
-Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
-More specialized predicates are used for delivering text when necessary,
-such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
-for ParticipaBR articles.
-A participant URI is unique throughout the provenance (e.g. the same for
-the same participant in all Twitter snapshots).
-To enable annotations which differ when the snapshot changes,
-\texttt{po:Observation} class instances are used in the triple
-\textttt{<participant\_uri> po:observation <observation\_uri>}.
-The observation instances are then linked to the snapshot and the
-data.
-
-Instance URIs are built using the URI from the class they derive from plus a hashtag character,
-a provenance string (e.g. \textttt{facebook-legacy} or
-\textttt{participabr-legacy}), and an identifier;
-i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
-All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
-All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
-\textttt{irc-legacy-labmacambira} or
-\textttt{email-legacy-linux.audio.devel1-20000}.
-
-Data was gathered from:
+\noindent Data was gathered from:
 \begin{itemize}
@@
 
-Integration and uniformity of access is obtained through linked data
-representation, as described in Section~\ref{queries}.
-This section introduces the underlying data in very concise terms.
-One should access the Supporting Information document for
-a more information about the dataset.
-
 \subsection{Snapshots}
-Of central importance to presented database is the concept of a snapshot.
+Of central importance to the presented dataset is the concept of a snapshot.
 A snapshot is herein a set of data gathered together, at a contiguous time
@@
 group, prospected at the same time, is a snapshot.
+Table~\ref{tsnap} holds the number of snapshots from each provenance in LOSD.
+
+\input{../tables/nsnapshots}
 
 \subsection{Facebook data}
-Friendship ego networks (networks whose nodes represent friends of a user)
+Friendship ego networks (networks whose nodes represent friends of a user and links represent friendships)
 were donated from individual users in 2013 and 2014.
@@
 but replies and user mentions are also kept in the database.
+LOSD kept the main data fields returned by Twitter API, such as creation date, text, and hashtags.
 
@@
 Public IRC (Internet Relay Chat) logs were used to obtain IRC snapshots.
-The database has records of relations yield by directed messages and
-mentions.
+LOSD holds records of relations yield by directed messages and
+mentions and the text related to each message.
 
@@
 Most importantly, each message instance holds the ID of the message it is
-a reply to, if any, which enables the achievement of interaction networks~\cite{bird,stab}
+a reply to, if any, which enables the achievement of canonical interaction networks~\cite{bird,stab}.
 
 \subsection{ParticipaBR data}
-ParticipaBR was a Brazilian federal platform for social participation,
+ParticipaBR is a Brazilian federal platform for social participation,
 the most important platform from the myriad of social participation platforms
@@
 Texts are derived from blog posts and networks may be obtained from
-friendship and interaction criteria.
+both friendship and interaction criteria.
 
-\subsection{Cidade Democr\'atica data}
+\subsection{Cidade Democrática data}
 Cidade Democrática is a Brazilian civil society social participation portal.
-Data gathered is complex in the number of types of instances.
+Data gathered is complex in the number of types of instances as noticeable in the Supplementary document.
 
 \subsection{AA data}
-AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications,
-it is based on testifying and sharing ongoing work.
+AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications by sharing sentences about ongoing work and testifying about their quality.
 The data was gathered from different versions of the system and from an IRC
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
+\section{Methods}\label{methods}
 LOSD is expressed as linked data through RDF
@@
 \subsection{Data acquisition}\label{acq}
-Research involving human subjects raise well-known ethic issues~\cite{}.
+Research involving human subjects raise well-known ethic issues~\cite{eth}.
 Furthermore, the online social networking platforms often restrict access
 to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
-In this context, research using data, e.g. from Facebook, is considered
-problematic if not prohibited.
+Although Twitter and GMANE (email) data is available through public APIs,
+the other sources are more problematic.
+In this context, research using data, e.g. from Facebook, is precarious
+if not prohibited.
 We reached a potential solution to this issue through continuated research with
-collaborators in the fields of phylosophy, social sciences, and psychology.
+collaborators in the fields of phylosophy, social sciences, antropology and psychology.
 The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
@@
 and keeping the procedures, data, algorithms and results publicly available.
-This practices enabled collecting and sharing the data in LOSD.
-Notice that, although Twitter data is available through public APIs, the other sources are more problematic.
+These practices enabled collecting and sharing the data in LOSD,
+and was called ``Anthropological Physics'' because it relies in the
+etnographic tecnique of investigating the social structures
+by studying the researcher, as to ammeliorate the ethic issues.
+Accordingly, aside from the data acquired from Twitter, all data in LOSD
+was gathered from social instances where the first author is a participant.
+For example, all Facebook ego networks were donated to the first author by
+users which know him and were interested in his research and intervention on the networks,
+and all email and Facebook groups data was obtained from groups where the first author is a member.
 
-% download via netviz when fb allowed it, donations
-% twitter, bots, gmane
-% contact with developers of AA, PBR, CD
-% Anthropological physics
+Facebook data was obtained through the Netvizz software~\cite{netviz},
+Facebook allowed downloading friendships and interaction data as available in LOSD.
+Data from Twitter was obtained through the Twitter streamming API.
+Data From IRC was obtained through bots that kept logs of the messages.
+Email data was obtained through the GMANE public API.
+Data from ParticipaBR, Cidade Democrática, and AA was obtained through contact
+of the researcher with the developers of the platforms.
 
@@
 In practice, the publication of linked data can me summarized
-by 1) the use of RDF to publish data on the web and 2) the use of RDF
+by 1) the use of RDF and 2) the use of RDF
 links to interlink data from different sources~\cite{ld1,ld2}.
-The web is expected to be interconnected and to grow by the systematic application of four
+The web is expected to be interconnected and to
+grow by the systematic application of four
 steps~\cite{lee1}:
@@
 
-The Linked Open Data~\cite{lod} is a constantly growing cloud of data,
+The Linked Open Data~\cite{lod} cloud is a constantly growing cloud of data,
 the global data space, which is usually
@@
 \subsection{Data-driven ontology synthesis}\label{ont}
-OWL Ontologies are tools to describe taxonomies and the
-structure of knowledge~\cite{owl1}.
 Most ontologies are created by domain experts even though the data they
 arrange is often given by a software system and has a predefined
-structure~\cite{owl2}.
-
+structure~\cite{owl1,owl2}.  
 We developed a simple ontology synthesis method that probes
@@
 The results are OWL code and diagrams which are available in the
-Supporting Information document of this article.
+Supplementary document of this article.
 The method can be extended to comprise further OWL axioms and restrictions,
@@
 
-\section{Usage notes}
+\subsection{Standardization}
+The data is embedded into URIs and triples, i.e. translated to RDF.
+URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
+which are identified herein with the prefix ``\textttt{po:}''.
+Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
+Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
+All class instances, such as participants, messages, friendships and
+interactions, are linked to
+snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
+Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
+Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
+More specialized predicates are used for delivering text when necessary,
+such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
+for ParticipaBR articles.
+A participant URI is unique throughout the provenance (e.g. the same for
+the same participant in all Twitter snapshots).
+To enable annotations which differ when the snapshot changes,
+\texttt{po:Observation} class instances are used in the triple
+\textttt{<participant\_uri> po:observation <observation\_uri>}.
+The observation instances are then linked to the snapshot and the
+data.
+
+Instance URIs are built using the URI from the class they derive from plus a hashtag character,
+a provenance string (e.g. \textttt{facebook-legacy} or
+\textttt{participabr-legacy}), and an identifier;
+i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
+All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
+All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
+\textttt{irc-legacy-labmacambira} or
+\textttt{email-legacy-linux.audio.devel1-20000}.
+
+\section{Usage notes}\label{usage}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-30 misc/overallText.tex
adiantada adaptacao p revista data

@@
 Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 49 have network interaction links, 89 have network friendship links; 43 have text content from messages.
+
+# The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters.
+Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.

### 2019-07-01 mdpi/template.tex
arrumados trabalhos relacionados e bibliografia

@@
 yielding diverse results from poorly related sources.
-Social networks linked data is currently scarce and not stable
-or rigorous enough to be a public reference on statistical physics
-and social networks research~\cite{l1,l2}.
-
-This work presents a linked and open dataset
-derived from online human social networks.
-The dataset is named LOSD (Linked Open Social Data),
-and has been achieved by linking social data into a dataset derived from diverse
+Social networks linked datasets are currently most often restricted to a specific platform or express statistics of social networks, as described in Section~\ref{srel}.
+This work presents 
+a dataset named LOSD (Linked Open Social Data)
+that has been achieved by linking social data derived from diverse
 provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
@@
 research involving complex networks, specially if social networks and textual content are pinpointed.
-In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant ),
-and should be further used e.g. in researching complex networks and data visualization (FAPESP grant )
+In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant 140860/2013-4),
+and should be further used e.g. in researching complex networks and data visualization (FAPESP grant 2017/05838-3)
 which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
-
 Data collection was performed by a potentially innovative method which places the researcher
@@
 Section~\ref{methods} describes the methods used for data gathering, translation into RDF linked data, and the synthesis of auxiliary structures.
-Section~\ref{usage} contains direction for using the data.
+Section~\ref{usage} holds directions for using the data.
 
-\subsection{Related work}
-Most available datasets on human social networks rely on only one source~\cite{o1,o2,o3,o4},
-on specialized communities such as related to research or movies~\cite{s1,s2,s3,s4,s5},
-or present only statistics about social networks, and not the data that yields the networks~\cite{st1,st2,st3,st4}.
-Contributions which are most similar to LOSD are mentioned in this section.
-In~\cite{fb1}, data obtained only from Facebook is reported.
-With such a focus, the authors made available a number of statistics,
-but the data is offline at least since 2010.
+\subsection{Related work}\label{srel}
+Most available datasets on human social networks rely only on one source (e.g.~\cite{nat,fb1}),
+on specialized communities such as related to research or movies (e.g.~\cite{s1},
+or present only statistics about social networks, and not the data that yields the networks (e.g.~\cite{st1}.
 Linked human social network data is scarcely found in the scientific community.
@@
 dedicated to free software developers linked by trust relatioships.
-There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2,t3,t4}),
+There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2}),
 most often with a specific focus (users, retweets, hashtags, etc)
 and not presented to the scientific community by means of a research article.
-In summary, no human social network dataset presented to the
+The datasets which are most closely related to LOSD are KONECT~\cite{konect},
+SNAP Datasets~\cite{snapnets}, the Network Repository~\cite{nr}, and ICON~\cite{icon}.
+None of them are (RDF) linked data but express the networks in formats specialized for networks.
+In total, these datasets make available thousands of social networks, from diverse provenance and of various types (e.g. simple, directed, weighted, bipartite, temporal networks).
+Compared to LOSD, KONECT, SNAP, Network Repository, and ICON are complementary, providing a further diversity of networks, but only as simplified data, i.e.
+in specialized formats, most often expressed as columns of node ids that are connected and metadata (such as timestaps for temporal networks).
+In summary, no linked human social network dataset presented to the
 scientific community containts the diverse provenance used in LOSD,
+or provides such a wide range of metadata,
 potentially because researchers did not have the means to acquire and
-share the data as described in Section~\ref{acq}.
+share the data as described in Section~\ref{acq},
+and because linked data standards are (still) not widely adopted in 
+every research domain.
 
@@
 Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.
-Appendix~\ref{aoverall} holds the queries used to obtain such data.
+Section~\ref{queries} holds the queries used to obtain such data.
 This section describes de data in LOSD in very brief terms.
@@
 \subsection{Data acquisition}\label{acq}
-Research involving human subjects raise well-known ethic issues~\cite{eth}.
+Research involving human subjects raise well-known ethical issues,
+although ethical regulation for (non-medical) research in social systems is not always endorsed by scientists~\cite{eth}.
 Furthermore, the online social networking platforms often restrict access
-to the data and declare illegal the use of their data through web crowlers~\cite{ile1,ile2}.
+to the data and user may consider illegal the use of their data through web crowlers~\cite{ile1,ile2,ile3}.
 Although Twitter and GMANE (email) data is available through public APIs,
@@
 recommendation, is a model for data
-interchange~\cite{w3cr}.
+interchange.
 It is based on the idea of making statements about resources in the form
@@
 directed multi-graph.
-RDF may be stored in a type of database called a triplestore~\cite{rdf}.
+RDF may be stored in a type of database called a triplestore.~\cite{rdf}
 
@@
 with which all the expressive capabilities of RDFS are
-available, but one can also, among other things~\cite{owl}:
+available, but one can also, among other things~\cite{ont}:
 \begin{itemize}
@@
 
-For the data described in this article, the following features are most relevant~\cite{rdfs,owl}:
+For the data described in this article, the following features are most relevant~\cite{rdfs,ont}:
 \begin{itemize}
@@
 \subsection{Data-driven ontology synthesis}\label{ont}
-Most ontologies are created by domain experts even though the data they
-arrange is often given by a software system and has a predefined
-structure~\cite{owl1,owl2}.  
+Most ontologies are created by domain experts~\cite{ont}
+even though the data they arrange is often given by a software system
+and has thus a predefined structure.  
 We developed a simple ontology synthesis method that probes
@@
 the steps implemented are as follows:
-\begin{enumerate}[leftmargin=0cm]
-    \item Obtain all distinct classes with the query:\\
-        \h\textttt{SELECT DISTINCT ?class\_uri WHERE \{ ?s a ?class\_uri .\}}
-\item For each class, obtain the properties that occur as predicates in triples where the subject is an instance of the class:\\
-	\h\textttt{SELECT DISTINCT ?property\_uri WHERE \{ ?s a <class\_uri> . ?s ?property\_uri ?o .\}}\\
-Such properties are used to assert existential and universal restrictions for the class.
-\item Compare the total number of individuals (\textttt{?cs1}) of the class (\textttt{class\_uri}) with
-	the number of such individuals (\textttt{?cs2}) that are subjects of at least one triple where 
-        the predicate is the property (\textttt{property\_uri}) and the object is
-        an instance of the same class (\textttt{class2\_uri}).
-	The queries are:\\
-	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE \{ ?s a <class\_uri> \}}\\
-        \h\textttt{SELECT DISTINCT ?oc WHERE \{\\
-	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a ?oc . \\ 
-        \h\}}\\
-        Then, for each object class (\textttt{object\_curi}) in \textttt{?oc}:\\
-	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs2) WHERE \{\\
-	\h\h ?s a <class\_uri>. ?s <property\_uri> ?o . ?o a <object\_curi> .\\         \h\}}\\
-        When \textttt{?cs1} matches \textttt{?cs2}, an existential restriction is found. And with the query:
-	\h\textttt{SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE \{\\
-	\h\h ?s a <class\_uri> . ?s <property\_uri> ?o . ?o a ?oc .\\
-        \h\h FILTER(str(?oc) != 'object\_curi')\\
-        \h\}}\\
-       When \textttt{?cs} is 0, a universal restriction is found. 
-\item To keep a record of further restrictions, get all object classes or datatypes where the subject is an instance of the class and the predicate is the property:\\
-	\h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
-        \h\h ?s a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
-        \h\}}
-    \item Obtain all distinct properties:\\
-        \h\textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
-    \item Check if each property is functional, i.e. if it
+\begin{enumerate}
+    \item Obtain all distinct classes with the query:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?class_uri WHERE { ?s a ?class_uri .}
+\end{lstlisting}
+\item For each class, obtain the properties that occur as predicates in triples where the subject is an instance of the class:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?property_uri WHERE { ?s a <class_uri> . ?s ?property_uri ?o .}
+\end{lstlisting}
+\item Compare the total number of individuals (\te{?cs1}) of the class (\te{class\_uri}) with
+	the number of such individuals (\te{?cs2}) that are subjects of at least one triple where 
+        the predicate is the property (\te{property\_uri}) and the object is
+        an instance of the same class (\te{class2\_uri}).
+	The queries are:
+\begin{lstlisting}[language=spq]
+SELECT (COUNT(DISTINCT ?s) as ?cs1) WHERE { ?s a <class_uri> }
+SELECT DISTINCT ?oc WHERE {
+  ?s a <class_uri> . ?s <property_uri> ?o . ?o a ?oc .
+}
+\end{lstlisting}
+        Then, for each object class (\textttt{object\_curi}) in \textttt{?oc}:
+\begin{lstlisting}[language=spq]
+SELECT (COUNT(DISTINCT ?s) as ?cs2) WHERE {
+  ?s a <class_uri> . ?s <property_uri> ?o . ?o a <object_curi> .
+}
+\end{lstlisting}
+        When \te{?cs1} matches \te{?cs2}, an existential restriction is found.
+        And with the following query, when \te{?cs} is 0, a universal restriction is found:
+\begin{lstlisting}[language=spq]
+SELECT (COUNT(DISTINCT ?s) as ?cs) WHERE {
+  ?s a <class_uri> . ?s <property_uri> ?o . ?o a ?oc .
+  FILTER(str(?oc) != 'object_curi')
+}
+\end{lstlisting}
+  \item Retrieve all object classes or datatypes where the subject is an instance of the class and the predicate is the property:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE {
+  ?s a <class_uri>. ?s <property_uri> ?o . OPTIONAL { ?o a ?co . }
+}
+\end{lstlisting}
+  \item Obtain all distinct properties:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?p WHERE { ?s ?p ?o }
+\end{lstlisting}
+  \item For each property, check if it is functional, i.e. if it
         occurs at most once with each subject.
         This is performed by counting the objects and further verifying
-        that they are at most one. The query is:\\
-        \h\textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
-            <property\_uri> ?o \} GROUP BY ?s}
+        that they are at most one. The query is:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT (COUNT(?o) as ?co) WHERE {
+  ?s <property_uri> ?o 
+} GROUP BY ?s
+\end{lstlisting}
   \item For each property, find the incident range (possible subjects) and domain (possible objects) with the
-        queries:\\
-        \h\textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{\\
-        \h\h ?s <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \}\\
-        \h\}} \\
-        and \\
-        \h\textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
-%     \item For each instance of each class, get all distinct predicates.
-%         For each predicate, check if all instances of the class
-%         hold such relationship (existential restriction):\\
-%         \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
-%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
-%         \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
-%                 a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
-%     \item and if all instances that hold such relationship are instances of the class
-%         (universal restriction):\\
-%         \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
-%     \item Draw each class, each property and the overall figure.
+        queries:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE {
+  ?s <property_uri> ?o . OPTIONAL { ?o a ?co . }
+} 
+SELECT DISTINCT ?cs WHERE { ?s <property_uri> ?o . ?s a ?cs . }
+\end{lstlisting}
   \item Render diagrams to achieve figures as exemplified in Section~\ref{sdia} and in the Supporting Information document.
-%      \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
-%          statements to better organize knowledge and link to third party
-%          ontologies and data. This step is left aside in current LOSD as 
 \end{enumerate}
@@
 \subsection{Standardization}
-The data is embedded into URIs and triples, i.e. translated to RDF.
+LOSD data is represented through URIs and triples by translation of various formats to RDF.
 URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
-which are identified herein with the prefix ``\textttt{po:}''.
-Classes and properties are built by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
+which are identified herein with the prefix ``\te{po:}''.
+Classes and properties are built by adding a suffix to the root, as in \te{po:Participant} or \te{po:text}.
 Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
@@
 interactions, are linked to
-snapshots through the triple \textttt{<instance> po:snapshot <snapshot\_uri>}.
-Message texts, including comments, are objects in the triple: \textttt{<message\_id> po:text <message\_text>}.
-Preprocessed texts are objects of triples: \textttt{<message\_id> po:cleanText <message\_text>}.
-More specialized predicates are used for delivering text when necessary,
-such as \textttt{po:htmlBodyText} and \textttt{po:cleanBodyText} used
+snapshots through the triple \te{<instance> po:snapshot <snapshot\_uri>}.
+Message texts, including comments, are objects in a triple: \te{<message\_id> po:text '...the text...'}.
+Preprocessed texts are objects of triples: \te{<message\_id> po:cleanText '...te text...'}.
+More specialized predicates are used for text storage when necessary,
+such as \te{po:htmlBodyText} and \te{po:cleanBodyText} used
 for ParticipaBR articles.
@@
 To enable annotations which differ when the snapshot changes,
-\texttt{po:Observation} class instances are used in the triple
-\textttt{<participant\_uri> po:observation <observation\_uri>}.
-The observation instances are then linked to the snapshot and the
+\te{po:Observation} class instances are used in the triple
+\te{<participant\_uri> po:observation <observation\_uri>}.
+Observation instances are then linked to the snapshot and the
 data.
@@
 Instance URIs are built using the URI from the class they derive from plus a hashtag character,
-a provenance string (e.g. \textttt{facebook-legacy} or
+a provenance string (e.g. \te{facebook-legacy} or
 \textttt{participabr-legacy}), and an identifier;
-i.e. \textttt{po:Participant\#<provenance-legacy>-<id>}.
+e.g. \textttt{po:Participant\#<provenance-legacy>-<id>}.
 All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
@@
 
+The authors chose to express all URIs in LOSD locally (i.e. within a local namespace), and linkage with external namespaces may be achieved e.g. by using subclass and ``same as'' relations.
+Also, each snapshot holds specific \te{po:observation} for each participant, allowing for specific metadata such as provided by the original data or as may be provided uppon analysis.
+LOSD does not hold data about each snapshot (such as statistics of the networks), and linkage to external namespaces (such as FOAF and Dublin Core)
+to avoid overcomplicating this initial contribution and unecessary use of storage space.
+
 \section{Usage notes}\label{usage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt} 
-\subsection{Software tools}
-The authors made available software for rendering analyses and
-the data set itself.
-\subsubsection{Triplification routines}
+
+\subsection{Triplification routines}
 For each social platform there is a \emph{triplification} routine,
-i.e. a script for expressinf the data as RDF.
-Original formats and further observations are presented in
+i.e. a script for expressing the data as RDF.
+Original formats and further information are presented in
 Table~\ref{tab:provenance}.
@@
   }
-\subsubsection{Topological and textual analysis}\label{ana}
-Routines are available for taking topological and textual measures from
-the LOSD data.
-  Auxiliary routines, such as for performing principal component analysis (PCA)
-  and obtaining statistics derived from the Kolmogorov-Smirnov test~\cite{kol}, are available to ease pattern recognition.
-
-% \subsubsection{Multimedia rendering}\label{media}
-% It is a core purpose of framework to provide routines for rendering
-% audiovisualizations of the data.
-% Social structures are rendered into music, images and video animations
-% through the Percolation toolbox~\cite{percolation} in association with
-% the Music and Visuals toolboxes~\cite{music,visuals}.
-
-% \subsubsection{Migration from deprecated toolboxes}
-% Routines mentioned in Sections~\ref{ana} and~\ref{media} are being migrated from deprecated
-% toolboxes~\cite{gmaneLegacy,percolationLegacy} into newly designed
-% toolboxes~\cite{percolation,visuals}. TTM, recover?
 
 \subsection{Diagrams of the data and auxiliary tables}\label{sdia}
-LOSD navigation may be assisted through diagrams which describes
+LOSD navigation may be assisted by diagrams which describe
 the structure from each provenance.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[H]
     \centering
     \includegraphics[width=0.5\textwidth]{../ontologies/facebook-legacy-AntonioAnzoategui18022013Friendship.ttl/draw}
-    \caption{A diagram of the core structure that entails the friendship networks
+    \caption{A diagram of the core structure that yields the friendship networks
     of the Facebook snapshots.
@@
 All queries assume the preamble \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
-\begin{enumerate}[leftmargin=0cm]
-	\item Retrieve the number of participants:\\
-            \h\textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
-            ?author a po:Participant . \} }
+\begin{itemize}
+  \item Retrieve the number of participants:
+\begin{lstlisting}[language=spq]
+SELECT (COUNT(DISTINCT ?author) as ?c) WHERE {
+  ?author a po:Participant . 
+}
+\end{lstlisting}
   \item Retrieve the number of relations, be them interactions or
-            friendships:\\
-            \h\textttt{SELECT (COUNT(?interaction) as ?c) WHERE \{\\
-            \h        \h \{ ?interaction a po:Friendship \} UNION \{ ?interaction
-                    a po:Interaction \}\\
-            \h        \h  UNION \{ ?interaction po:retweetOf
-                    ?message \} \\
-            \h\h  UNION \{ ?interaction po:replyTo ?message \}\\
-            \h        \h UNION \{ ?interaction po:directedTo ?participant
-                    \}\\
-            \h\} }
-                  \item Retrieve all text produced by a specific user with URI \textttt{user\_uri}:\\
-            \h\textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{\\
-            \h        \h ?activity po:author <user\_uri> . ?activity po:text ?text .\\
-            \h\}}
+      friendships:
+\begin{lstlisting}[language=spq]
+SELECT (COUNT(?interaction) as ?c) WHERE {
+  { ?interaction a po:Friendship } UNION { ?interaction a po:Interaction }
+  UNION { ?interaction po:retweetOf ?message }
+  UNION { ?interaction po:replyTo ?message }
+  UNION { ?interaction po:directedTo ?participant }
+}
+\end{lstlisting}
+            \item Retrieve all text produced by a specific user with URI \textttt{user\_uri}:
+\begin{lstlisting}[language=spq]
+SELECT (CONCAT(?text) as ?texts) WHERE {
+  ?activity po:author <user_uri> . ?activity po:text ?text .
+}
+\end{lstlisting}
   \item List 1000 users (URIs and names) with the most friendships and the number of
-            friendships in descending order by the number of friendships:\\
-            \h\textttt{SELECT DISTINCT ?participant (COUNT(?friendship) as ?c) WHERE \{\\
-            \h    \h ?friendship a po:Friendship . ?friendship po:member ?participant . \\
-            \h\} ORDER BY DESC(?c) LIMIT 1000}
-        \item Retrieve text messages with the word ``crucial'' (case insensitive):\\
-            \h\textttt{SELECT ?text WHERE \{ \\
-            \h        \h ?activity po:text ?text . FILTER regex(?text, 'crucial', 'i')\\
-            \h\}}
-\item List participants and respective full names whose name has the substring ``Amanda'':\\
-    \h\textttt{SELECT DISTINCT ?participant ?name WHERE \{\\
-    \h\h ?participant po:observation ?obs . ?obs po:name ?name .\\
-    \h\h FILTER regex(?name, 'Amanda', 'i') \\
-    \h\}}
-  \item Return all pairs of friends of a participant (with URI \textttt{participant\_uri}) which are friends themselves:\\
-     \h\textttt{SELECT DISTINCT ?friend1 ?friend2 WHERE \{\\
-     \h\h       ?friendship1 po:member <participant\_uri> . ?friendship1 po:member ?friend1 .\\
-     \h\h       ?friendship2 po:member <participant\_uri> . ?friendship2 po:member ?friend2 .\\
-     \h\h       ?friendship3 po:member ?friend1 . ?friendship3 po:member ?friend2 .\\
-     \h\}}
-\item Return all interactions from replies in a snapshot:\\
-    \h\textttt{SELECT ?from ?to WHERE \{\\
-    \h      \h  ?message1 po:snapshot <snapshot\_uri> .  ?message2 po:replyTo ?message1 .\\
-    \h      \h  ?message1 po:author ?from .  ?message2 po:author ?to .\\
-    \h\}}
-\end{enumerate}
+      friendships in descending order by the number of friendships:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?participant (COUNT(?friendship) as ?c) WHERE {
+  ?friendship a po:Friendship . ?friendship po:member ?participant . 
+} ORDER BY DESC(?c) LIMIT 1000
+\end{lstlisting}
+  \item Retrieve text messages with the word ``crucial'' (case insensitive):
+\begin{lstlisting}[language=spq]
+SELECT ?text WHERE { 
+  ?activity po:text ?text . FILTER regex(?text, 'crucial', 'i')
+}
+\end{lstlisting}
+  \item List participants and respective full names whose name has the substring ``Amanda'':
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?participant ?name WHERE {
+  ?participant po:observation ?obs . ?obs po:name ?name .
+  FILTER regex(?name, 'Amanda', 'i') 
+}
+\end{lstlisting}
+  \item Return all pairs of friends of a participant (with URI \textttt{participant\_uri}) which are friends themselves:
+\begin{lstlisting}[language=spq]
+SELECT DISTINCT ?friend1 ?friend2 WHERE {
+  ?friendship1 po:member <participant_uri> . ?friendship1 po:member ?friend1 .
+  ?friendship2 po:member <participant_uri> . ?friendship2 po:member ?friend2 .
+  ?friendship3 po:member ?friend1 . ?friendship3 po:member ?friend2 .
+}
+\end{lstlisting}
+  \item Return all interactions from replies in a snapshot:
+\begin{lstlisting}[language=spq]
+SELECT ?from ?to WHERE {
+  ?message1 po:snapshot <snapshot_uri> .  ?message2 po:replyTo ?message1 .
+  ?message1 po:author ?from .  ?message2 po:author ?to .
+}
+\end{lstlisting}
+\end{itemize}
 
@@
 and metadata $m$ such as node and link weights, text related to the nodes, etc.
-This section presents the queries for obtaining networks using LOSD
+This section presents queries for obtaining networks using LOSD
 which are promptly envisioned by the authors.
@@
 links, and restrains in the set of nodes and of links,
-are determined by the research being developed.
+are determined by the research or application being developed.
 Furthermore, we do not focus here on multilayer networks (e.g. bipartite networks), or on multigraphs (or muldimensional networks), i.e. on networks with more than one type of node or of link.
@@
   ?f po:member ?a1 . ?f po:member ?a2 .
-
 }
@@
 
-Again, this is not an exaustive list.
+This is not an exaustive list.
 For example, other instances which are related in LOSD, and have authors,
 also yield valid criteria for interaciton networks.
-For a less canonical example for obtaining networks, links may be derived from sufficiently similar vocabulary used by participants, or from activity is sufficiently similar dates, which yield networks which are not interaction or friendship networks.
+For a less canonical example for obtaining networks, links may be derived from sufficiently similar vocabulary used by participants, or from activity with sufficiently similar dates, which yield networks which are not interaction or friendship networks.
+Figure~\ref{fnet} describes most often criteria used to obtain social networks.
 % mk figure with two participants and their relation though response, retweet, or friendship TTM
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{../misc/netSynth____}
+    \caption{Most usual paradigm for considering social networks:
+    actors (i.e. participants, authors) are linked through interactions
+    (e.g. replies, likes, comments)
+    or relations (e.g. friendships, similar activity patterns).}\label{fnet}
+\end{figure}
+
+\subsection{Data access}\label{sac}
+We highly encourage interested parties to take advantage of the API at \url{https://apidocs.data.world/toolkit/api}.
+The user responsible for LOSD has id ``rfabbri'' and the dataset id is ``linked-open-social-data''.
+For making use of the API, and exploiting further services provided by the platform (such as integrations with data analysis software), the researcher needs to register in Data.World, which is fast and free, and facilitates assistance by the platform maintainers.
+Anonynous (not authenticated) SPARQL queries may be performed using the endpoint at: \url{https://rfabbri.linked.data.world/sparql/linked-open-social-data}, kindly provided by the Data.World staff.
+
+\subsection{Example network}\label{sex}
+Figure~\ref{caleb} exemplifies a network obtained through a SPARQL query made to LOSD.
+The nodes are colored and sized according to network metrics, as explained in the caption.
+With \te{http://purl.org/socialparticipation/po/Snapshot\#facebook-legacy-CalebLuporini13042013} as the \te{snaphot\_uri}, the query performed to find the network is:
+\begin{lstlisting}[language=spq]
+PREFIX po: <http://purl.org/socialparticipation/po/>
+SELECT ?a1 ?a2 WHERE {
+  ?f a po:Friendship . ?f po:snapshot <snapshot_uri> .
+  ?f po:member ?a1, ?a2 .
+  FILTER(?a1 != ?a2 && ?a1 > ?a2)
+}
+\end{lstlisting}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../misc/caleb}
+    \caption{
+      A network retrieved from LOSD as described in Section~\ref{sex}.
+  Nodes are colored according to clustering coefficient (greatest values are green, smallest values are pink) and their sizes are proportional to degree (number of links).
+The network has 1032 nodes and 24653 links, average degree of 47.77, average clustering coefficient of 0.47, diameter of 7 (greatest shortest path between nodes), etc.
+    }\label{caleb}
+\end{figure}
 
@@
 \label{conclusions}
-This article outlines LOSD, a large human online social networking dataset (with almost 35 million triples) with diverse provenance.
-Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, are also described to the convenience of the newcommer
+This article outlined LOSD, a large human online social networking dataset, with almost 32 million triples and diverse provenance.
+Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcommer
 and for a consistent exposition.
-The resulting dataset should be expanded upon need or request from interested parties.
-All data should be available online in the \url{http://linkedopensocialdata.org}
-address in near future to fulfill the purpose of being a common
-repertoire in current research.
-All data is publicly available though the endpoint \url{xxxyyyzzz},
-through which the scientific community now has consistent and curated data
+All data is publicly available, and the scientific community now has consistent and curated data
 that enables research validation, continuity, and benchmarking.
-The diagrams and tables of the Supporting Information document of this article
-for further directions
-on the available structures and for an overview complement.
+The diagrams and tables of the Supplementary document of this article
+has further directions on the available structures and is an overview complement.
+
+The authors plan on using LOSD for analyzing simple, bipartite
+and dynamic (i.e. time-evolving) networks.
+The research and development of data visualization interfaces using LOSD
+is envisioned by the authors (FAPESP grant ).
+Directly related to publications, the authors should submite a short paper
+describing and exploring the data collection procedure outlined in Section~\ref{acq} and the data-driven ontology synthesis outlined in Section~\ref{ont}.
+For enhancing LOSD, it should be expanded upon need or request from interested parties.
+Also, software packages for LOSD data retrieval and analysis should be made available.
+Routines for anonymization of LOSD data may be desired in order to preserve privacy of the participants~\cite{ieee}.
+Careful linkage of LOSD classes and entities with third party data (e.g. DBPedia or other Data.World datasets) should make LOSD a five start dataset~\cite{5star}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{The authors thank all participants in the experiments which  donated the data they gathered as described in Section~\ref{}.}
-
+\acknowledgments{
+The authors acknowledge the financial support of the São Paulo State Research Foundation (FAPESP grant 2017/05838-3) and the National Council for Scientific and Technological Development (CNPq, grant 140860/2013-4).
+The authors thank Data.World staff for providing the technical assistance when for providing LOSD,
+for providing the authors the possibility of uploading LOSD which is above their standard size limits,
+and for providing the authors and the scientific community the public query endpoint mentioned in Section~\ref{sac}.
+The views expressed do not reflect the official policy or position of FAPESP, CNPq or Data.World.
+The authors thank all participants in the experiments which resulted in the donation of the data they gathered as described in Section~\ref{acq}.
+}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-% \begin{thebibliography}{999}
-% % Reference 1
-% \bibitem[Author1(year)]{ref-journal}
-% Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% % Reference 2
-% \bibitem[Author2(year)]{ref-book}
-% Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-% \end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-07-01 mdpi/template.tex
acrescentado sobre virtual ethnography

@@
 etnographic tecnique of investigating the social structures
-by studying the researcher, as to ammeliorate the ethic issues.
+by studying the researcher, as to ammeliorate the ethical issues
+(not to be confused e.g. with ``virtual ethnography''~\cite{veth}).
 Accordingly, aside from the data acquired from Twitter, all data in LOSD
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{This research was funded by CNPq (grant number XXX) and FAPESP (grant number ).}
+\funding{This research was funded by CNPq (grant 14068/2013-4) and FAPESP (grant 2017/05838-3).}
 
@@
 %% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{}
-\unskip
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
+% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+% \appendix
+% \section{}
+% \unskip
+% \subsection{}
+% The appendix is an optional section that can contain details and data supplemental to the main text.
 For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
 Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+% 
+% \section{}
+% All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
 

### 2019-07-02 mdpi/template.tex
acrescentado sobre virtual ethnography
finalizada revisao do artigo

@@
 % Full title of the paper (Capitalized)
-\Title{LOSD: A Linked Open Social Dataset for Scientific Benchmarking}
+\Title{LOSD: A Linked Open Social Dataset for Scientific Research}
 
@@
 diverse data, most often not accessible to researchers other than the publishing authors.
-In order to provide the scientific community a common repertoire,
+In order to enhance the scientific community's common repertoire,
 this work presents an open dataset with diverse provenance of 
@@
 Facebook, Twitter, IRC, Email.
-Data from ParticipaBR, AA, and Cidade Democrática are also supplied as samples of less known instances.
+Data from Participa.BR, AA, and Cidade Democrática are also supplied as samples of less known instances.
 The linked data representation was adopted in order to
 comply with current best practices, homogenize access, facilitate discovery
-and analyzes which integrate third party and provided data.
+and analyzes which integrate third party and provided data,
+and make available a dataset with multiple metadata that admits on-demand retrieval of structures.
 Futhermore, the method for obtaing and making the data available is potentially novel,
-and may be of use for other research groups.
+and may be of use for collecting and sharing additional data and for other research groups.
 This document presents a description, usage remarks and overall statistics of the dataset,
@@
 The research on human social and complex networks often
-rely on data that express real phenomena~\cite{c1,c2,c3}.
+relies on data that express real phenomena~\cite{c1,c2,c3}.
 Such data is regularly not made publicly available,
@@
 and does not favor subsequent work.
-Furthermore, there is a lack of open datasets for benchmarking results
+Furthermore, there is a shortcoming of open datasets for benchmarking results
 related to complex and social networks,
 yielding diverse results from poorly related sources.
-Social networks linked datasets are currently most often restricted to a specific platform or express statistics of social networks, as described in Section~\ref{srel}.
+Social networks linked datasets are currently most often restricted to a specific platform
+or express statistics of social networks, as described in Section~\ref{srel}.
 This work presents 
@@
 that has been achieved by linking social data derived from diverse
-provenance, namely: Facebook, Twitter, IRC, Email groups, ParticipaBR, Cidade Democrática, and AA.
+provenance, namely: Facebook, Twitter, IRC, Email groups, Participa.BR, Cidade Democrática, and AA.
 Such data may be of use as a common and public repertoire for scientific
 research involving complex networks, specially if social networks and textual content are pinpointed.
-In fact, the data has already been used in scientific research~\cite{stab,thesis} (CNPq grant 140860/2013-4),
-and should be further used e.g. in researching complex networks and data visualization (FAPESP grant 2017/05838-3)
+In fact, LOSD has already been used in scientific research~\cite{stab,thesis} (CNPq grant 140860/2013-4),
+and should be further exploited e.g. in researching complex networks and data visualization (FAPESP grant 2017/05838-3)
 which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
@@
 This paper is organized as follows. Next subsection briefly discusses related work.
-Section~\ref{materials} describes the data currently available.
-Section~\ref{methods} describes the methods used for data gathering, translation into RDF linked data, and the synthesis of auxiliary structures.
-Section~\ref{usage} holds directions for using the data.
+Section~\ref{materials} describes the data currently available in LOSD.
+Section~\ref{methods} describes the methods used for data gathering,
+translation into RDF linked data, and the synthesis of auxiliary structures.
+Section~\ref{usage} holds directions for LOSD usage.
 
@@
 Most available datasets on human social networks rely only on one source (e.g.~\cite{nat,fb1}),
-on specialized communities such as related to research or movies (e.g.~\cite{s1},
+on specialized communities such as related to research or movies (e.g.~\cite{s1}),
 or present only statistics about social networks, and not the data that yields the networks (e.g.~\cite{st1}.
-Linked human social network data is scarcely found in the scientific community.
-In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud, and made available a dataset from the Advogato platform,
-dedicated to free software developers linked by trust relatioships.
-There are various datasets which makes available data obtained from Twitter (e.g.~\cite{tw1,tw2}),
-most often with a specific focus (users, retweets, hashtags, etc)
+In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud,
+and made available a dataset from the Advogato platform,
+dedicated to free software developers connected by trust relatioships.
+There are various datasets which provide data obtained from Twitter (e.g.~\cite{tw1,tw2}),
+most often with a specific focus (users, retweets, hashtags, user mentions, etc)
 and not presented to the scientific community by means of a research article.
-The datasets which are most closely related to LOSD are KONECT~\cite{konect},
+The datasets which are most closely related to LOSD are: KONECT~\cite{konect},
 SNAP Datasets~\cite{snapnets}, the Network Repository~\cite{nr}, and ICON~\cite{icon}.
 None of them are (RDF) linked data but express the networks in formats specialized for networks.
-In total, these datasets make available thousands of social networks, from diverse provenance and of various types (e.g. simple, directed, weighted, bipartite, temporal networks).
-Compared to LOSD, KONECT, SNAP, Network Repository, and ICON are complementary, providing a further diversity of networks, but only as simplified data, i.e.
-in specialized formats, most often expressed as columns of node ids that are connected and metadata (such as timestaps for temporal networks).
+In total, these datasets make available thousands of networks,
+a fraction of them are social networks from diverse provenance and
+of various types (e.g. simple, directed, weighted, bipartite, temporal networks).
+Compared to LOSD, KONECT, SNAP, Network Repository, and ICON are complementary, providing a further diversity of networks,
+but only as simplified data, i.e. in specialized formats, most often expressed as columns of node ids
+that are connected and metadata (such as timestaps for temporal networks).
 In summary, no linked human social network dataset presented to the
@@
 % scripts/percolationUse2.py
-The database consists of 31,996,740 triples, 3,030,433 network links yield by main interactions or relations, 350,284 participants and 245,194,377 characters.
-Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction edges, 89 have friendship edges; 43 have text content from messages.
+The LOSD dataset consists of 31,996,740 triples, 350,284 participants,
+3,030,433 network links yield by most important interactions or relations, and 245,194,377 characters.
+Among all snapshots, 63 are ego snapshots, 54 are group snapshots; 50 have interaction links,
+89 have friendship links; 43 have text content associated to participant activity.
 Section~\ref{queries} holds the queries used to obtain such data.
-This section describes de data in LOSD in very brief terms.
+This section describes the data in LOSD in very brief terms.
 The Supplementary document is provided for further information about the dataset.
@@
     \item Netvizz software~\cite{netvizz} and subsequent donation by users (Facebook);
-    \item donation by system administrators (AA, ParticipaBR, Cidade Democr\'atica).
+    \item donation by system administrators (AA, Participa.BR, Cidade Democr\'atica).
 \end{itemize}
@@
 group, prospected at the same time, is a snapshot.
-Table~\ref{tsnap} holds the number of snapshots from each provenance in LOSD.
+Table~\ref{tsnap} denotes the number of snapshots from each provenance in LOSD.
 
@@
 Friendship ego networks (networks whose nodes represent friends of a user and links represent friendships)
-were donated from individual users in 2013 and 2014.
+were donated from individual users in 2013-5.
 Friendship and interaction networks from groups were gathered from
@@
 but replies and user mentions are also kept in the database.
-LOSD kept the main data fields returned by Twitter API, such as creation date, text, and hashtags.
+LOSD kept the main data fields returned by Twitter API,
+such as creation date, text, and hashtags.
 
@@
 
-\subsection{ParticipaBR data}
-ParticipaBR is a Brazilian federal platform for social participation,
-the most important platform from the myriad of social participation platforms
-that were available in 2011-2018~\cite{spbr}.
+\subsection{Participa.BR data}
+Participa.BR is a Brazilian federal platform for social participation,
+one of the most important platforms from the myriad of social participation platforms
+that were available in Brazil in 2011-2018~\cite{spbr}.
 Texts are derived from blog posts and networks may be obtained from
@@
 Cidade Democrática is a Brazilian civil society social participation portal.
-Data gathered is complex in the number of types of instances as noticeable in the Supplementary document.
+Data gathered is complex in the number of types of instances
+as noticeable in the Supplementary document.
 
 \subsection{AA data}
-AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications by sharing sentences about ongoing work and testifying about their quality.
+AA (Algorithmic Autoregulation~\cite{aa,aa2}) is a software to keep track of dedications by
+sharing sentences about ongoing work and testifying about their quality.
 The data was gathered from different versions of the system and from an IRC
@@
 \section{Methods}\label{methods}
-LOSD is expressed as linked data through RDF
+LOSD is expressed as RDF linked data
 and is ontologicaly described through the
@@
 \subsection{Data acquisition}\label{acq}
-Research involving human subjects raise well-known ethical issues,
-although ethical regulation for (non-medical) research in social systems is not always endorsed by scientists~\cite{eth}.
+Research involving human subjects raises well-known ethical issues,
+although ethical regulation for (non-medical) research in social systems
+is not always endorsed by scientists~\cite{eth}.
 Furthermore, the online social networking platforms often restrict access
-to the data and user may consider illegal the use of their data through web crowlers~\cite{ile1,ile2,ile3}.
-Although Twitter and GMANE (email) data is available through public APIs,
+to the data and users may consider illegal the use of their data through web crowlers~\cite{ile1,ile2,ile3}.
+Although Twitter and GMANE (email) data are available through public APIs,
 the other sources are more problematic.
@@
 collaborators in the fields of phylosophy, social sciences, antropology and psychology.
-The conceptual framework is presented in~\cite{antphy,antphy2}, and in essence
+The conceptual framework is presented in~\cite{antphy,antphy2,antphy3}, and in essence
 consists in researching first the social networks related to the researchers,
@@
 and was called ``Anthropological Physics'' because it relies in the
-etnographic tecnique of investigating the social structures
+etnographic tecnique of investigating social structures
 by studying the researcher, as to ammeliorate the ethical issues
@@
 Accordingly, aside from the data acquired from Twitter, all data in LOSD
-was gathered from social instances where the first author is a participant.
+was gathered from social instances where the first author was a participant.
 For example, all Facebook ego networks were donated to the first author by
-users which know him and were interested in his research and intervention on the networks,
-and all email and Facebook groups data was obtained from groups where the first author is a member.
+users which knew him and were interested in his research and intervention on the networks,
+and all email and Facebook groups data was obtained from groups where the first author was a member.
 
 Facebook data was obtained through the Netvizz software~\cite{netviz},
-Facebook allowed downloading friendships and interaction data as available in LOSD.
+when Facebook allowed downloading friendships and interaction data as available in LOSD.
 Data from Twitter was obtained through the Twitter streamming API.
@@
 Email data was obtained through the GMANE public API.
-Data from ParticipaBR, Cidade Democrática, and AA was obtained through contact
+Data from Participa.BR, Cidade Democrática, and AA was obtained through contact
 of the researcher with the developers of the platforms.
+These sources are expreesed in Table~\ref{tab:provenance}.
 
@@
 The yielded web of data is composed of documents on the web
-such as the web of HTML documents.
+in similarity to the web of HTML documents.
 In practice, the publication of linked data can me summarized
@@
 \begin{itemize}
-    \item Use URIs to identify things~\cite{uri}.
+  \item Use URIs (Uniform Resource Identifiers) to identify things~\cite{uri}.
     \item Use HTTP URIs.
@@
 Turtle and Manchester, which all, in essence, represent a labeled and
-directed multi-graph.
+directed multigraph.
 RDF may be stored in a type of database called a triplestore.~\cite{rdf}
@@
 format asserts that ``the paper has color white'':\\
-\texttt{http://example.org/Thing\#Paper http://example.org/hasColor\\
+\te{http://example.org/Thing\#Paper http://example.org/hasColor\\
 http://example.org/Color\#White .}
@@
   \item properties, which are concepts that are used as predicates and thus relate concepts;
-  \item special relations between concepts that state that one concept is more general than the other\footnote{This
-          kind of relation is called hypernymy. Examples: mammal is a hypernym of monkey, drink is a hypernym of beer.};
+  \item special relations between concepts that state that one concept (superclass) is more general than the other (subclass)\footnote{This
+          kind of relation is called a hypernymy. Examples: mammal is a hypernym of monkey, drink is a hypernym of beer.};
   \item the subjects and objects that can occurs in a triple where a specific property is a predicate.
@@
 
-One can also write an ontology using the OWL protocol,
+One may also write an ontology using the OWL protocol,
 with which all the expressive capabilities of RDFS are
@@
 The advantage is the greater power to represent conceptualizations.
-Using ontologies enables automated reasoning.
-For example, if a property \textttt{:r} is known to relate only mokeys to trees,
-then, if there is a triple \textttt{:a :r :b}, the instance \textttt{:a} is considered a monkey,
-and instance \textttt{:b} is considered a tree,
+Using ontologies enables automated reasoning (usually called ``inference'').
+A very simple example: if a property \te{:r} is known to relate only mokeys to trees,
+then, if there is a triple \te{:a :r :b}, the instance \te{:a} is considered a monkey,
+and instance \te{:b} is considered a tree,
 even if there is no explicit declaration similar
-to \textttt{:a rdf:type :Mokey}.
+to \te{:a rdf:type :Mokey}.
 
@@
 \begin{itemize}
-  \item Using RDFS, one may define a property domain, i.e. the class instances that are allowed as the object in a triple with the property.
-  \item Using RDFS, one may define a property range, i.e. the class instances that are allowed as the predicate in a triple with the property.
-  \item Using OWL, one may define an existential restriction for a class \textttt{:C},
-    i.e. declare that an instance \textttt{:C\#I} of such class has at least one relation with a property \textttt{:p} and a class \textttt{:C2} in the form
-    \textttt{:C\#I :p :C2\#I}, where \textttt{:C2\#I} is an instance of the class \textttt{:C2}.
-  \item Using OWL, one may define an universal restriction for a class \textttt{:C},
-    i.e. declare that any relation of an instance \textttt{:C\#I} of such class, with a property \textttt{:p} of the form \textttt{:C\#I :p :CX\#I},
-    implies that \textttt{:CX\#I} is an instance of a specific class, e.g. \textttt{:C2}.
+  \item Using RDFS, one may define a property domain, i.e. the class instances that are allowed as the subject in a triple with the property.
+  \item Using RDFS, one may define a property range, i.e. the class instances that are allowed as the object in a triple with the property.
+  \item Using OWL, one may define an existential restriction for a class \te{:C} using a property \te{:p} and another class \te{:C2},
+    i.e. declare that $\forall$ \te{:C\#I}, if \te{:C\#I rdf:type :C} then $\exists$ \te{:C2\#I rdf:type :C2} such that \te{:C\#I :p :C2\#I}.
+  \item Using OWL, one may define a universal restriction for a class \te{:C} using a property \te{:p} and another class \te{:C2},
+    i.e. declare that $\forall$ \te{:C\#I}, if \te{:C\#I rdf:type :C} then
+    \te{:C\#I :p :C2\#I} implies that \te{:C2\#I rdf:type :C2}.
 \end{itemize}
@@
 Most ontologies are created by domain experts~\cite{ont}
-even though the data they arrange is often given by a software system
+even though the data they arrange is often output by a software system
 and has thus a predefined structure.  
@@
         the predicate is the property (\te{property\_uri}) and the object is
-        an instance of the same class (\te{class2\_uri}).
+        an instance of the same class (\te{object\_curi}).
 	The queries are:
@@
 \end{lstlisting}
-        Then, for each object class (\textttt{object\_curi}) in \textttt{?oc}:
+        Then, for each object class (\te{object\_curi}) in \te{?oc}:
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item Render diagrams to achieve figures as exemplified in Section~\ref{sdia} and in the Supporting Information document.
+  \item Render diagrams to achieve figures as exemplified in Section~\ref{sdia} and provided in the Supplementary document.
 \end{enumerate}
@@
 LOSD data is represented through URIs and triples by translation of various formats to RDF.
-URIs are built in the namespace \url{http://purl.org/socialparticipation/participationontology/}
+URIs are built in the namespace \te{http://purl.org/socialparticipation/participationontology/}
 which are identified herein with the prefix ``\te{po:}''.
@@
 Message texts, including comments, are objects in a triple: \te{<message\_id> po:text '...the text...'}.
-Preprocessed texts are objects of triples: \te{<message\_id> po:cleanText '...te text...'}.
+Preprocessed texts are objects of triples: \te{<message\_id> po:cleanText '...the text...'}.
 More specialized predicates are used for text storage when necessary,
 such as \te{po:htmlBodyText} and \te{po:cleanBodyText} used
-for ParticipaBR articles.
+for Participa.BR articles.
 A participant URI is unique throughout the provenance (e.g. the same for
@@
 a provenance string (e.g. \te{facebook-legacy} or
-\textttt{participabr-legacy}), and an identifier;
-e.g. \textttt{po:Participant\#<provenance-legacy>-<id>}.
-All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
-All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
-\textttt{irc-legacy-labmacambira} or
-\textttt{email-legacy-linux.audio.devel1-20000}.
+\te{participabr-legacy}), and an identifier;
+e.g. \te{po:Participant\#<provenance-legacy>-<id>}.
+All snapshot URIs follow the formation rule: \te{po:<SnapshotProvenance>\#<snapshot\_id>}.
+All snapshot ids follow the formation rule: \te{<platform>-legacy-<further\_identifier>}; e.g.
+\te{irc-legacy-labmacambira} or
+\te{email-legacy-linux.audio.devel1-20000}.
 
 The authors chose to express all URIs in LOSD locally (i.e. within a local namespace), and linkage with external namespaces may be achieved e.g. by using subclass and ``same as'' relations.
-Also, each snapshot holds specific \te{po:observation} for each participant, allowing for specific metadata such as provided by the original data or as may be provided uppon analysis.
-LOSD does not hold data about each snapshot (such as statistics of the networks), and linkage to external namespaces (such as FOAF and Dublin Core)
+Also, each snapshot holds a specific \te{po:Observation} instance related to each participant,
+allowing for specific metadata such as provided by the original data or as may be provided uppon analysis.
+Current LOSD does not hold statistics about each snapshot, or linkage to external namespaces (such as FOAF and Dublin Core),
 to avoid overcomplicating this initial contribution and unecessary use of storage space.
@@
 \section{Usage notes}\label{usage}
-
 \subsection{Triplification routines}
@@
 \begin{tabular}{ l | p{3cm} p{3cm} c }
-    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
+    \textbf{social platform} & \textbf{original format} & \textbf{further note} & \textbf{toolbox} \\\specialrule{1.5pt}{1pt}{1pt}
     AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation~\cite{participation} \\\hline
@@
     IRC & plain text log & obtained through Supybot logging & Social \\\hline
-    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
+    Participa.BR & PostgreSQL database & donated by admins & Participation \\\hline
     Twitter & JSON & obtained through Twitter streaming API & Social \\
@@
 the structure from each provenance.
-Such diagrams are in the Supporting Information document
+Such diagrams are in the Supplementary document
 together with tables to facilitate the appreciation of the provided data.
@@
     \caption{A diagram of the core structure that yields the friendship networks
-    of the Facebook snapshots.
-A green edge denotes an OWL existential class restriction; an inverted nip denotes an OWL universal class restriction; a full (non-dashed) edge denotes an OWL functional property axiom (see Section~\ref{sont}).
-Further information and complete diagrams for each provenance are available in the Supporting Information document.}\label{dia}
+    of the Facebook snapshots. A green edge denotes an OWL existential class restriction; an inverted arrow tip denotes an OWL universal class restriction;
+    a full (non-dashed) edge denotes an OWL functional property axiom (see Sections~\ref{sont} and~\ref{ont}).
+Further information and complete diagrams for each provenance are available in the Supplementary document.}\label{dia}
 \end{figure}
@@
 There are numerous useful and general purpose SPARQL queries to be performed against LOSD.
-In this section, a selection of some of the most basic queries are made available due to their potential to be varied.
-All queries assume the preamble \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
+In this section, a selection of some of the most basic queries are provided due to their potential to be varied.
+All queries assume the preamble \te{PREFIX po: <http://purl.org/socialparticipation/po/>}.
 \begin{itemize}
@@
 \end{lstlisting}
-  \item Retrieve the number of relations, be them interactions or
+  \item Retrieve the number of most relevant relations, be them interactions or
       friendships:
@@
 \end{lstlisting}
-            \item Retrieve all text produced by a specific user with URI \textttt{user\_uri}:
+            \item Retrieve all text produced by a specific user with URI \te{user\_uri}:
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item List participants and respective full names whose name has the substring ``Amanda'':
+  \item List participants and respective snapshots:
 \begin{lstlisting}[language=spq]
-SELECT DISTINCT ?participant ?name WHERE {
-  ?participant po:observation ?obs . ?obs po:name ?name .
-  FILTER regex(?name, 'Amanda', 'i') 
+SELECT DISTINCT ?participant ?snap WHERE {
+  ?participant po:observation ?obs . ?obs po:snapshot ?snap .
 }
 \end{lstlisting}
-  \item Return all pairs of friends of a participant (with URI \textttt{participant\_uri}) which are friends themselves:
+  \item Return all pairs of friends of a participant (with URI \te{participant\_uri}) which are friends themselves:
 \begin{lstlisting}[language=spq]
@@
   ?friendship3 po:member ?friend1 . ?friendship3 po:member ?friend2 .
+  FILTER(
+    str(?friend1) != 'participant_uri' && str(?friend2) != 'participant_uri'
+    && str(?friend1) != str(?friend2)
+  )
 }
 \end{lstlisting}
-  \item Return all interactions from replies in a snapshot:
+  \item Return all interactions from replies in a snapshot (with URI \te{snapshot\_uri}):
 \begin{lstlisting}[language=spq]
@@
 A complex network $N=(n,l,m)$ may be regarded essentially as
-a set of nodes $n$, a set of links $l$ that related the nodes,
+a set of nodes $n$, a set of links $l$ that relate the nodes,
 and metadata $m$ such as node and link weights, text related to the nodes, etc.
@@
 Many other networks may be obtained because the criteria that entail
-links, and restrains in the set of nodes and of links,
+links and nodes, and restrictions to the set of nodes and of links,
 are determined by the research or application being developed.
-Furthermore, we do not focus here on multilayer networks (e.g. bipartite networks), or on multigraphs (or muldimensional networks), i.e. on networks with more than one type of node or of link.
-Here is a list with short descriptions of the networks and the corresponding SPARQL queries:
+Furthermore, we do not focus here on multilayer networks (e.g. bipartite networks),
+or on multigraphs (or muldimensional networks),
+i.e. on networks with more than one type of node or of link.
+Here is a list with short descriptions of the resulting networks and the corresponding SPARQL queries:
 \begin{itemize}
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item Overall facebook friendship network derived only from ego networks:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item Overall facebook interaction network available in LOSD:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
-  \item Email interaction network for email list shapshot with URI \verb|http://xxxx|, with the all (preliminarily cleaned) text related to each message (node atribute):
+\end{lstlisting}
+  \item Email interaction network for email list shapshot with URI \verb|snapshot_uri|,
+    with the all (preliminarily cleaned) text related to each message (node atribute):
 \begin{lstlisting}[language=spq]
 SELECT ?a1 ?a2 ?t1 ?t2 WHERE { 
-  ?m po:snapshot <http://xxxx> . ?m a po:EmailMessage .
-  ?m po:author ?a1 . ?m po:replyTo ?m2 . ?m2 po:author ?a2 .
+  ?m po:snapshot <snapshot_uri> . ?m po:author ?a1 .
+  ?m po:replyTo ?m2 . ?m2 po:author ?a2 .
   ?m po:cleanText ?t1 . ?m2 po:cleanText ?t2 .
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item IRC users related by messages directed to each other, and the text of each message (link attribute):
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item Twitter users related by retweets:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item Twitter users related by user mentions and the text of the tweets:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
-  \item Friendship network from ParticipaBR, with the date when each frienship was established:
+\end{lstlisting}
+  \item Friendship network from Participa.BR, with the date when each frienship was established:
 \begin{lstlisting}[language=spq]
@@
   ?s po:shapshotID 'participabr-legacy' . ?f po:snapshot ?s .
-  ?f a po:Friendship . 
-  ?f po:member ?a1 . ?f po:member ?a2 . ?f po:createdAt ?d .
+  ?f a po:Friendship .  ?f po:member ?a1 . ?f po:member ?a2 .
+  ?f po:createdAt ?d .
 }
-\end{lstlisting}\vspace{-0.2cm}
-  \item Interaction network from ParticipaBR obtained by comments on articles:
+\end{lstlisting}
+  \item Interaction network from Participa.BR obtained by comments on articles:
 \begin{lstlisting}[language=spq]
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
-  \item Interaction network from ParticipaBR obtained by comments and votes on comments:
+\end{lstlisting}
+  \item Interaction network from Participa.BR obtained by comments and votes on comments:
 \begin{lstlisting}[language=spq]
@@
   ?c a po:Comment .  ?c po:author ?a1 .
-  { ?v a po:Vote . ?v po:reference ?c . ?v po:author ?a2 .
+  {
+    ?v a po:Vote . ?v po:reference ?c . ?v po:author ?a2 .
   } UNION {
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item Interaction network from Cidade Democrática obtained by comments on Topics:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
   \item AA users, related by AA session validations:
@@
 }
-\end{lstlisting}\vspace{-0.2cm}
+\end{lstlisting}
 \end{itemize}
@@
 This is not an exaustive list.
-For example, other instances which are related in LOSD, and have authors,
-also yield valid criteria for interaciton networks.
-For a less canonical example for obtaining networks, links may be derived from sufficiently similar vocabulary used by participants, or from activity with sufficiently similar dates, which yield networks which are not interaction or friendship networks.
+For a less canonical example for obtaining networks,
+links may be derived from sufficiently similar vocabulary used by participants,
+or from activity with sufficiently similar dates,
+which yield networks which are not (standard) interaction or friendship networks.
 Figure~\ref{fnet} describes most often criteria used to obtain social networks.
@@
     \includegraphics[width=0.5\textwidth]{../misc/netSynth____}
-    \caption{Most usual paradigm for considering social networks:
+    \caption{Most usual paradigm for the achievement of networks from data related to human social structures:
     actors (i.e. participants, authors) are linked through interactions
@@
 \subsection{Data access}\label{sac}
-We highly encourage interested parties to take advantage of the API at \url{https://apidocs.data.world/toolkit/api}.
+The authors highly encourage interested parties to take advantage of the API at \url{https://apidocs.data.world/toolkit/api}.
 The user responsible for LOSD has id ``rfabbri'' and the dataset id is ``linked-open-social-data''.
 For making use of the API, and exploiting further services provided by the platform (such as integrations with data analysis software), the researcher needs to register in Data.World, which is fast and free, and facilitates assistance by the platform maintainers.
-Anonynous (not authenticated) SPARQL queries may be performed using the endpoint at: \url{https://rfabbri.linked.data.world/sparql/linked-open-social-data}, kindly provided by the Data.World staff.
+Anonymous (not authenticated) SPARQL queries may be performed using the endpoint at: \url{https://rfabbri.linked.data.world/sparql/linked-open-social-data}, kindly provided by the Data.World staff.
+Also, a very simple Python package for using LOSD has been released by the authors\footnote{Instructions for installation and usage at: \url{https://pypi.org/project/losd/}.}.
+This package contains functions to retrieve data and render networks from the data.
 
@@
   ?f po:member ?a1, ?a2 .
-  FILTER(?a1 != ?a2 && ?a1 > ?a2)
+  FILTER(?a1 != ?a2)
 }
@@
 \label{conclusions}
-This article outlined LOSD, a large human online social networking dataset, with almost 32 million triples and diverse provenance.
+This article outlined LOSD, a large dataset of human and complex networks, with almost 32 million triples and diverse provenance.
 Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcommer
 and for a consistent exposition.
-All data is publicly available, and the scientific community now has consistent and curated data
+All data is publicly available to provide the scientific community with a consistent and curated linked data dataset
 that enables research validation, continuity, and benchmarking.
 The diagrams and tables of the Supplementary document of this article
-has further directions on the available structures and is an overview complement.
+has further directions on the available structures and is a complementary overview.
 
@@
 The research and development of data visualization interfaces using LOSD
-is envisioned by the authors (FAPESP grant ).
-Directly related to publications, the authors should submite a short paper
-describing and exploring the data collection procedure outlined in Section~\ref{acq} and the data-driven ontology synthesis outlined in Section~\ref{ont}.
+is envisioned by the authors (FAPESP grant 2017/05838-3).
+Directly related to publications, the authors should submit a short paper
+describing and further exploring the data collection procedure outlined in Section~\ref{acq},
+and the data-driven ontology synthesis outlined in Section~\ref{ont}.
 For enhancing LOSD, it should be expanded upon need or request from interested parties.
-Also, software packages for LOSD data retrieval and analysis should be made available.
+Careful linkage of LOSD classes and entities with third party data (e.g. DBPedia or other Data.World datasets) should make LOSD a five start dataset~\cite{5star}.
 Routines for anonymization of LOSD data may be desired in order to preserve privacy of the participants~\cite{ieee}.
-Careful linkage of LOSD classes and entities with third party data (e.g. DBPedia or other Data.World datasets) should make LOSD a five start dataset~\cite{5star}.
+Finally, software packages for providing standard networks and data analysis using LOSD should be made available.
 
@@
 %% optional
-\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
+\supplementary{The Supplementary document, available online at \linksupplementary{s1}, provides diagrams of LOSD data, further descriptions and elementary statistics.}
 
@@
 for providing the authors the possibility of uploading LOSD which is above their standard size limits,
-and for providing the authors and the scientific community the public query endpoint mentioned in Section~\ref{sac}.
+and for providing the authors and the scientific community the public query endpoint mentioned in Section~\ref{sac},
+specially Patrick McGarry (Community Head) and Bryon Jacob (CTO).
 The views expressed do not reflect the official policy or position of FAPESP, CNPq or Data.World.

### 2019-07-02 mdpi/template.tex
revisado supplementary document e passado spellcheck

@@
 and make available a dataset with multiple metadata that admits on-demand retrieval of structures.
-Futhermore, the method for obtaing and making the data available is potentially novel,
+Furthermore, the method for obtaining and making the data available is potentially novel,
 and may be of use for collecting and sharing additional data and for other research groups.
@@
 at the core of data collection, and keeps the resources (e.g. code, data, documents) open,
-in order to ammeliorate ethical issues involved in studying human social systems.
+in order to ameliorate ethical issues involved in studying human social systems.
 
@@
 and made available a dataset from the Advogato platform,
-dedicated to free software developers connected by trust relatioships.
+dedicated to free software developers connected by trust relationships.
 There are various datasets which provide data obtained from Twitter (e.g.~\cite{tw1,tw2}),
@@
 but only as simplified data, i.e. in specialized formats, most often expressed as columns of node ids
-that are connected and metadata (such as timestaps for temporal networks).
+that are connected and metadata (such as timestamps for temporal networks).
 In summary, no linked human social network dataset presented to the
-scientific community containts the diverse provenance used in LOSD,
+scientific community contains the diverse provenance used in LOSD,
 or provides such a wide range of metadata,
@@
 comprises a snapshot; the tweets from the MAMA music event is a
-snapshot; the friendship, interaction and posts structures of a facebook
+snapshot; the friendship, interaction and posts structures of a Facebook
 group, prospected at the same time, is a snapshot.
@@
 LOSD is expressed as RDF linked data
-and is ontologicaly described through the
+and is ontologically described through the
 data-driven ontology synthesis method presented in Section~\ref{ont}.
@@
 if not prohibited.
-We reached a potential solution to this issue through continuated research with
-collaborators in the fields of phylosophy, social sciences, antropology and psychology.
+We reached a potential solution to this issue through continued research with
+collaborators in the fields of philosophy, social sciences, anthropology and psychology.
 The conceptual framework is presented in~\cite{antphy,antphy2,antphy3}, and in essence
@@
 and was called ``Anthropological Physics'' because it relies in the
-etnographic tecnique of investigating social structures
-by studying the researcher, as to ammeliorate the ethical issues
+ethnographic technique of investigating social structures
+by studying the researcher, as to ameliorate the ethical issues
 (not to be confused e.g. with ``virtual ethnography''~\cite{veth}).
@@
 when Facebook allowed downloading friendships and interaction data as available in LOSD.
-Data from Twitter was obtained through the Twitter streamming API.
+Data from Twitter was obtained through the Twitter streaming API.
 Data From IRC was obtained through bots that kept logs of the messages.
@@
 of the researcher with the developers of the platforms.
-These sources are expreesed in Table~\ref{tab:provenance}.
+These sources are expressed in Table~\ref{tab:provenance}.
 
@@
 Using ontologies enables automated reasoning (usually called ``inference'').
-A very simple example: if a property \te{:r} is known to relate only mokeys to trees,
+A very simple example: if a property \te{:r} is known to relate only monkeys to trees,
 then, if there is a triple \te{:a :r :b}, the instance \te{:a} is considered a monkey,
@@
 even if there is no explicit declaration similar
-to \te{:a rdf:type :Mokey}.
+to \te{:a rdf:type :Monkey}.
 
@@
 Also, each snapshot holds a specific \te{po:Observation} instance related to each participant,
-allowing for specific metadata such as provided by the original data or as may be provided uppon analysis.
+allowing for specific metadata such as provided by the original data or as may be provided upon analysis.
 Current LOSD does not hold statistics about each snapshot, or linkage to external namespaces (such as FOAF and Dublin Core),
-to avoid overcomplicating this initial contribution and unecessary use of storage space.
+to avoid overcomplicating this initial contribution and unnecessary use of storage space.
 
@@
 Furthermore, we do not focus here on multilayer networks (e.g. bipartite networks),
-or on multigraphs (or muldimensional networks),
+or on multigraphs (or multidimensional networks),
 i.e. on networks with more than one type of node or of link.
@@
 \begin{itemize}
-  \item Overall facebook friendship network available in LOSD:
+  \item Overall Facebook friendship network available in LOSD:
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item Overall facebook friendship network derived only from ego networks:
+  \item Overall Facebook friendship network derived only from ego networks:
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item Overall facebook interaction network available in LOSD:
+  \item Overall Facebook interaction network available in LOSD:
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item Email interaction network for email list shapshot with URI \verb|snapshot_uri|,
-    with the all (preliminarily cleaned) text related to each message (node atribute):
+  \item Email interaction network for email list snapshot with URI \verb|snapshot_uri|,
+    with the all (preliminarily cleaned) text related to each message (node attribute):
 \begin{lstlisting}[language=spq]
@@
 \end{lstlisting}
-  \item Friendship network from Participa.BR, with the date when each frienship was established:
+  \item Friendship network from Participa.BR, with the date when each friendship was established:
 \begin{lstlisting}[language=spq]
@@
 
-This is not an exaustive list.
+This is not an exhaustive list.
 For a less canonical example for obtaining networks,
@@
 Figure~\ref{fnet} describes most often criteria used to obtain social networks.
-% mk figure with two participants and their relation though response, retweet, or friendship TTM
 
@@
 This article outlined LOSD, a large dataset of human and complex networks, with almost 32 million triples and diverse provenance.
-Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcommer
+Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcomer
 and for a consistent exposition.

### 2019-07-02 supplementaryDocument.tex
revisado supplementary document e passado spellcheck

--- 2016-08-01 supportingInformation.tex
@@
 %
-\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
+\newcommand{\te}[1] {\texttt{\footnotesize#1}}
 \newcommand{\h} {\hphantom ~ }
-% \newcommand{\textttt}[1] {\mbox{\texttt{\footnotesize#1}}}
-% \newcommand{\textttt}[1] {
+% \newcommand{\te}[1] {\mbox{\texttt{\footnotesize#1}}}
+% \newcommand{\te}[1] {
 % \begin{verbatim} #1 \end{verbatim}
@@
 %
-\title{Linked Open Social Data for Scientific Benchmarking (Supporting
-Information document)}
+\title{LOSD: A Linked Open Social Dataset for Scientific Research
+  (Supplementary document)}
 %
-\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
+\author[pwa]{Renato Fabbri\corref{corresponding}\fnref{kii-url}}
 \ead{fabbri@usp.br}
@@
 \cortext[corresponding]{Corresponding author}
-\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
-University, Brazil}
+\address[pwa]{
+Institute of Mathematical and Computer Sciences, University of São Paulo (ICMC/USP), São Carlos, Brazil
+}
+\address[pwr]{
+São Carlos Institute of Physics, University of São Paulo (IFSC/USP), São Carlos, Brazil
+}
 %
+\fntext[kii-url]{\textit{URL:} \url{https://icmc.usp.br/}}
 \fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
@@
 \begin{abstract}
-This is a Supporting Information document which exposes ontological
-diagrams and auxiliary tables for the Linked Open Social Database.
+This is a Supplementary document which provides ontological
+diagrams and auxiliary tables for the LOSD dataset.
 The main document of the article is in~\cite{losd}.
@@
 %
-\begin{keyword}
-Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email,
-Complex Networks, Text Mining
-%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
-\end{keyword}
-
 \end{frontmatter}
@@
 In this document we provide diagrams
-for the social protocols in the Linked Open Social Database:
-Facebook, Twitter, IRC, Email, ParticipaBR, Cidade Democrática and AA.
-Each social protocol diagram was broken in two, one presents the relations
-among main classes (blue nodes) and data types (orange nodes),
-the other presents metadata for the
-snapshots.
+for the provenances in LOSD, namely
+Facebook, Twitter, IRC, Email, Participa.BR, Cidade Democrática and AA.
+Each provenance has two diagrams: one presents the relations
+among classes (blue nodes) and data types (orange nodes),
+the other presents metadata for the snapshot provenance.
 Every class instance is related to the snapshot instance
-by the triple \textttt{class\_uri po:snapshot snapshot\_uri}.
+by the triple \te{class\_uri po:snapshot snapshot\_uri}.
 Such triples are omitted for simplicity.
 Due to the large number of relations, the rendering of diagrams are
-automatized and displays some overlaps.
-Even so, the images are useful for grasping what is in the current
-database and for conducting explorations.
-Edges in the diagrams have:
+automatized and displays some overlaps of the nodes related to each class or datatype.
+Even so, the images are useful for grasping what is available in current LOSD
+and exploration assistance.
+Edges in the diagrams have (see OWL official documentation or the article~\cite{losd}:
 \begin{itemize}
     \item green color if representing an OWL existential
-class restriction (all individuals from the class present at least one triple with
-the property as predicate);
-    \item inverted nip if representing an OWL universal class
-        restriction (all individuals presenting triples with the
-        property as predicate are from the class);
+class restriction;
+    \item inverted arrow tip if representing an OWL universal class
+        restriction;
     \item full edges (non-dashed) if representing a functional property
-        axiom (there is at most one triple with the property as the
-        predicate for each individual).
+        axiom.
 \end{itemize}
 
-% \textheight = 100pt
-% \pdfpageheight 300pt
 Furthermore, this document ends with two sets of tables, one with
-references for snapshot groups, such as wikipedia or
+references for snapshot groups, such as Wikipedia or
 contact links, 
@@
 \section{Facebook data}
-Each Facebook snapshot is yield by either an user, from which the
-friends constitute a friendship network, or a group, which participants
+Each Facebook snapshot is yield by either a user, from which the
+friends are in the friendship network, or a group, in which participants
 can yield friendship and interaction networks and posts information with
 text and some metadata.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 
@@
 Each Twitter snapshot is yield by a hashtag.
-Retweets (\textttt{po:retweetOf} are usually considered the interactions between users.
-The database present also \textttt{po:replyTo} and \textttt{po:userMention}
-which might be useful in understanding the networking.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+Retweets (\te{po:retweetOf} are usually considered the interactions between users.
+The database present also \te{po:replyTo} and \te{po:userMention}
+which may be useful in understanding the networking and provides additional networks.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 
@@
 Each IRC snapshot is yield by an IRC channel.
-IRC messages are either server messages (e.g. join and exit)
-marked with \textttt{po:systemMessage true} and having an \textttt{po:impliedUser user\_uri},
-or user messages, which yield interactions through \textttt{po:directedTo} and \textttt{po:mentions} properties.
-Text messages without the user names are delivered through the \textttt{po:cleanText} property.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+IRC messages are either 1) server messages (e.g. join and exit)
+marked with \te{po:systemMessage true} and having an \te{po:impliedUser user\_uri},
+or 2) user messages, which yield interactions through \te{po:directedTo} and \te{po:mentions} properties.
+Text messages without the user names are provided through the \te{po:cleanText} property.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 \incgraph[
@@
 Each email snapshot is yield by an Email list.
-Interactions are yield through \textttt{po:replyTo} relations
-although \textttt{po:to} and \textttt{po:cc} might also be considered.
-The email body is given by \textttt{po:text} relations while
-\textttt{po:cleanText} links to text with lines removed where they are
-trivially from previous messages or computer code.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+Interactions are entailed by \te{po:replyTo} relations
+although \te{po:to} and \te{po:cc} may also be considered.
+The email body is given by \te{po:text} relations while
+\te{po:cleanText} links to text with lines removed where they are
+obviously from previous messages or computer code.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 \incgraph[
@@
 
-\section{ParticipaBR data}
-The ParticipaBR snapshot is yield by a data dump donated by the system
-administrators of the federal portal of social participation ParticipaBR.
-Articles can have parent articles (\textttt{po:parent}), be step of a
-collection of articles (\textttt{po:stepOf}) and be a mediation of other
-articles (\textttt{po:mediationOf}).
-Interactions are yield by comments which are \textttt{po:replyTo} other
+\section{Participa.BR data}
+The Participa.BR snapshot is yield by a data dump donated by the system
+administrators of the federal portal of social participation Participa.BR.
+Articles can have parent articles (\te{po:parent}), can be a step of a
+collection of articles (\te{po:stepOf}) and can be a mediation of other
+articles (\te{po:mediationOf}).
+Interactions are entailed by comments which are \te{po:replyTo} other
 comments or which are made directly to an article.
-This snapshot holds also friendship structures.
+This snapshot also holds friendship structures.
 The language used is mainly Brazilian Portuguese, but English and
 Spanish are also incident.
-Due to the higher complexity of the diagram, an additional figure is
-given rendered with another layout algorithm.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+Due to the higher complexity of the diagram, an additional figure,
+obtained through another layout algorithm, is given.
+Further information is found in the following diagrams, in the tables at 
+the end of this document or in the main document of this article~\cite{losd}.
 \incgraph[
   overlay={\node[red,below right] at (page.north west) {\Huge
-  ParticipaBR};}
+  Participa.BR};}
     paper=graphics
@@
   overlay={\node[red,below right] at (page.north west) {\Huge
-  ParticipaBR};}
+  Participa.BR};}
     paper=graphics
@@
 This snapshot holds a complex structure of both
-Topics/Inspirati-ons/Observatories/Supports/Competitions/Prizes
-and of State/City/Neighbor-hood/Place.
+Topics/Inspirations/Observatories/Supports/Competitions/Prizes
+and of State/City/Neighborhood/Place.
 The language used is mainly Brazilian Portuguese.
-Due to the higher complexity of the diagram, an additional figure is
-given rendered with another layout algorithm.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+Due to the higher complexity of the diagram, an additional figure,
+obtained through another layout algorithm, is given.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 \incgraph[
@@
 administrators and by a mined IRC log.
-The system pursue simplicity and most of data consists of detached
-shouts with \textttt{po:text} and \textttt{po:author}.
-Further information is found on the following diagrams, the tables on
-the end of this document or in the main document of this
-article~\cite{losd}.
+The system aims at simplicity and most of the data consists of
+shouts with \te{po:text} and \te{po:author}.
+Further information is found in the following diagrams, in the tables at
+the end of this document or in the main document of this article~\cite{losd}.
 \incgraph[
@@
 \begin{center}
-\caption{Different Twitter snapshots are yield by different hashtags. In this
-table we present each snapshot with the respective hashtag and a
+\caption{Different Twitter snapshots are yield by different hashtags.
+  This table presents each snapshot with the respective hashtag and a
 reference to the subject.}\label{tab:provenance}
@@
 \begin{center}
-\caption{Different IRC snapshots are yield by different channels. In this
-table we present each snapshot with the respective channel and a
+\caption{Different IRC snapshots are yield by different IRC channels.
+  This table presents each snapshot with the respective channel and a
 reference to the subject.}\label{tab:provenance}
@@
 \begin{center}
-\caption{Different Email snapshots are yield by different email lists. In this
-table we present each snapshot with the respective list and a
+\caption{Different Email snapshots are yield by different email lists.
+  This table presents each snapshot with the respective list and a
 reference to the subject.}\label{tab:provenance}
@@
 \begin{center}
-\caption{References for the snapshots of the detached instances
-ParticipaBR, Cidade Democrática and AA.}\label{tab:provenance}
+\caption{References for the snapshots of
+Participa.BR, Cidade Democrática and AA.}\label{tab:provenance}
 \begin{tabular}{| l || p{4cm} | p{3cm} | }\hline
     \textbf{social protocol} & \textbf{observations} & \textbf{reference} \\\hline\hline
-    ParticipaBR & a Brazilian federal portal of social participation & \url{http://www.participa.br/} \\\hline
+    Participa.BR & a Brazilian federal portal of social participation & \url{http://www.participa.br/} \\\hline
     Cidade Demorática & a Brazilian civil society portal of social participation & \url{http://www.cidadedemocratica.org.br/} \\\hline

### 2019-07-02 supplementaryDocument.tex
bib

@@
     Cidade Demorática & a Brazilian civil society portal of social participation & \url{http://www.cidadedemocratica.org.br/} \\\hline
-    AA & the Algorithmic Autoregulation software development methodology & \cite{aarticle} \\\hline
+    AA & the Algorithmic Autoregulation software development methodology & \cite{aarticle,aa2} \\\hline
 \end{tabular}\end{center}
